### 2019-12-25 README.Rmd
Init

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SampleSizePlanner

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of SampleSizePlanner is to ...

## Installation

You can install the released version of SampleSizePlanner from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("SampleSizePlanner")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(SampleSizePlanner)
## basic example code
```

What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:

```{r cars}
summary(cars)
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub!

### 2019-12-25 R/app_ui.R

@@ -1,0 +1,29 @@
+#' @import shiny
+app_ui <- function() {
+  tagList(
+    # Leave this function for adding external resources
+    golem_add_external_resources(),
+    # List the first level UI elements here 
+    fluidPage(
+      h1("SampleSizePlanner"),
+      mod_question_module_ui("question_module_ui_1")
+    )
+  )
+}
+
+#' @import shiny
+golem_add_external_resources <- function(){
+  
+  addResourcePath(
+    'www', system.file('app/www', package = 'SampleSizePlanner')
+  )
+ 
+  tags$head(
+    golem::activate_js(),
+    golem::favicon()
+    # Add here all the external resources
+    # If you have a custom.css in the inst/app/www
+    # Or for example, you can add shinyalert::useShinyalert() here
+    #tags$link(rel="stylesheet", type="text/css", href="www/custom.css")
+  )
+}

### 2019-12-25 R/mod_question_module.R

@@ -1,0 +1,46 @@
+# Module UI
+  
+#' @title   mod_question_module_ui and mod_question_module_server
+#' @description  A shiny Module.
+#'
+#' @param id shiny id
+#' @param input internal
+#' @param output internal
+#' @param session internal
+#'
+#' @rdname mod_question_module
+#'
+#' @keywords internal
+#' @export 
+#' @importFrom shiny NS tagList 
+mod_question_module_ui <- function(id){
+  ns <- NS(id)
+  tagList(
+
+  )
+}
+    
+# Module Server
+    
+#' @rdname mod_question_module
+#' @export
+#' @keywords internal
+    
+mod_question_module_server <- function(input, output, session, data){
+  ns <- session$ns
+  
+  output$question <- renderText({
+    
+  })
+  
+  # add reactive ui print question
+  # add reactive ui print button with option left under the question (use actionBttn)
+  # add reactive ui print button with option right under the question (use actionBttn)
+
+  }
+    
+## To be copied in the UI
+# mod_question_module_ui("question_module_ui_1")
+    
+## To be copied in the server
+# callModule(mod_question_module_server, "question_module_ui_1")

### 2019-12-25 dev/01_start.R

@@ -1,0 +1,65 @@
+# Building a Prod-Ready, Robust Shiny Application.
+# 
+# Each step is optional. 
+# 
+# 1 - On init
+# 
+## 1.1 - Fill the descripion & set options
+## 
+## Add information about the package that will contain your app
+
+golem::fill_desc(
+  pkg_name = "SampleSizePlanner", 
+  pkg_title = "Application to Select the Fitting Sample Size Planning Method", 
+  pkg_description = "The package contains a Shiny application that through several questions step by step helps researchers to find the sample size planning method that best fits their needs. The application also helps to determine the parameters needed to run the given method. Finally, the application helps researchers to report the outcome of their sample size plan.",  
+  author_first_name = "Marton", 
+  author_last_name = "Kovacs",  
+  author_email = "marton.balazs.kovacs@gmail.com",    
+  repo_url = NULL # The (optional) URL of the GitHub Repo
+)     
+
+## Use this desc to set {golem} options
+
+golem::set_golem_options()
+
+## 1.2 - Set common Files 
+## 
+## If you want to use the MIT licence, README, code of conduct, lifecycle badge, and news
+
+usethis::use_mit_license( name = "Marton Kovacs" )  # You can set another licence here
+usethis::use_readme_rmd( open = FALSE )
+# usethis::use_code_of_conduct()
+usethis::use_lifecycle_badge( "Experimental" )
+
+usethis::use_news_md( open = FALSE )
+usethis::use_git()
+
+## 1.3 - Add a data-raw folder
+## 
+## If you have data in your package
+# usethis::use_data_raw( name = "flowchart_data", open = FALSE )
+
+## 1.4 - Init Tests
+## 
+## Create a template for tests
+
+golem::use_recommended_tests()
+
+## 1.5 : Use Recommended Package
+
+golem::use_recommended_deps()
+
+## 1.6 Add various tools
+
+# If you want to change the favicon (default is golem's one)
+golem::remove_favicon()
+golem::use_favicon() # path = "path/to/ico". Can be an online file. 
+
+# Add helper functions 
+golem::use_utils_ui()
+golem::use_utils_server()
+
+# You're now set! 
+# go to dev/02_dev.R
+rstudioapi::navigateToFile( "dev/02_dev.R" )
+

### 2019-12-25 dev/run_dev.R

@@ -1,0 +1,12 @@
+# Set options here
+options(golem.app.prod = FALSE) # TRUE = production mode, FALSE = development mode
+
+# Detach all loaded packages and clean your environment
+golem::detach_all_attached()
+# rm(list=ls(all.names = TRUE))
+
+# Document and reload your package
+golem::document_and_reload()
+
+# Run the application
+SampleSizePlanner::run_app()

### 2019-12-25 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(SampleSizePlanner)
+
+test_check("SampleSizePlanner")

### 2020-05-06 inst/app/www/about.Rmd
working stage with messy (beautiful) UI
This is a working app with a nice UI that is not efficient according to some people. Therefore I need to make it boring. Note: not all the sample size calculation methods are working currently and not all of them are added to the app. flowchart still needs to be updated.

---
title: ""
output: html_document
---

### About the project

Important text here!

### How to use the application?

Important instructions here!

### Contributors
- Don van Ravenzwaaij
- Rink Hoekstra
- Marton Kovacs
- Balazs Aczel

### 2020-05-06 R/PowerOptim.R

@@ -1,0 +1,23 @@
+PowerOptim <- function (Fun, Range, Arguments, Opt) {
+  Ns = Range
+  Res = c(Fun(Ns[1], Arguments[1], Arguments[2]), 
+          Fun(Ns[2], Arguments[1], Arguments[2]))
+  if (!(Opt>min(Res)&Opt<max(Res))) {return (NULL)}
+  
+  while ((min(Ns[Res>Opt])-max(Ns[Res<Opt]))>1)
+  {
+    ResL = Res[which(Ns==max(Ns[Res<Opt]))]
+    ResH = Res[which(Ns==min(Ns[Res>Opt]))]
+    NL = max(Ns[Res<Opt])
+    NH = min(Ns[Res>Opt])
+    NewN = round((NH-NL)*((Opt-ResL)/(ResH-ResL))+NL)
+    while (NewN%in%Ns)
+    {
+      NewN = ifelse(Res[length(Res)]>Opt, NewN - 1, NewN + 1)
+    }
+    Ns = c(Ns, NewN)
+    Res = c(Res, Fun(Ns[length(Ns)], Arguments[1], Arguments[2]))
+  }
+
+  c(min(Ns[Res>Opt]), Res[which(Ns==min(Ns[Res>Opt]))])
+}

### 2020-05-06 R/ROPEpow.R

@@ -1,0 +1,31 @@
+ROPEpow <- function (n1, Band, delta, alpha = .05, Tol = 1e-4, granularity = 300)
+{
+  n2 = n1
+  t = seq (qt (.005, n1+n2-2, delta/sqrt(1/n1+1/n2)), 
+           qt (.995, n1+n2-2, delta/sqrt(1/n1+1/n2)), length = granularity)
+  Upper = Lower = 0.5
+  i = 0
+  while (!(Lower<(alpha/2) & Upper>(1-alpha/2)))
+  {
+    i = i+1
+    Upper = cdf_t(Band, t[i], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
+                   prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = Tol)
+    Lower = cdf_t(-Band, t[i], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
+                   prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = Tol)
+  }
+  Upper = Lower = 0.5
+  j = length (t)+1
+  while (!(Lower<(alpha/2) & Upper>(1-alpha/2)))
+  {
+    j = j-1
+    Upper = cdf_t(Band, t[j], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
+                   prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = Tol)
+    Lower = cdf_t(-Band, t[j], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
+                   prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = Tol)
+  }
+  npower = pt(t[j], n1+n2-2, delta/sqrt(1/n1+1/n2))-
+    pt(t[i], n1+n2-2, delta/sqrt(1/n1+1/n2))
+  
+  # cat ("\n n1 is ", n1, ", power is ", npower, sep = "")
+  npower
+}

### 2020-05-06 R/SeqBFpow.R

@@ -1,0 +1,26 @@
+SeqBFpow <- function (delta, thresh = 10, power = 0.8, nRep = 1000) {
+  Ns = BFs = c()
+  for (i in 1:nRep)
+  {
+    cat ("\n", i)
+    n1 = n2 = 4
+    Plac = rnorm (n1, 0, 1)
+    Treat = rnorm (n2, delta, 1)
+    BF = as.vector (ttestBF (x = Treat, y = Plac, nullInterval = c(0, Inf), rscale = 1/sqrt(2)))[1]
+    
+    while (BF<thresh & BF>(1/thresh) & n1<1500)
+    {
+      n1 = n2 = n1+1
+      Plac = c(Plac, rnorm (1, 0, 1))
+      Treat = c(Treat, rnorm (1, delta, 1))
+      BF = as.vector (ttestBF (x = Treat, y = Plac, nullInterval = c(0, Inf), rscale = 1/sqrt(2)))[1]
+    }
+    Ns[i] = n1
+    BFs[i] = BF
+  }
+  nopt = sort(Ns)[round (length(Ns)*power)]
+  BFHA = mean (BFs>thresh)
+  cat ("\n n1 is ", nopt, ", power is ", power, 
+       ", the proportion of times HA is supported is ", BFHA, sep = "")
+  c(nopt, BFHA)
+}

### 2020-05-06 R/SuperBFpow.R

@@ -1,0 +1,27 @@
+SuperBFpow <- function (n1, Band, delta, thresh = 10, Tol = 1e-4, granularity = 300) {
+  n2 = n1
+  t = seq(qt(.005, n1+n2-2, delta/sqrt(1/n1+1/n2)), 
+          qt(.995, n1+n2-2, delta/sqrt(1/n1+1/n2)), length = granularity)
+  BF = 1
+  i = 0
+  while (BF<thresh)
+  {
+    i = i+1
+    BF = exp(BayesFactor::ttest.tstat(t = t[i], n1 = n1, n2 = n2, nullInterval = c(0, Inf), rscale = 1/sqrt(2))[['bf']])
+  }
+  BF = 1
+  j = length (t)+1
+  while (BF<thresh)
+  {
+    j = j-1
+    BF = exp(BayesFactor::ttest.tstat(t = t[j], n1 = n1, n2 = n2, nullInterval = c(0, Inf), rscale = 1/sqrt(2))[['bf']])
+  }
+  npower = pt(t[j], n1+n2-2, delta/sqrt(1/n1+1/n2))-
+    pt(t[i], n1+n2-2, delta/sqrt(1/n1+1/n2))
+
+  output <-
+    tibble::tibble(n1 = npower[1],
+                   npower = npower[2])
+  
+  return(output)
+}

### 2020-05-06 R/TOSTss.R

@@ -1,4 +1,4 @@
-TOSTss <- function (Opt, Band, delta, sigma = 1, nr = 1, alpha = .05){
+TOSTss <- function(Opt, Band, delta, sigma = 1, nr = 1, alpha = .05){
   
   n1 = 4
   sigsq = sigma^2
@@ -8,7 +8,7 @@ TOSTss <- function (Opt, Band, delta, sigma = 1, nr = 1, alpha = .05){
   coevecc = c(1, rep(c(4, 2), numint / 2 - 1), 4, 1)
   npower = 0
   
-  while (npower < Opt & n1 < 10001){
+  while(npower < Opt & n1 < 10001){
     n1 = n1 + 1
     n2 = nr * n1
     df = n1 + n2 - 2

### 2020-05-06 R/app_server.R

@@ -1,8 +1,42 @@
 #' @import shiny
 app_server <- function(input, output,session) {
-  # List the first level callModules here
-  callModule(mod_article_module_server, "article_module_ui_1")
-  mod_menu_output <- callModule(mod_menu_module_server, "menu_module_ui_1")
-  mod_question_output <- callModule(mod_question_module_server, "question_module_ui_1")
-  callModule(mod_tost_module_server, "tost_module_ui_1", input_data = mod_question_output$question, menu_call = mod_menu_output$method)
+  mod_article_module_server("article_module_ui_1")
+  mod_about_server("about_ui_1")
+  menu_out <- mod_menu_module_server("menu_module_ui_1")
+  question_out <- mod_question_module_server("question_module_ui_1")
+  mod_tost_module_server("tost_module_ui_1",
+                         activate_menu = menu_out$activate,
+                         method_menu = menu_out$method,
+                         activate_question = question_out$activate,
+                         method_question = question_out$method)
+  mod_traditional_power_server("traditional_power_ui_1",
+                               activate_menu = menu_out$activate,
+                               method_menu = menu_out$method,
+                               activate_question = question_out$activate,
+                               method_question = question_out$method)
+  mod_power_curve_server("power_curve_ui_1",
+                         activate_menu = menu_out$activate,
+                         method_menu = menu_out$method,
+                         activate_question = question_out$activate,
+                         method_question = question_out$method)
+  mod_apriori_precision_server("apriori_precision_ui_1",
+                               activate_menu = menu_out$activate,
+                               method_menu = menu_out$method,
+                               activate_question = question_out$activate,
+                               method_question = question_out$method)
+  mod_rope_server("rope_ui_1",
+                  activate_menu = menu_out$activate,
+                  method_menu = menu_out$method,
+                  activate_question = question_out$activate,
+                  method_question = question_out$method)
+  mod_interval_equiv_bf_server("interval_equiv_bf_ui_1",
+                               activate_menu = menu_out$activate,
+                               method_menu = menu_out$method,
+                               activate_question = question_out$activate,
+                               method_question = question_out$method)
+  mod_bf_threshold_server("bf_threshold_ui_1",
+                          activate_menu = menu_out$activate,
+                          method_menu = menu_out$method,
+                          activate_question = question_out$activate,
+                          method_question = question_out$method)
 }

### 2020-05-06 R/app_ui.R

@@ -3,11 +3,12 @@ app_ui <- function() {
   tagList(
     # Leave this function for adding external resources
     golem_add_external_resources(),
-    # Using the package: shinyjs
-    shinyjs::useShinyjs(),
+    
     # List the first level UI elements here 
     fluidPage(
-      h1("SampleSizePlanner"),
+      fluidRow(
+      h1("Sample Size Planner"), align = "center", style = "color: #FF6356;"),
+      mod_about_ui("about_ui_1"),
       mod_article_module_ui("article_module_ui_1"),
       mod_menu_module_ui("menu_module_ui_1"),
       mod_question_module_ui("question_module_ui_1"),
@@ -25,7 +26,13 @@ golem_add_external_resources <- function(){
  
   tags$head(
     golem::activate_js(),
-    golem::favicon()
+    golem::favicon(),
+    # Using the package: shinyjs
+    shinyjs::useShinyjs(),
+    tags$style("#question_module_ui_1-question{color: #FF6356;
+                                               font-size: 30px;
+                                               font-style: bold;
+               }")
     # Add here all the external resources
     # If you have a custom.css in the inst/app/www
     # Or for example, you can add shinyalert::useShinyalert() here

### 2020-05-06 R/cdf_t.R

@@ -1,0 +1,31 @@
+cdf_t <- function(x,
+                  t,
+                  n1,
+                  n2 = NULL,
+                  independentSamples = FALSE,
+                  prior.location,
+                  prior.scale,
+                  prior.df,
+                  rel.tol = .Machine$double.eps^0.25) {
+  
+  out <- integrate(posterior_t,
+                   lower = -Inf, upper = x,
+                   t = t, n1 = n1, n2 = n2,
+                   independentSamples = independentSamples,
+                   prior.location = prior.location,
+                   prior.scale = prior.scale,
+                   prior.df = prior.df,
+                   rel.tol = rel.tol)$value
+  
+  # catch numerical errors
+  if (out > 1 & out < 1.001) {
+    out <- 1
+    warning(
+      "Numerical integration yields a CDF value slightly larger than 1. The CDF value has been replaced by 1.",
+      call. = FALSE
+    )
+  }
+  
+  return(out)
+  
+}

### 2020-05-06 R/mod_interval_equiv_bf.R

@@ -1,0 +1,93 @@
+# Module UI
+  
+#' @title   mod_interval_equiv_bf_ui and mod_interval_equiv_bf_server
+#' @description  A shiny Module.
+#'
+#' @param id shiny id
+#' @param input internal
+#' @param output internal
+#' @param session internal
+#'
+#' @rdname mod_interval_equiv_bf
+#'
+#' @keywords internal
+#' @export 
+#' @importFrom shiny NS tagList 
+mod_interval_equiv_bf_ui <- function(id){
+  
+  tagList(
+  
+  )
+}
+    
+# Module Server
+    
+#' @rdname mod_interval_equiv_bf
+#' @export
+#' @keywords internal
+    
+mod_interval_equiv_bf_server <- function(id, activate_menu, method_menu, activate_question, method_question){
+  
+  moduleServer(id, function(input, output, session) {
+    
+    modal <- function() {
+      
+      modalDialog(
+        footer = actionButton(NS(id, "close_modal"), label = "Close"),
+        h1("Interval Equiv BF"),
+        h2("Effect size"),
+        sliderInput(NS(id, "delta"), "Delta", min = 0, max = 1, value = 0, step = 0.1),
+        h2("Threshold"),
+        selectInput(NS(id, "tresh"), "Threshold", choices = c(3, 6, 10),
+                    selected = 10),
+        h2("Prior"),
+        selectInput(NS(id, "distribution"), "Choose a distribution:",
+                    choices = c("cdf_t", "cdf_normal"),
+                    selected = "cdf_t"),
+        selectInput(NS(id, "scale"), "Choose a scale:",
+                    choices = "1/sqrt(2)",
+                    selected = "1/sqrt(2)"),
+        numericInput(NS(id, "location"), "Choose a location:", value = 0),
+        numericInput(NS(id, "df"), "Choose a df:", value = 1),
+        actionButton(NS(id, "calculate"), "Ready, set, go!"),
+        textOutput(NS(id, "output$eq_bf_output"))
+      )
+    }
+    
+    observeEvent(activate_menu(), {
+      
+      if(method_menu() == "Interval Equiv BF") {
+        showModal(modal())}
+      
+    })
+    
+    observeEvent(activate_question(), {
+      
+      if(method_question() == "Interval Equiv BF") {
+        showModal(modal())}
+      
+    })
+    
+    observeEvent(input$close_modal, {
+      
+      removeModal()
+      
+    })
[…]

### 2020-05-06 R/mod_question_module.R

@@ -14,28 +14,31 @@
 #' @export 
 #' @importFrom shiny NS tagList 
 mod_question_module_ui <- function(id){
-  ns <- NS(id)
-  
+
   tagList(
     fixedPanel(
-      textOutput(ns("question")),
+      textOutput(NS(id, "question")),
       bottom = "50%", left = "40%", width = "auto"
     ),
     fixedPanel(
-      uiOutput(ns("left_button")),
+      uiOutput(NS(id, "left_button")),
       bottom = "30%", left = "10%", width = "auto"
     ),
     fixedPanel(
-      uiOutput(ns("right_button")),
+      uiOutput(NS(id, "right_button")),
       bottom = "30%", right = "10%", width = "auto"
     ),
     fixedPanel(
       shinyWidgets::actionBttn(
-        inputId = ns("over"),
+        inputId = NS(id, "over"),
         label = "Start over",
         style = "minimal",
         color = "danger"),
-      top = "5%", right = "2%", width = "auto"
+      top = "8%", right = "4%", width = "auto"
+    ),
+    fixedPanel(
+      uiOutput(NS(id, "open")),
+      bottom = "30%", right = "45%", left = "45%", width = "auto"
     )
   )
 }
@@ -46,9 +49,9 @@ mod_question_module_ui <- function(id){
 #' @export
 #' @keywords internal
     
-mod_question_module_server <- function(input, output, session){
+mod_question_module_server <- function(id){
   
-  ns <- session$ns
+  moduleServer(id, function(input, output, session) {
   
   current <- reactiveVal(value = dplyr::filter(question_data, id == 1))
   
@@ -97,18 +100,29 @@ mod_question_module_server <- function(input, output, session){
     }
   })
   
+  output$open <- renderUI({
+    if(current()$is_method == 1L) {
+      shinyWidgets::actionBttn(
+        inputId = session$ns("open_modal"),
+        label = "Show me more!",
+        style = "minimal",
+        color = "danger")
+      }
+  })
+  
   output$question <- renderText({
     current()$question
   })
   
-  return(list(
-    question = reactive({current()$question})
-  ))
-
+  list(
+    method = reactive(current()$question),
+    activate = reactive(input$open_modal)
+    )
+  })
   }
     
 ## To be copied in the UI
 # mod_question_module_ui("question_module_ui_1")
     
 ## To be copied in the server
-# callModule(mod_question_module_server, "question_module_ui_1")
+# mod_question_module_server("question_module_ui_1")

### 2020-05-06 R/mod_tost_module.R

@@ -14,13 +14,8 @@
 #' @export 
 #' @importFrom shiny NS tagList 
 mod_tost_module_ui <- function(id){
-  
-  ns <- NS(id)
+
   tagList(
-    fixedPanel(
-      uiOutput(ns("open")),
-      bottom = "30%", right = "50%", width = "auto"
-    )
   )
 }
     
@@ -30,38 +25,41 @@ mod_tost_module_ui <- function(id){
 #' @export
 #' @keywords internal
     
-mod_tost_module_server <- function(input, output, session, input_data, menu_call){
+mod_tost_module_server <- function(id, activate_menu, method_menu, activate_question, method_question){
   
+  moduleServer(id, function(input, output, session) {
   modal <- function() {
-    ns <- session$ns
     
     modalDialog(
-      easyClose = TRUE,
-      footer = modalButton("Close Modal"),
+      footer = actionButton(NS(id, "close_modal"), label = "Close modal"),
       h1("Let's have a TOAST!"),
-      sliderInput(ns("opt"), "Power", min = 0, max = 1, value = 0.8, step = 0.1),
-      sliderInput(ns("band"), "Band", min = 0, max = 1, value = 0.2, step = 0.1),
-      sliderInput(ns("delta"), "Delta", min = 0, max = 1, value = 0, step = 0.1),
-      actionButton(ns("calculate"), "Ready, set, go!"),
-      textOutput(ns("tost_output_n1")),
-      textOutput(ns("tost_output_npower")),
-      textOutput(ns("error"))
+      sliderInput(NS(id, "opt"), "Power", min = 0, max = 1, value = 0.8, step = 0.1),
+      sliderInput(NS(id, "band"), "Band", min = 0, max = 1, value = 0.2, step = 0.1),
+      sliderInput(NS(id, "delta"), "Delta", min = 0, max = 1, value = 0, step = 0.1),
+      actionButton(NS(id, "calculate"), "Ready, set, go!"),
+      textOutput(NS(id, "tost_output_n1")),
+      textOutput(NS(id, "tost_output_npower"))
     )
   }
   
-  observeEvent(input$open_modal, {
-    showModal(modal())},
-    ignoreInit = TRUE)
+  observeEvent(activate_menu(), {
+    
+    if(method_menu() == "TOST") {
+    showModal(modal())}
+    
+    })
   
+  observeEvent(activate_question(), {
+    
+    if(method_question() == "TOST") {
+      showModal(modal())}
+    
+  })
   
-  output$open <- renderUI({
[…]

### 2020-05-11 inst/app/www/on_load_placeholder.Rmd
working plotly version

---
title: ''
output: html_document
---

+ Please set the parameters and run the sample size planning calculation with the "Calculate sample size" button.
+ After running the calculation you can download the results in a general template text in word format.
+ After donwload, please complement the text where it is need with your reasoning behind each decision and include the text in your manuscript.

### 2020-05-11 R/mod_power_curve.R

@@ -14,9 +14,17 @@
 #' @export 
 #' @importFrom shiny NS tagList 
 mod_power_curve_ui <- function(id){
-
   tagList(
-  
+    sidebarPanel(
+      h1("Power curve"),
+      sliderInput(NS(id, "opt"), "Power", min = 0, max = 1, value = 0.8, step = 0.1),
+      sliderInput(NS(id, "delta"), "Delta", min = 0, max = 1, value = c(0.1, 0.9), step = 0.1),
+      actionButton(NS(id, "calculate"), "Calculate sample size ")),
+    mainPanel(
+      plotly::plotlyOutput(NS(id, "power_curve_plot"))
+      # div(id = "download_power_curve",
+      #     downloadButton(NS(id, "report")))
+    )
   )
 }
     
@@ -29,60 +37,59 @@ mod_power_curve_ui <- function(id){
 mod_power_curve_server <- function(id, activate_menu, method_menu, activate_question, method_question){
 
     moduleServer(id, function(input, output, session) {
-    modal <- function() {
       
-      modalDialog(
-        footer = actionButton(NS(id, "close_modal"), label = "Close"),
-        size = "m",
-        h1("Power curve"),
-        sliderInput(NS(id, "opt"), "Power", min = 0, max = 1, value = 0.8, step = 0.1),
-        sliderInput(NS(id, "delta"), "Delta", min = 0, max = 1, value = c(0.1, 0.9), step = 0.1),
-        actionButton(NS(id, "calculate"), "Ready, set, go!"),
-        plotOutput(NS(id, "power_curve_plot"))
-      )
-    }
-    
-    observeEvent(activate_menu(), {
+      waitress <- waiter::Waitress$new("#tost_ui_1-tost_output", theme = "overlay", infinite = TRUE, hide_on_render = TRUE)
       
-      if(method_menu() == "Power curve") {
-        showModal(modal())}
+      observe({
+        if (input$calculate) {
+          shinyjs::enable("report")
+          shinyjs::runjs("$('#power_curve_plot').removeAttr('title');")
+        } else{
+          shinyjs::disable("report")
+          shinyjs::runjs("$('#power_curve_plot').attr('title', 'Please run the calculation first');")
+        }
+      })
       
-    })
-    
-    observeEvent(activate_question(), {
-      
-      if(method_question() == "Power curve") {
-        showModal(modal())}
-      
-    })
-    
-    observeEvent(input$close_modal, {
-      
-      removeModal()
-      
-    })
-    
[…]

### 2020-06-28 inst/app/www/aipe_output.Rmd
working baseline version

---
title: "&nbsp;"
output:
  html_document: default
  word_document: default
params:
  n_min: NA
  n_max: NA
  output_n: NA
  input_power: NA
  output_confidence: NA
  band: NA
  delta: NA
---

```{r, include = FALSE}
library(knitr)
```

In order to calculate the sample size we choose the accuracy in parameter estimation (AIPE; REF) method. We choose the expected effect size to be `r expression (delta)` = `r params$delta`, because ... The desired power was `r params$input_power` as ... Because of limits in resources the minimum sample size of interest was `r params$n_min` and the maximum sample size of interest was `r params$n_max`. The resulting confidence of the sample size estimation method was `r params$output_confidence` with the estimated sample size of `r params$output_n`.

### 2020-06-28 R/EqBFpow.R

@@ -1,39 +1,0 @@
-EqBFpow = function (n1, Band, delta, thresh = 10, Tol = 1e-4, granularity = 300)
-{
-  n2 = n1
-  t = seq (qt (.005, n1+n2-2, delta/sqrt(1/n1+1/n2)), 
-           qt (.995, n1+n2-2, delta/sqrt(1/n1+1/n2)), length = granularity)
-  BF = 1
-  i = 0
-  while (BF<thresh)
-  {
-    i = i+1
-    Upper = cdf_t (Band, t[i], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
-                   prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = Tol)
-    Lower = cdf_t (-Band, t[i], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
-                   prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = Tol)
-    PostDens = Upper - Lower
-    PriorDens = pcauchy (Band, scale = 1/sqrt(2)) - 
-      pcauchy (-Band, scale = 1/sqrt(2))
-    BF = (PostDens / PriorDens) / ((1-PostDens) / (1-PriorDens))
-  }
-  BF = 1
-  j = length (t)+1
-  while (BF<thresh)
-  {
-    j = j-1
-    Upper = cdf_t (Band, t[j], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
-                   prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = Tol)
-    Lower = cdf_t (-Band, t[j], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
-                   prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = Tol)
-    PostDens = Upper - Lower
-    PriorDens = pcauchy (Band, scale = 1/sqrt(2)) - 
-      pcauchy (-Band, scale = 1/sqrt(2))
-    BF = (PostDens / PriorDens) / ((1-PostDens) / (1-PriorDens))
-  }
-  npower = pt(t[j], n1+n2-2, delta/sqrt(1/n1+1/n2))-
-    pt(t[i], n1+n2-2, delta/sqrt(1/n1+1/n2))
-  
-  # cat ("\n n1 is ", n1, ", power is ", npower, sep = "")
-  npower
-}

### 2020-06-28 R/PowerOptim.R

@@ -1,23 +1,0 @@
-PowerOptim <- function (Fun, Range, Arguments, Opt) {
-  Ns = Range
-  Res = c(Fun(Ns[1], Arguments[1], Arguments[2]), 
-          Fun(Ns[2], Arguments[1], Arguments[2]))
-  if (!(Opt>min(Res)&Opt<max(Res))) {return (NULL)}
-  
-  while ((min(Ns[Res>Opt])-max(Ns[Res<Opt]))>1)
-  {
-    ResL = Res[which(Ns==max(Ns[Res<Opt]))]
-    ResH = Res[which(Ns==min(Ns[Res>Opt]))]
-    NL = max(Ns[Res<Opt])
-    NH = min(Ns[Res>Opt])
-    NewN = round((NH-NL)*((Opt-ResL)/(ResH-ResL))+NL)
-    while (NewN%in%Ns)
-    {
-      NewN = ifelse(Res[length(Res)]>Opt, NewN - 1, NewN + 1)
-    }
-    Ns = c(Ns, NewN)
-    Res = c(Res, Fun(Ns[length(Ns)], Arguments[1], Arguments[2]))
-  }
-
-  c(min(Ns[Res>Opt]), Res[which(Ns==min(Ns[Res>Opt]))])
-}

### 2020-06-28 R/SeqBFpow.R

@@ -1,26 +1,0 @@
-SeqBFpow <- function (delta, thresh = 10, power = 0.8, nRep = 1000) {
-  Ns = BFs = c()
-  for (i in 1:nRep)
-  {
-    cat ("\n", i)
-    n1 = n2 = 4
-    Plac = rnorm (n1, 0, 1)
-    Treat = rnorm (n2, delta, 1)
-    BF = as.vector (ttestBF (x = Treat, y = Plac, nullInterval = c(0, Inf), rscale = 1/sqrt(2)))[1]
-    
-    while (BF<thresh & BF>(1/thresh) & n1<1500)
-    {
-      n1 = n2 = n1+1
-      Plac = c(Plac, rnorm (1, 0, 1))
-      Treat = c(Treat, rnorm (1, delta, 1))
-      BF = as.vector (ttestBF (x = Treat, y = Plac, nullInterval = c(0, Inf), rscale = 1/sqrt(2)))[1]
-    }
-    Ns[i] = n1
-    BFs[i] = BF
-  }
-  nopt = sort(Ns)[round (length(Ns)*power)]
-  BFHA = mean (BFs>thresh)
-  cat ("\n n1 is ", nopt, ", power is ", power, 
-       ", the proportion of times HA is supported is ", BFHA, sep = "")
-  c(nopt, BFHA)
-}

### 2020-06-28 R/SuperBFpow.R

@@ -1,27 +1,0 @@
-SuperBFpow <- function (n1, Band, delta, thresh = 10, Tol = 1e-4, granularity = 300) {
-  n2 = n1
-  t = seq(qt(.005, n1+n2-2, delta/sqrt(1/n1+1/n2)), 
-          qt(.995, n1+n2-2, delta/sqrt(1/n1+1/n2)), length = granularity)
-  BF = 1
-  i = 0
-  while (BF<thresh)
-  {
-    i = i+1
-    BF = exp(BayesFactor::ttest.tstat(t = t[i], n1 = n1, n2 = n2, nullInterval = c(0, Inf), rscale = 1/sqrt(2))[['bf']])
-  }
-  BF = 1
-  j = length (t)+1
-  while (BF<thresh)
-  {
-    j = j-1
-    BF = exp(BayesFactor::ttest.tstat(t = t[j], n1 = n1, n2 = n2, nullInterval = c(0, Inf), rscale = 1/sqrt(2))[['bf']])
-  }
-  npower = pt(t[j], n1+n2-2, delta/sqrt(1/n1+1/n2))-
-    pt(t[i], n1+n2-2, delta/sqrt(1/n1+1/n2))
-
-  output <-
-    tibble::tibble(n1 = npower[1],
-                   npower = npower[2])
-  
-  return(output)
-}

### 2020-06-28 R/TOSTss.R

@@ -1,34 +1,0 @@
-TOSTss <- function(Opt, Band, delta, sigma = 1, nr = 1, alpha = .05){
-  
-  n1 = 4
-  sigsq = sigma^2
-  numint = 1000
-  lc = numint + 1
-  cl = 1e-10
-  coevecc = c(1, rep(c(4, 2), numint / 2 - 1), 4, 1)
-  npower = 0
-  
-  while(npower < Opt & n1 < 10001){
-    n1 = n1 + 1
-    n2 = nr * n1
-    df = n1 + n2 - 2
-    tcrit = qt(1 - alpha, df)
-    nfac = 1/n1 + 1/n2
-    var = sigsq * nfac
-    std = sqrt(var)
-    cu = (df * Band^2) / (var * tcrit^2)
-    int = cu - cl
-    intl = int / numint
-    cvec = cl + intl * (0:numint)
-    wcpdf = (intl / 3) * coevecc * dchisq(cvec, df)
-    st = sqrt(cvec / df) * tcrit
-    npower = sum(wcpdf * (pnorm((Band - delta) /std - st) - pnorm((-Band - delta) / std + st)))
-  }
-  
-  output <-
-    tibble::tibble(n1 = round(n1, 4),
-                   n2 = round(n2, 4),
-                   npower = round(npower, 4))
-  
-  return(output)
-}

### 2020-06-28 R/mod_about.R

@@ -16,13 +16,7 @@
 mod_about_ui <- function(id){
   
   tagList(
-    fixedPanel(
-      shinyWidgets::actionBttn(inputId = NS(id, "open_about"),
-                               label = "How to use",
-                               style = "minimal",
-                               color = "danger"),
-      top = "8%", left = "4%", width = "auto"
-    )
+    includeMarkdown("inst/app/www/about.Rmd")
   )
 }
     
@@ -35,17 +29,8 @@ mod_about_ui <- function(id){
 mod_about_server <- function(id){
   
   moduleServer(id, function(input, output, session) {
-  modal <- function() {
+
     
-    modalDialog(
-      easyClose = TRUE,
-      footer = modalButton("Close Modal"),
-      includeMarkdown("inst/app/www/about.Rmd"))
-    }
-  
-  observeEvent(input$open_about, {
-    showModal(modal())})
-  
   })
   }
     

### 2020-06-28 R/mod_download.R

@@ -1,0 +1,59 @@
+#' download UI Function
+#'
+#' @description A shiny Module.
+#'
+#' @param id,input,output,session Internal parameters for {shiny}.
+#'
+#' @noRd 
+#'
+#' @importFrom shiny NS tagList 
+mod_download_ui <- function(id){
+  tagList(
+    div(class = "download-btn",
+        downloadButton(NS(id, "report")))
+  )
+}
+    
+#' download Server Function
+#'
+#' @noRd 
+mod_download_server <- function(id, activate, input_file, params, format = "word_document"){
+  moduleServer(id, function(input, output, session) {
+  # Create output file extension based on the format
+  ext <- switch(format,
+                word_document = ".doc",
+                html_document = ".html")
+  
+  # Add downloadbutton enable logic
+  observe({
+    if (activate()) {
+      shinyjs::enable("report")
+      shinyjs::runjs("$('.download-btn').removeAttr('title');")
+    } else{
+      shinyjs::disable("report")
+      shinyjs::runjs("$('.download-btn').attr('title', 'Please run the calculation first');")
+    }
+  })
+  
+  # Download the output
+  output$report <- downloadHandler(
+    filename = function() {
+      paste0(tools::file_path_sans_ext(input_file), "_", Sys.Date(), ext)
+    },
+    content = function(file) {
+      report_path <- file.path("inst/app/www/", input_file)
+      file.copy(input_file, report_path, overwrite = TRUE)
+      callr::r(
+        render_report,
+        list(input = report_path, output = file, format = format, params = params())
+      )
+    })
+  })
+}
+    
+## To be copied in the UI
+# mod_download_ui("download")
+    
+## To be copied in the server
+# mod_download_server("download")
+ 

### 2020-06-28 R/mod_flowchart_module.R

@@ -1,54 +1,0 @@
-# Module UI
-  
-#' @title   mod_flowchart_module_ui and mod_flowchart_module_server
-#' @description  A shiny Module.
-#'
-#' @param id shiny id
-#' @param input internal
-#' @param output internal
-#' @param session internal
-#'
-#' @rdname mod_flowchart_module
-#'
-#' @keywords internal
-#' @export 
-#' @importFrom shiny NS tagList 
-mod_flowchart_module_ui <- function(id){
-  
-  tagList(
-    shinyglide::screen(
-      DiagrammeR::grVizOutput(NS(id, "dg"))
-    ),
-    fixedPanel(
-      shinyWidgets::actionBttn(
-        inputId = NS(id, "flowchart"),
-        label = "Open flowchart", 
-        style = "minimal",
-        color = "danger"),
-      bottom = "5%", left = "6%", width = "auto"
-    )
-  )
-}
-    
-# Module Server
-    
-#' @rdname mod_flowchart_module
-#' @export
-#' @keywords internal
-    
-mod_flowchart_module_server <- function(id){
-  moduleServer(id, function(input, output, session) {
-  
-  graph <- open_graph("inst/app/www/flowchart.dgr")
-  
-  output$dg <- renderGrViz({
-    render_graph(graph, layout = "neato")
-  })
-  })
-}
-    
-## To be copied in the UI
-# mod_flowchart_module_ui("flowchart_module_ui_1")
-    
-## To be copied in the server
-# mod_flowchart_module_server("flowchart_module_ui_1")

### 2020-06-28 R/mod_interval_equiv_bf.R

@@ -1,93 +1,0 @@
-# Module UI
-  
-#' @title   mod_interval_equiv_bf_ui and mod_interval_equiv_bf_server
-#' @description  A shiny Module.
-#'
-#' @param id shiny id
-#' @param input internal
-#' @param output internal
-#' @param session internal
-#'
-#' @rdname mod_interval_equiv_bf
-#'
-#' @keywords internal
-#' @export 
-#' @importFrom shiny NS tagList 
-mod_interval_equiv_bf_ui <- function(id){
-  
-  tagList(
-  
-  )
-}
-    
-# Module Server
-    
-#' @rdname mod_interval_equiv_bf
-#' @export
-#' @keywords internal
-    
-mod_interval_equiv_bf_server <- function(id, activate_menu, method_menu, activate_question, method_question){
-  
-  moduleServer(id, function(input, output, session) {
-    
-    modal <- function() {
-      
-      modalDialog(
-        footer = actionButton(NS(id, "close_modal"), label = "Close"),
-        h1("Interval Equiv BF"),
-        h2("Effect size"),
-        sliderInput(NS(id, "delta"), "Delta", min = 0, max = 1, value = 0, step = 0.1),
-        h2("Threshold"),
-        selectInput(NS(id, "tresh"), "Threshold", choices = c(3, 6, 10),
-                    selected = 10),
-        h2("Prior"),
-        selectInput(NS(id, "distribution"), "Choose a distribution:",
-                    choices = c("cdf_t", "cdf_normal"),
-                    selected = "cdf_t"),
-        selectInput(NS(id, "scale"), "Choose a scale:",
-                    choices = "1/sqrt(2)",
-                    selected = "1/sqrt(2)"),
-        numericInput(NS(id, "location"), "Choose a location:", value = 0),
-        numericInput(NS(id, "df"), "Choose a df:", value = 1),
-        actionButton(NS(id, "calculate"), "Ready, set, go!"),
-        textOutput(NS(id, "output$eq_bf_output"))
-      )
-    }
-    
-    observeEvent(activate_menu(), {
-      
-      if(method_menu() == "Interval Equiv BF") {
-        showModal(modal())}
-      
-    })
-    
-    observeEvent(activate_question(), {
-      
-      if(method_question() == "Interval Equiv BF") {
-        showModal(modal())}
-      
-    })
-    
-    observeEvent(input$close_modal, {
-      
-      removeModal()
-      
-    })
-    
[…]

### 2020-06-28 R/mod_power_curve.R

@@ -1,103 +1,0 @@
-# Module UI
-  
-#' @title   mod_power_curve_ui and mod_power_curve_server
-#' @description  A shiny Module.
-#'
-#' @param id shiny id
-#' @param input internal
-#' @param output internal
-#' @param session internal
-#'
-#' @rdname mod_power_curve
-#'
-#' @keywords internal
-#' @export 
-#' @importFrom shiny NS tagList 
-mod_power_curve_ui <- function(id){
-  tagList(
-    sidebarPanel(
-      h1("Power curve"),
-      sliderInput(NS(id, "opt"), "Power", min = 0, max = 1, value = 0.8, step = 0.1),
-      sliderInput(NS(id, "delta"), "Delta", min = 0, max = 1, value = c(0.1, 0.9), step = 0.1),
-      actionButton(NS(id, "calculate"), "Calculate sample size ")),
-    mainPanel(
-      plotly::plotlyOutput(NS(id, "power_curve_plot"))
-      # div(id = "download_power_curve",
-      #     downloadButton(NS(id, "report")))
-    )
-  )
-}
-    
-# Module Server
-    
-#' @rdname mod_power_curve
-#' @export
-#' @keywords internal
-    
-mod_power_curve_server <- function(id, activate_menu, method_menu, activate_question, method_question){
-
-    moduleServer(id, function(input, output, session) {
-      
-      waitress <- waiter::Waitress$new("#tost_ui_1-tost_output", theme = "overlay", infinite = TRUE, hide_on_render = TRUE)
-      
-      observe({
-        if (input$calculate) {
-          shinyjs::enable("report")
-          shinyjs::runjs("$('#power_curve_plot').removeAttr('title');")
-        } else{
-          shinyjs::disable("report")
-          shinyjs::runjs("$('#power_curve_plot').attr('title', 'Please run the calculation first');")
-        }
-      })
-      
-      power_curve_result <- eventReactive(input$calculate, {
-        Band <- NA
-        noptim <- NULL
-        delta <- seq(input$delta[1], input$delta[2], 0.01)
-        for (i in 1:length(delta)) {
-          noptim[i] = PowerOptim(Fun = Tpow, Range = c(5, 2000), Arguments = c(Band, delta[i]), Opt = input$opt)$n1
-          }
-        return(
-          list(
-            noptim = noptim,
-            delta = delta
-            )
-          )
-        })
-
-      # params <- eventReactive(input$calculate, {
-      #   list(
-      #   )
-      # })
[…]

### 2020-06-28 R/mod_ssp_bf_thresh.R

@@ -1,0 +1,78 @@
+# Module UI
+  
+#' @title   mod_ssp_bf_thresh_ui and mod_ssp_bf_thresh_server
+#' @description  A shiny Module.
+#'
+#' @param id shiny id
+#' @param input internal
+#' @param output internal
+#' @param session internal
+#'
+#' @rdname mod_ssp_bf_thresh
+#'
+#' @keywords internal
+#' @export 
+#' @importFrom shiny NS tagList 
+mod_ssp_bf_thresh_ui <- function(id){
+  tagList(
+    sidebarLayout(
+    sidebarPanel(
+      h1("Decide BF threshold"),
+      sliderInput(NS(id, "delta"), "Delta", min = 0, max = 2, value = 0.5, step = 0.1),
+      selectInput(NS(id, "thresh"), "Threshold", choices = c(10, 6, 3), selected = 10),
+      actionButton(NS(id, "calculate"), "Calculate sample size")),
+    mainPanel(
+      mod_preview_ui(NS(id, "preview")),
+      mod_download_ui(NS(id, "download"))
+      )
+    )
+  )
+}
+    
+# Module Server
+    
+#' @rdname mod_ssp_bf_thresh
+#' @export
+#' @keywords internal
+    
+mod_ssp_bf_thresh_server <- function(id){
+  
+  moduleServer(id, function(input, output, session) {
+    # Setup loadingbar
+    waitress <- waiter::Waitress$new("#bf_thresh-preview-show_preview", theme = "overlay", infinite = TRUE, hide_on_render = TRUE)
+    
+    # Calculate results
+    bf_thresh_result <- eventReactive(input$calculate, {
+      Waitress$start()
+      ssp_bf_thresh(opt = 0.8,
+                      band = 0.2,
+                      delta = input$delta,
+                      thresh = as.integer(input$thresh))
+      })
+    
+    # Send params to RMD
+    params <- eventReactive(input$calculate, {
+      list(
+        n1 = bf_thresh_result()$n1,
+        output_power = bf_thresh_result()$npower,
+        input_power = 0.8,
+        band = 0.2,
+        delta = input$delta,
+        thresh = input$thresh
+      )
+    })
+    
+    # Render preview
+    mod_preview_server("preview", activate = reactive(input$calculate), input_file = "bf_thresh_output.Rmd", params = params)
+    
+    # Download the output
[…]

### 2020-06-28 R/mod_ssp_power_traditional.R

@@ -1,0 +1,79 @@
+# Module UI
+  
+#' @title   mod_ssp_power_traditional_ui and mod_ssp_power_traditional_server
+#' @description  A shiny Module.
+#'
+#' @param id shiny id
+#' @param input internal
+#' @param output internal
+#' @param session internal
+#'
+#' @rdname mod_ssp_power_traditional
+#'
+#' @keywords internal
+#' @export 
+#' @importFrom shiny NS tagList 
+mod_ssp_power_traditional_ui <- function(id){
+  tagList(
+    sidebarLayout(
+    sidebarPanel(
+      h1("Traditional Power"),
+      sliderInput(NS(id, "opt"), "Power", min = 0, max = 1, value = 0.8, step = 0.1),
+      sliderInput(NS(id, "delta"), "Delta", min = 0, max = 1, value = 0.5, step = 0.1),
+      numericInput(NS(id, "n_min"), "Minimum sample size", 20),
+      numericInput(NS(id, "n_max"), "Maximum sample size", 200),
+      actionButton(NS(id, "calculate"), "Calculate sample size")),
+    mainPanel(
+      mod_preview_ui(NS(id, "preview")),
+      mod_download_ui(NS(id, "download"))
+      )
+    )
+  )
+  }
+    
+# Module Server
+    
+#' @rdname mod_ssp_power_traditional
+#' @export
+#' @keywords internal
+    
+mod_ssp_power_traditional_server <- function(id, activate_menu, method_menu, activate_question, method_question){
+  
+  moduleServer(id, function(input, output, session) {
+    waitress <- waiter::Waitress$new("#traditional-preview-show_preview", theme = "overlay", infinite = TRUE, hide_on_render = TRUE)
+    
+    # Calculate results
+    traditional_result <- eventReactive(input$calculate, {
+      waitress$start()
+      ssp_power_traditional(n_min = input$n_min,
+                        n_max = input$n_max,
+                        delta = input$delta,
+                        opt = input$opt)
+    })
+    
+    # Send params to RMD
+    params <- eventReactive(input$calculate, {
+      list(
+        n1 = traditional_result()$n1,
+        output_power = traditional_result()$npower,
+        input_power = input$opt,
+        delta = input$delta,
+        n_min = input$n_min,
+        n_max = input$n_max
+      )
+    })
+
+    # Render preview
+    mod_preview_server("preview", activate = reactive(input$calculate), input_file = "traditional_output.Rmd", params = params)
+    
[…]

### 2020-06-28 R/mod_ssp_rope.R

@@ -1,0 +1,76 @@
+# Module UI
+  
+#' @title   mod_ssp_rope_ui and mod_ssp_rope_server
+#' @description  A shiny Module.
+#'
+#' @param id shiny id
+#' @param input internal
+#' @param output internal
+#' @param session internal
+#'
+#' @rdname mod_ssp_rope
+#'
+#' @keywords internal
+#' @export 
+#' @importFrom shiny NS tagList 
+mod_ssp_rope_ui <- function(id){
+  tagList(
+    sidebarLayout(
+    sidebarPanel(
+      h1("ROPE"),
+      sliderInput(NS(id, "opt"), "Power", min = 0, max = 1, value = 0.8, step = 0.1),
+      sliderInput(NS(id, "band"), "Band", min = 0, max = 1, value = 0.2, step = 0.1),
+      sliderInput(NS(id, "delta"), "Delta", min = 0, max = 2, value = 0, step = 0.1),
+      actionButton(NS(id, "calculate"), "Calculate sample size")),
+    mainPanel(
+      mod_preview_ui(NS(id, "preview")),
+      mod_download_ui(NS(id, "download"))
+    )
+    )
+  )
+}
+
+# Module Server
+    
+#' @rdname mod_ssp_rope
+#' @export
+#' @keywords internal
+    
+mod_ssp_rope_server <- function(id){
+  
+  moduleServer(id, function(input, output, session) {
+    # Setup loadingbar
+    waitress <- waiter::Waitress$new("#rope-preview-show_preview", theme = "overlay", infinite = TRUE, hide_on_render = TRUE)
+    
+    # Calculate results
+    rope_result <- eventReactive(input$calculate, {
+      waitress$start()
+      ssp_rope(opt = input$opt, band = input$band, delta = input$delta)
+    })
+    
+    # Send params to RMD
+    params <- eventReactive(input$calculate, {
+      list(
+        n1 = rope_result()$n1,
+        output_power = rope_result()$npower,
+        input_power = input$opt,
+        band = input$band,
+        delta = input$delta
+      )
+    })
+    
+    # Render preview
+    mod_preview_server("preview", activate = reactive(input$calculate), input_file = "rope_output.Rmd", params = params)
+    
+    # Download the output
+    mod_download_server("download", activate = reactive(input$calculate), input_file = "rope_output.Rmd", params = params)
+
+  })
+}
+    
+## To be copied in the UI
[…]

### 2020-06-28 R/mod_ssp_tost.R

@@ -1,0 +1,79 @@
+# Module UI
+  
+#' @title   mod_ssp_tost_ui and mod_ssp_tost_server
+#' @description  A shiny Module.
+#'
+#' @param id shiny id
+#' @param input internal
+#' @param output internal
+#' @param session internal
+#'
+#' @rdname mod_ssp_tost
+#'
+#' @keywords internal
+#' @export 
+#' @importFrom shiny NS tagList 
+mod_ssp_tost_ui <- function(id){
+  tagList(
+    sidebarLayout(
+    sidebarPanel(
+      h1("TOST"),
+      sliderInput(NS(id, "opt"), "Power", min = 0, max = 1, value = 0.8, step = 0.1),
+      sliderInput(NS(id, "band"), "Band", min = 0, max = 1, value = 0.2, step = 0.1),
+      sliderInput(NS(id, "delta"), "Delta", min = 0, max = 2, value = 0, step = 0.1),
+      actionButton(NS(id, "calculate"), "Calculate sample size")
+      ),
+    mainPanel(
+      mod_preview_ui(NS(id, "preview")),
+      mod_download_ui(NS(id, "download"))
+      )
+    )
+  )
+}
+    
+# Module Server
+    
+#' @rdname mod_ssp_tost
+#' @export
+#' @keywords internal
+    
+mod_ssp_tost_server <- function(id){
+
+  moduleServer(id, function(input, output, session) {
+    # Setup loadingbar
+    waitress <- waiter::Waitress$new("#tost-preview-show_preview", theme = "overlay", infinite = TRUE, hide_on_render = TRUE)
+    
+    # Calculate results
+    tost_result <- eventReactive(input$calculate, {
+      waitress$start()
+      ssp_tost(opt = input$opt,
+                 band = input$band,
+                 delta = input$delta)
+      })
+    
+    # Send params to RMD
+    params <- eventReactive(input$calculate, {
+      list(
+        n1 = tost_result()$n1,
+        n2 = tost_result()$n2,
+        output_power = tost_result()$npower,
+        input_power = input$opt,
+        band = input$band,
+        delta = input$delta
+        )
+      })
+    
+    # Render preview
+    mod_preview_server("preview", activate = reactive(input$calculate), input_file = "tost_output.Rmd", params = params)
+    
+    # Download the output
+    mod_download_server("download", activate = reactive(input$calculate), input_file = "tost_output.Rmd", params = params)
+    })
[…]

### 2020-06-28 R/mod_tost.R

@@ -1,97 +1,0 @@
-# Module UI
-  
-#' @title   mod_tost_ui and mod_tost_server
-#' @description  A shiny Module.
-#'
-#' @param id shiny id
-#' @param input internal
-#' @param output internal
-#' @param session internal
-#'
-#' @rdname mod_tost
-#'
-#' @keywords internal
-#' @export 
-#' @importFrom shiny NS tagList 
-mod_tost_ui <- function(id){
-  tagList(
-    sidebarPanel(
-      h1("TOST"),
-      sliderInput(NS(id, "opt"), "Power", min = 0, max = 1, value = 0.8, step = 0.1),
-      sliderInput(NS(id, "band"), "Band", min = 0, max = 1, value = 0.2, step = 0.1),
-      sliderInput(NS(id, "delta"), "Delta", min = 0, max = 1, value = 0, step = 0.1),
-      actionButton(NS(id, "calculate"), "Calculate sample size!")
-      ),
-    mainPanel(
-      htmlOutput(NS(id, "tost_output")),
-      downloadButton(NS(id, "report"))
-      )
-  )
-}
-    
-# Module Server
-    
-#' @rdname mod_tost
-#' @export
-#' @keywords internal
-    
-mod_tost_server <- function(id){
-
-  moduleServer(id, function(input, output, session) {
-    
-    waitress <- waiter::Waitress$new("#tost_ui_1-tost_output", theme = "overlay", infinite = TRUE, hide_on_render = TRUE)
-    
-    tost_result <- eventReactive(input$calculate, {
-      waitress$start()
-      TOSTss(Opt = input$opt,
-             Band = input$band,
-             delta = input$delta)
-      })
-    
-    params <- eventReactive(input$calculate, {
-      list(
-        n1 = tost_result()$n1,
-        n2 = tost_result()$n2,
-        output_power = tost_result()$npower,
-        input_power = input$opt,
-        band = input$band,
-        delta = input$delta
-        )
-      })
-    
-    output$tost_output <- renderUI({
-      if (input$calculate) {
-        report_path <- file.path("inst/app/www/", "tost_output.Rmd")
-        file <- paste0("tost_preview", ".html")
-        callr::r(
-          render_report,
-          list(input = report_path, output = file, format = "html_document", params = params())
-          )
-        return(tags$iframe(style="height:400px; width:100%", src = "www/tost_preview.html", seamless = "seamless"))
-        } else {
[…]

### 2020-06-28 R/power_optim.R

@@ -1,0 +1,33 @@
+power_optim <- function(fun, range, delta, opt, ...) {
+  Ns = range
+  Res = c(fun(Ns[1], delta, ...), 
+          fun(Ns[2], delta, ...))
+  if (!(opt > min(Res) & opt < max(Res))) {
+    return(
+      list(
+        n1 = NA_real_,
+        npower = NA_real_
+        )
+      )
+    }
+  
+  while ((min(Ns[Res>opt])-max(Ns[Res<opt]))>1) {
+    ResL = Res[which(Ns==max(Ns[Res<opt]))]
+    ResH = Res[which(Ns==min(Ns[Res>opt]))]
+    NL = max(Ns[Res<opt])
+    NH = min(Ns[Res>opt])
+    NewN = round((NH-NL)*((opt-ResL)/(ResH-ResL))+NL)
+    while (NewN%in%Ns) {
+      NewN = ifelse(Res[length(Res)]>opt, NewN - 1, NewN + 1)
+      }
+    Ns = c(Ns, NewN)
+    Res = c(Res, fun(Ns[length(Ns)], delta, ...))
+    }
+  
+  return(
+    list(
+      n1 = min(Ns[Res>opt]),
+      npower = Res[which(Ns==min(Ns[Res>opt]))]
+      )
+    )
+}

### 2020-06-28 R/ssp_aipe.R

@@ -1,0 +1,11 @@
+ssp_aipe <- function(n_min, n_max, delta, opt) {
+  result <- power_optim(fun = traditional, range = c(n_min, n_max), delta = delta, opt = opt)
+  if(!is.null(result)) {
+    return(
+      list(n1 = result[1] %>% purrr::flatten_dbl(),
+           npower = result[2] %>% purrr::flatten_dbl())
+    )
+  } else {
+    stop("Your chosen confidence level cannot be achieved within this range of sample sizes!")
+  }
+}

### 2020-06-28 R/ssp_app.R

@@ -1,0 +1,9 @@
+ssp_app <- function(confidence, closeness) {
+  n1 <- ceiling((qnorm((sqrt(confidence) + 1) / 2) / closeness)^2)
+  
+  return(
+    list(
+      n1 = n1
+    )
+  )
+}

### 2020-06-28 R/ssp_eq_bf.R

@@ -1,0 +1,41 @@
+eq_bf <- function(n1, band, delta, thresh = 10, tol = 1e-4, granularity = 300) {
+  n2 = n1
+  t = seq(qt(.005, n1+n2-2, delta/sqrt(1/n1+1/n2)), 
+          qt(.995, n1+n2-2, delta/sqrt(1/n1+1/n2)), length = granularity)
+  bf = 1
+  i = 0
+  while (bf < thresh)
+  {
+    i = i + 1
+    upper = cdf_t(band, t[i], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
+                   prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)
+    lower = cdf_t(-band, t[i], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
+                   prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)
+    post_dens = upper - lower
+    prior_dens = pcauchy(band, scale = 1/sqrt(2)) - 
+      pcauchy(-band, scale = 1/sqrt(2))
+    bf = (post_dens / prior_dens) / ((1 - post_dens) / (1 - prior_dens))
+  }
+  bf = 1
+  j = length(t) + 1
+  while (bf < thresh)
+  {
+    j = j-1
+    upper = cdf_t(band, t[j], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
+                   prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)
+    lower = cdf_t(-band, t[j], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
+                   prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)
+    post_dens = upper - lower
+    prior_dens = pcauchy(band, scale = 1/sqrt(2)) - 
+      pcauchy(-band, scale = 1/sqrt(2))
+    bf = (post_dens / prior_dens) / ((1 - post_dens) / (1 - prior_dens))
+  }
+  npower = pt(t[j], n1+n2-2, delta/sqrt(1/n1+1/n2)) - pt(t[i], n1+n2-2, delta/sqrt(1/n1+1/n2))
+  
+  npower
+}
+
+ssp_eq_bf <- function(opt, band, delta, thresh) {
+  est <- ssp_tost(opt = opt, band = band, delta = delta)[1] %>% purrr::flatten_dbl()
[…]

### 2020-06-28 R/ssp_power_curve.R

@@ -1,0 +1,30 @@
+ssp_power_curve <- function(delta1, delta2, opt, animated = FALSE) {
+  noptim <- NULL
+  delta <- seq(delta1, delta2, 0.01)
+  for (i in 1:length(delta)) {
+    noptim[i] = power_optim(fun = traditional, range = c(5, 2000), delta[i], opt = opt)$n1
+    }
+  noptim
+  gg <- ggplot2::ggplot() +
+    ggplot2::aes(x = delta,
+                 y = noptim,
+                 text = paste("n:", noptim,
+                              "<br>delta: ", delta)) +
+    ggplot2::geom_point(shape = 21, colour = "black", fill = "white", size = 3, stroke = 0.5) +
+    ggplot2::scale_y_continuous(limits = c(0, 1500),
+                                labels = c("0", "500", "1000", "1500"),
+                                breaks = c(0, 500, 1000, 1500)) +
+    ggplot2::scale_x_continuous(limits = c(0, delta2),
+                                labels = as.character(delta),
+                                breaks = delta) +
+    ggplot2::labs(x = "Delta",
+                  y = "Planned sample size") +
+    ggplot2::theme_classic()
+  if (animated) {
+    gg %>% 
+    plotly::ggplotly(tooltip = "text") %>%
+    plotly::config(displayModeBar = F)
+  } else {
+    gg
+  }
+}

### 2020-06-28 R/ssp_power_traditional.R

@@ -1,0 +1,7 @@
+traditional <- function(n1, delta, alpha = .05) {
+  1 - pt(qt((1-alpha), n1*2-2), n1*2-2, delta/sqrt(2/n1))
+}
+
+ssp_power_traditional <- function(n_min, n_max, delta, opt) {
+  power_optim(fun = traditional, range = c(n_min, n_max), delta = delta, opt = opt)
+}

### 2020-06-28 R/ssp_tost.R

@@ -1,0 +1,32 @@
+ssp_tost <- function(opt, band, delta, sigma = 1, nr = 1, alpha = .05) {
+  n1 = 4
+  sigsq = sigma^2
+  numint = 1000
+  lc = numint + 1
+  cl = 1e-10
+  coevecc = c(1, rep(c(4, 2), numint / 2 - 1), 4, 1)
+  npower = 0
+  
+  while (npower < opt & n1 < 10001) {
+    n1 = n1 + 1
+    n2 = nr * n1
+    df = n1 + n2 - 2
+    tcrit = qt(1 - alpha, df)
+    nfac = 1/n1 + 1/n2
+    var = sigsq * nfac
+    std = sqrt(var)
+    cu = (df * band^2) / (var * tcrit^2)
+    int = cu - cl
+    intl = int / numint
+    cvec = cl + intl * (0:numint)
+    wcpdf = (intl / 3) * coevecc * dchisq(cvec, df)
+    st = sqrt(cvec / df) * tcrit
+    npower = sum(wcpdf * (pnorm((band - delta) /std - st) - pnorm((-band - delta) / std + st)))
+  }
+  
+  return(
+    list(n1 = round(n1, 4),
+         n2 = round(n2, 4),
+         npower = round(npower, 4))
+    )
+}

### 2020-06-28 dev/02_dev.R

@@ -6,34 +6,40 @@
 # 2. All along your project
 
 ## 2.1 Add modules
-### TODO: Change names everywhere
-golem::add_module( name = "question" )
-golem::add_module( name = "article" )
-golem::add_module( name = "flowchart" )
-golem::add_module( name = "about" )
-golem::add_module( name = "tost" )
-golem::add_module( name = "menu" )
-golem::add_module( name = "traditional_power" )
-golem::add_module( name = "power_curve" )
-golem::add_module( name = "apriori_precision" )
-golem::add_module( name = "rope" )
-golem::add_module( name = "interval_equiv_bf" )
-golem::add_module( name = "bf_threshold" )
+
+golem::add_module(name = "question")
+golem::add_module(name = "article")
+golem::add_module(name = "flowchart")
+golem::add_module(name = "about")
+golem::add_module(name = "tost")
+golem::add_module(name = "menu")
+golem::add_module(name = "traditional_power")
+golem::add_module(name = "power_curve")
+golem::add_module(name = "apriori_precision")
+golem::add_module(name = "rope")
+golem::add_module(name = "interval_equiv_bf")
+golem::add_module(name = "bf_threshold")
+golem::add_module(name = "AIPE")
+golem::add_module(name = "BFDA")
+golem::add_module(name = "download")
+golem::add_module(name = "preview")
 
 ## 2.2 Add dependencies
 
-usethis::use_package( "waiter" )
-usethis::use_package( "DiagrammeR" )
-usethis::use_package( "shinyglide" )
-usethis::use_package( "tibble" )
-usethis::use_package( "shinyWidgets" )
-usethis::use_package( "shinyjs" )
-usethis::use_package( "shinyBS" )
-usethis::use_package( "BayesFactor" )
+usethis::use_package("waiter")
+usethis::use_package("plotly")
+usethis::use_package("ggplot2")
+usethis::use_package("DiagrammeR")
+usethis::use_package("tibble")
+usethis::use_package("shinyWidgets")
+usethis::use_package("shinyjs")
+usethis::use_package("shinyBS")
+usethis::use_package("BayesFactor")
+usethis::use_package("klippy")
 
 ## 2.3 Add tests
 
-usethis::use_test( "app" )
+usethis::use_test("app")
 
 ## 2.4 Add a browser button
 
@@ -41,9 +47,9 @@ golem::browser_button()
 
 ## 2.5 Add external files
 
[…]

### 2020-06-28 informedTtest_functions.R

@@ -1,0 +1,419 @@
+
+integrand_t <- function(delta, t, n, nu, mu.delta, gamma, kappa) {
+  
+  suppressWarnings(
+    dt(x = t, df = nu, ncp = sqrt(n) * delta) *
+    1 / gamma * dt( (delta - mu.delta) / gamma, df = kappa)
+  )
+  
+}
+
+posterior_t <- function(delta,
+                        t,
+                        n1,
+                        n2 = NULL,
+                        independentSamples = FALSE,
+                        prior.location,
+                        prior.scale,
+                        prior.df,
+                        rel.tol = .Machine$double.eps^0.25) {
+  
+  neff <- ifelse(independentSamples, n1 * n2 / (n1 + n2), n1)
+  nu <- ifelse(independentSamples, n1 + n2 - 2, n1 - 1)
+  
+  mu.delta <- prior.location
+  gamma <- prior.scale
+  kappa <- prior.df
+  
+  numerator <- suppressWarnings(
+    dt(x = t, df = nu, ncp = sqrt(neff) * delta) *
+    1 / gamma * dt( (delta - mu.delta) / gamma, df = kappa)
+  )
+  
+  denominator <- integrate(integrand_t,
+                           lower = -Inf, upper = Inf,
+                           t = t, n = neff, nu = nu,
+                           mu.delta = mu.delta,
+                           gamma = gamma,
+                           kappa = kappa,
+                           rel.tol = rel.tol)$value
+  
+  out <- numerator / denominator
+  out[is.na(out)] <- 0
+  
+  return(out)
+  
+}
+
+cdf_t <- function(x,
+                  t,
+                  n1,
+                  n2 = NULL,
+                  independentSamples = FALSE,
+                  prior.location,
+                  prior.scale,
+                  prior.df,
+                  rel.tol = .Machine$double.eps^0.25) {
+  
+  out <- integrate(posterior_t,
+                   lower = -Inf, upper = x,
+                   t = t, n1 = n1, n2 = n2,
+                   independentSamples = independentSamples,
+                   prior.location = prior.location,
+                   prior.scale = prior.scale,
+                   prior.df = prior.df,
+                   rel.tol = rel.tol)$value
+  
+  # catch numerical errors
+  if (out > 1 & out < 1.001) {
+    out <- 1
+    warning(
+      "Numerical integration yields a CDF value slightly larger than 1. The CDF value has been replaced by 1.",
+      call. = FALSE
+    )
+  }
+  
[…]

### 2020-07-03 inst/app/www/about.Rmd

@@
 ---
-title: ""
+title: "&nbsp;"
 output: html_document
@@
 
-Important text here!
+The __SampleSizePlanner__ shiny application enables researchers to calculate the sample size with 8 different methods in app and provides code to use an additional method from R.
+The aim of the app is to make the various sample size calculating methods available for a wide range of researchers and propagate the transparent reporting of the choose method.
+Therefore, a formatted template text containing the used parameters and the output of the calculations is available for download.
+This text can be complemented with the reasoning behind each decision.
 
-### How to use the application?
-
-Important instructions here!
+### Available sample size planning methods
+- AIPE [REF](link)
+- APP [REF](link)
+- BF threshold [REF](link)
+- Power curve [REF](link)
+- Traditional power calculation [REF](link)
+- Interval equiv bf [REF](link)
+- Rope [REF](link)
+- TOST [REF](link)
+- BFDA [REF](link)
 
@@
 - Balazs Aczel
+
+### Questions and comments
+
+If you have any questions or comments please write to Marton Kovacs to the following email address: marton.balazs.kovacs@gmail.com
+
+Or submit an issue on the apps Github page: [https://github.com/marton-balazs-kovacs/SampleSizePlanner](https://github.com/marton-balazs-kovacs/SampleSizePlanner)

### 2020-07-03 inst/app/www/on_load_placeholder.Rmd

@@
 ---
-title: ''
+title: "&nbsp;"
 output: html_document
@@
 + After running the calculation you can download the results in a general template text in word format.
-+ After donwload, please complement the text where it is need with your reasoning behind each decision and include the text in your manuscript.
++ After download, please complement the text where it is need with your reasoning behind each decision and include the text in your manuscript.
++ Please wait for a calulation to finish before you attempt to download the output or open a new tab.

### 2020-07-06 R/app_ui.R
Made the code simpler because of preformance issues

@@ -6,7 +6,7 @@ app_ui <- function() {
     navbarPage(
       title = "Sample Size Planner",
         tabPanel("Flowchart",
-                 mod_flowchart_ui("flowchart_ui_1")),
+                 tags$img(src = "www/flowchart.png", width = "80%", heigth = "80%", align = "left")),
         navbarMenu("Methods",
                    tabPanel("TOST",
                             mod_ssp_tost_ui("tost")),

### 2020-07-06 R/mod_download.R

@@ -1,49 +1,0 @@
-#' download UI Function
-#'
-#' @description A shiny Module.
-#'
-#' @param id,input,output,session Internal parameters for {shiny}.
-#'
-#' @noRd 
-#'
-#' @importFrom shiny NS tagList 
-mod_download_ui <- function(id){
-  tagList(
-    div(class = "download-btn",
-        downloadButton(NS(id, "report")))
-  )
-}
-    
-#' download Server Function
-#'
-#' @noRd 
-mod_download_server <- function(id, activate, output_text, method){
-  moduleServer(id, function(input, output, session) {
-  # Add downloadbutton enable logic
-  observe({
-    if (activate()) {
-      shinyjs::enable("report")
-      shinyjs::runjs("$('.download-btn').removeAttr('title');")
-    } else{
-      shinyjs::disable("report")
-      shinyjs::runjs("$('.download-btn').attr('title', 'Please run the calculation first');")
-    }
-  })
-  
-  # Download the output
-  output$report <- downloadHandler(
-    filename = function() {
-      paste0(method, "_", Sys.Date(), ".txt")
-    },
-    content = function(file) {
-      writeLines(output_text(), file)
-    })
-  })
-}
-    
-## To be copied in the UI
-# mod_download_ui("download")
-    
-## To be copied in the server
-# mod_download_server("download")
- 

### 2021-01-07 inst/ssp_rope_performance.Rmd
small changes

---
title: "rope_performance_test"
author: "Marton Kovacs"
date: "11/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
source("informedTtest_functions.R")
```

This document contains all the test that are carried out in order to improve the performance of the `ssp_rope` function.

# Parameters used for testing

```{r}
n1 = 314
band = 0.2
delta = 0
power = 0.1
alpha = .05
tol = 1e-4
granularity = 300
upper = 0.5
lower = 0.5
```

# Util functions

```{r}
# source("informedTtest_functions.R)
cdf_t <- function(x,
                  t,
                  n1,
                  n2 = NULL,
                  independentSamples = FALSE,
                  prior.location,
                  prior.scale,
                  prior.df,
                  rel.tol = .Machine$double.eps^0.25) {

  out <- integrate(posterior_t,
                   lower = -Inf, upper = x,
                   t = t, n1 = n1, n2 = n2,
                   independentSamples = independentSamples,
                   prior.location = prior.location,
                   prior.scale = prior.scale,
                   prior.df = prior.df,
                   rel.tol = rel.tol)$value

  # catch numerical errors
  if (out > 1 & out < 1.001) {
    out <- 1
    warning(
      "Numerical integration yields a CDF value slightly larger than 1. The CDF value has been replaced by 1.",
      call. = FALSE
    )
  }

  return(out)

}

# Vectorized cdf_t
v_cdf_t <- Vectorize(cdf_t)

# detect2 function
# This function works like purrr::map2 but with detect
detect2_index <- function(.x, .y, .f, .dir = c("forward", "backward")) {
  .dir <- rlang::arg_match(.dir, c("forward", "backward"))

  for (i in index(.x, .dir)) {
    if (.f(.x[[i]], .y[[i]])) {
      return(i)
    }
  }

  0L
}

index <- function(x, dir) {
  idx <- seq_along(x)
  if (dir == "backward") {
    idx <- rev(idx)
  }
  idx
}

calculate_and_check <- function (band, t, n1, tol, alpha) {
    upper = cdf_t (band, t, n1 = n1, n2 = n1, independentSamples = T, prior.location = 0, 
                   prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)
    lower = cdf_t (-band, t, n1 = n1, n2 = n1, independentSamples = T, prior.location = 0, 
                   prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)

    lower<(alpha/2) & upper>(1-alpha/2)
}

check <- function (upper, lower, alpha = .05) {
  lower<(alpha/2) & upper>(1-alpha/2)
}
```

# The original function

```{r}
rope <- function(n1, delta, band, alpha = .05, tol = 1e-4, granularity = 300) {
  n2 = n1
  t = seq(qt(.005, n1+n2-2, delta/sqrt(1/n1+1/n2)), 
          qt(.995, n1+n2-2, delta/sqrt(1/n1+1/n2)), length = granularity)
  upper = lower = 0.5
  i = 0
  while (!(lower<(alpha/2) & upper>(1-alpha/2))) {
    i = i+1
    upper = cdf_t(band, t[i], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
                  prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)
    lower = cdf_t(-band, t[i], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
                  prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)
  }
  upper = lower = 0.5
  j = length(t) + 1
  while (!(lower<(alpha/2) & upper>(1-alpha/2))) {
    j = j-1
    upper = cdf_t(band, t[j], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
                  prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)
    lower = cdf_t(-band, t[j], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
                  prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)
  }

  npower = pt(t[j], n1+n2-2, delta/sqrt(1/n1+1/n2)) - pt(t[i], n1+n2-2, delta/sqrt(1/n1+1/n2))

  cat ("\n n1 is ", n1, ", power is ", npower, sep = "")

}

rope(n1 = 314, band = 0.3, delta = 0)
```

# Profiling

```{r eval=FALSE}
profvis::profvis(rope(n1 = 314, delta = 0, band = 0.2))
```

# Fully vectorized version

```{r}
v_rope <- function (n1, delta, band, alpha = .05, tol = 1e-4, granularity = 300) {
    n2 = n1
    t = seq (qt (.005, n1+n2-2, delta/sqrt(1/n1+1/n2)), 
             qt (.995, n1+n2-2, delta/sqrt(1/n1+1/n2)), length = granularity)

    bound <- 
      list(
        Upper = v_cdf_t(band, t, n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
                        prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol),
        Lower = v_cdf_t(-band, t, n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
                        prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)
        )

    i <- detect2_index(bound[["Upper"]], bound[["Lower"]], check, .dir = "forward")
    j <- detect2_index(bound[["Upper"]], bound[["Lower"]], check, .dir = "backward")

    npower = pt(t[j], n1+n2-2, delta/sqrt(1/n1+1/n2)) - pt(t[i], n1+n2-2, delta/sqrt(1/n1+1/n2))

    cat ("\n n1 is ", n1, ", power is ", npower, sep = "")
}

v_rope(n1 = 314, band = 0.3, delta = 0)
```

# Purrr version

```{r}
purrr_rope <- function (n1, band, delta, alpha = .05, tol = 1e-4, granularity = 300) {
    n2 = n1
    t = seq (qt (.005, n1+n2-2, delta/sqrt(1/n1+1/n2)), 
             qt (.995, n1+n2-2, delta/sqrt(1/n1+1/n2)), length = granularity)

    i <- purrr::detect(t, ~ calculate_and_check(n1 = n1, band = band, t = .x, tol = tol, alpha = alpha), .dir = "forward")
    j <- purrr::detect(t, ~ calculate_and_check(n1 = n1, band = band, t = .x, tol = tol, alpha = alpha), .dir = "backward")

    npower = pt(j, n1+n2-2, delta/sqrt(1/n1+1/n2)) - pt(i, n1+n2-2, delta/sqrt(1/n1+1/n2))

    cat ("\n n1 is ", n1, ", power is ", npower, sep = "")
}

purrr_rope(n1 = 314, band = 0.3, delta = 0)
```

# Rcpp version

```{r}
Rcpp::sourceCpp('rcpp_functions.cpp')
Rcpp::sourceCpp('rcpp_informed_ttest_functions.cpp')

integrate_exp4()

test <- my_qt(314, 0, 300)

original <- seq (qt (.005, 314+314-2, 0/sqrt(1/314+1/314)), 
             qt (.995, 314+314-2, 0/sqrt(1/314+1/314)), length = 300)

microbenchmark::microbenchmark(
  my_qt(314, 0, 300),
  seq (qt (.005, 314+314-2, 0/sqrt(1/314+1/314)), 
             qt (.995, 314+314-2, 0/sqrt(1/314+1/314)), length = 300),
  times = 10000
)

rcpp_rope <- function(n1, delta, band, alpha = .05, tol = 1e-4, granularity = 300) {
  n2 = n1
  t = my_qt(n1, delta, granularity)
  t = as.vector(t)
  upper = lower = 0.5
  i = 0
  while (!(lower<(alpha/2) & upper>(1-alpha/2))) {
    i = i+1
    upper = cdf_t(band, t[i], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
                  prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)
    lower = cdf_t(-band, t[i], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
                  prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)
  }
  upper = lower = 0.5
  j = length(t) + 1
  while (!(lower<(alpha/2) & upper>(1-alpha/2))) {
    j = j-1
    upper = cdf_t(band, t[j], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
                  prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)
    lower = cdf_t(-band, t[j], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
                  prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)
  }

  npower = pt(t[j], n1+n2-2, delta/sqrt(1/n1+1/n2)) - pt(t[i], n1+n2-2, delta/sqrt(1/n1+1/n2))

  cat ("\n n1 is ", n1, ", power is ", npower, sep = "")

}
```

# Compare performance

```{r}
microbenchmark::microbenchmark(
    rope(n1 = 314, band = 0.3, delta = 0),
    # v_rope(314, 0.3, 0),
    # purrr_rope(n1 = 314, band = 0.3, delta = 0),
    rcpp_rope(n1 = 314, band = 0.3, delta = 0),
    times = 10
)  
```

### 2021-01-07 R/assertions.R

@@ -1,0 +1,27 @@
+is_not_null <- function(x) {
+  !is.null(x)
+}
+
+assertthat::on_failure(is_not_null) <- function(call, env) {
+  paste(deparse(substitute(x, env)), "argument must be defined")
+}
+
+is_positive_number <- function(x) {
+  assertthat::assert_that(is_not_null(x))
+  assertthat::assert_that(is.numeric(x))
+  x > 0
+}
+
+assertthat::on_failure(is_positive_number) <- function(call, env) {
+  paste(deparse(call$x), "must be a postive number")
+}
+
+# if (any(c(opt, band,delta) <= 0)) {
+#   rlang::abort("opt, band and delta must be positive numbers")
+# }
+
+# if (any(is.null(opt),
+#         is.null(band),
+#         is.null(delta))) {
+#   rlang::abort("opt, band, and delta must be defined")
+# }

### 2021-01-07 dev/01_start.R

@@ -37,7 +37,8 @@ usethis::use_git()
 ## 1.3 - Add a data-raw folder
 ## 
 ## If you have data in your package
-usethis::use_data_raw( name = "question_data", open = FALSE )
+# usethis::use_data_raw( name = "question_data", open = FALSE )
+usethis::use_data_raw(name = "rope_data", open = FALSE)
 
 ## 1.4 - Init Tests
 ## 

### 2021-01-07 dev/02_dev.R

@@ -35,7 +35,7 @@ usethis::use_package("shinyWidgets")
 usethis::use_package("shinyjs")
 usethis::use_package("shinyBS")
 usethis::use_package("BayesFactor")
-usethis::use_package("klippy")
+usethis::use_package("assertthat")
 
 ## 2.3 Add tests
 

### 2021-03-19 inst/app/www/about.Rmd
updated UI and function documentation

@@
 
-The __SampleSizePlanner__ shiny application enables researchers to calculate the sample size with 8 different methods in app and provides code to use an additional method from R.
+The __SampleSizePlanner__ shiny application enables researchers to calculate the sample size with 9 different methods in app and provides code to use an additional method from R.
 The aim of the app is to make the various sample size calculating methods available for a wide range of researchers and propagate the transparent reporting of the choose method.

### 2021-03-19 inst/app/www/introduction.Rmd
updated UI and function documentation

---
title: ''
output: html_document
---
<center>
### How to use the app

This application provides 9 different methods to determine the sample size for a two group design. Moreover, the app provides help to create an accompanying report text for your sample size determination.

The figure below shows the 9 sample size estimation methods in the app and their relations. If you would like to know more about each method you can read our [tutorials manuscript](http://corndog.io/).

To get started with the sample size estimation choose a method under the __Methods__ tab.

<img src="www/flowchart.jpg" width="80%"/>

Please visit the [About](http://corndog.io/) tab if you would like to know more about the project or if you have any suggestions or issues with the app.

</center>

### 2021-03-19 R/app_server.R

@@ -1,15 +1,12 @@
 #' @import shiny
 app_server <- function(input, output,session) {
-
-  # Create flowchart module
-  mod_flowchart_server("flowchart_ui_1")
   
   # Methods modules
   mod_ssp_tost_server("tost")
   mod_ssp_power_traditional_server("traditional")
   mod_ssp_power_curve_server("curve")
   mod_ssp_rope_server("rope")
-  mod_ssp_bf_thresh_server("bf_thresh")
+  mod_ssp_bf_predetermined_server("bf_predetermined")
   mod_ssp_app_server("app")
   mod_ssp_eq_bf_server("eq_bf")
   mod_ssp_aipe_server("aipe")

### 2021-03-19 R/jusfitication.R

@@ -1,0 +1,131 @@
+justification <- function(method, output_parameters) {
+  
+  if (method == "aipe") {
+    text <- 
+      glue::glue(
+      "In order to estimate the sample size, we used the accuracy in parameter estimation \\
+      (AIPE; Kelley and Rausch, 2006) method. We aimed for a {confidence_level}% confidence level, \\
+      because {confidence_level_justification}. The desired width was {width} because \\
+      {width_justification}. We expected an underlying population effect size of {delta}, {delta_justification}.
+      Based on these parameters, a minimal sample size of {n1} per group was estimated for our design.",
+      width = output_parameters$width,
+      delta = output_parameters$delta,
+      n1 = output_parameters$n1,
+      confidence_level = output_parameters$confidence_level,
+      confidence_level_justification = output_parameters$confidence_level_justification,
+      width_justification = output_parameters$width_justification,
+      delta_justification = output_parameters$delta_justification
+      )
+  } else if (method == "app") {
+    text <- 
+      glue::glue(
+      "In order to estimate the sample size, we used the a-priori precision \\
+      (APP; Trafimow and MacDonald, 2017) method. A-priori to data collection, \\
+      we wanted to be {confidence}% confident that both sample means lie within {closeness} SD of \\
+      the true population means. Based on these parameters, the resulting minimum \\
+      sample size was {n1} per group for our design.",
+      confidence = output_parameters$confidence,
+      closeness = output_parameters$closeness,
+      n1 = output_parameters$n1
+    )
+  } else if (method == "tost") {
+    text <-
+      glue::glue(
+        "In order to calculate an appropriate sample size for testing whether the \\
+        two groups are practically equivalent, we used the Two One-Sided Tests of \\
+        Equivalence (TOST; Schuirmann, 1981) method. We set the aimed TPR to be {tpr}, \\
+        because {tpr_justification}. We consider all effect sizes below {eq_band} \\
+        equivalent to zero, because {eq_band_justification}. The expected delta was {delta} \\
+        because {delta_justification}. Based on these parameters, a sample size of {n1} \\
+        per group was estimated in order to reach a TPR of {round(npower, 1)} with our design.",
+        eq_band = output_parameters$eq_band,
+        delta = output_parameters$delta,
+        n1 = output_parameters$n1,
+        npower = output_parameters$npower,
+        tpr = output_parameters$tpr,
[…]

### 2021-03-19 R/mod_preview.R

@@ -9,50 +9,65 @@
 #' @importFrom shiny NS tagList 
 mod_preview_ui <- function(id){
   tagList(
-    mainPanel(
-      h3("Instructions"),
-      tags$ul(tags$li("Please set the parameters and run the sample size planning calculation with the \"Calculate sample size\" button."),
-              tags$li("After running the calculation you can download the results in a general template text in word format."),
-              tags$li("After download, please complement the text where it is need with your reasoning behind each decision and include the text in your manuscript."),
-              tags$li("Please wait for a calulation to finish before you attempt to download the output or open a new tab.")),
-    div(style = "height:120px;", textOutput(NS(id, "show_preview"))),
-    div(class = "download-btn",
-        downloadButton(NS(id, "report"))))
-  )
+    # Show output
+    wellPanel(
+      # style = "height:120px;",
+      textOutput(NS(id, "show_preview"))
+      ),
+    # Download button
+    div(
+      class = "download-justification-btn",
+      downloadButton(NS(id, "report"), label = "Download"),
+      actionButton(NS(id, "clip"), label = "Copy")
+      )
+    )
 }
     
 #' preview Server Function
 #'
 #' @noRd 
-mod_preview_server <- function(id, activate, output_text, method){
+mod_preview_server <- function(id, activate, output_parameters, method){
   moduleServer(id, function(input, output, session) {
-  # Render preview
-  output$show_preview <- renderText({
-    req(activate())
-    output_text()
-    })
+    # Create justification text
+    justification_text <- reactive({
+      req(activate())
+      justification(
+        method = method,
+        output_parameters = isolate(output_parameters())
+        )
+      })
+    
+    # Show justification
+    output$show_preview <- renderText(justification_text())
+    
+    # Add downloadbutton enable logic
+    observe({
+      if (activate()) {
+        shinyjs::enable("report")
+        shinyjs::enable("clip")
+        shinyjs::runjs("$('.download-justification-btn').removeAttr('title');")
+        } else{
+          shinyjs::disable("report")
+          shinyjs::disable("clip")
+          shinyjs::runjs("$('.download-justification-btn').attr('title', 'Please run the calculation first');")
+          }
+      })
   
-  # Add downloadbutton enable logic
-  observe({
-    if (activate()) {
-      shinyjs::enable("report")
-      shinyjs::runjs("$('.download-btn').removeAttr('title');")
-    } else{
[…]

### 2021-03-19 R/mod_ssp_aipe.R

@@ -13,22 +13,103 @@
 #' @keywords internal
 #' @export 
 #' @importFrom shiny NS tagList 
-mod_ssp_aipe_ui <- function(id){
-  
+mod_ssp_aipe_ui <- function(id) {
   tagList(
+    # Method
+    h1("Accuracy In Parameter Estimation"),
     sidebarLayout(
-    # Settings
-    sidebarPanel(
-      h1("Accuracy In Parameter Estimation"),
-      sliderInput(NS(id, "opt"), "Confidence", min = 0, max = 1, value = 0.8, step = 0.1),
-      sliderInput(NS(id, "delta"), "Delta", min = 0, max = 2, value = 0.5, step = 0.1),
-      numericInput(NS(id, "n_min"), "Minimum sample size", value = 20, min = 5),
-      numericInput(NS(id, "n_max"), "Maximum sample size", value = 200, max = 2000),
-      actionButton(NS(id, "calculate"), "Calculate sample size")),
-    # Output
-      mod_preview_ui(NS(id, "preview"))
+      sidebarPanel(
+        # Panel title
+        h3("Determine your sample size", class = "subtitle"),
+        # Method description
+        p("A sample size estimation method used for parameter estimation. The approach aims to find the required sample size,
+          such that the confidence interval has a certain expected width. "),
+        # Calculation settings
+        ## Confidence level input
+        sliderInput(
+          NS(id, "confidence_level"),
+          label = "Confidence Level",
+          min = 0,
+          max = 1,
+          value = 0.8,
+          step = 0.1),
+        ## Delta input
+        sliderInput(
+          NS(id, "delta"),
+          label = "Delta",
+          min = 0,
+          max = 2,
+          value = 0.5,
+          step = 0.1),
+        ## Width input
+        sliderInput(
+          NS(id, "width"),
+          label = "Width",
+          min = 0,
+          max = 1,
+          value = 0.2,
+          step = 0.1),
+        # Run calculation
+        actionButton(NS(id, "calculate"), "Calculate sample size", class = "calculate-btn"),
+        # Show the results of the calculation
+        wellPanel(
+          class = "panel-calculate",
+          htmlOutput(NS(id, "calculate_output")))),
+      # Output
+      mainPanel(
+        wellPanel(
+          tabsetPanel(
+            type = "tabs",
+            tabPanel(
+              "Justification",
+              # Panel title
+              h3("Justify your sample size"),
+              # Justification for confidence level
+              selectizeInput(
[…]

### 2021-03-19 R/mod_ssp_app.R

@@ -13,16 +13,60 @@
 #' @keywords internal
 #' @export 
 #' @importFrom shiny NS tagList 
-mod_ssp_app_ui <- function(id){
-  
+mod_ssp_app_ui <- function(id) {
   tagList(
+    # Method
+    h1("APP (a-priori precision)"),
     sidebarLayout(
-    sidebarPanel(
-      h1("APP (a-priori precision)"),
-      sliderInput(NS(id, "closeness"), "Closeness", min = 0, max = 1, value = 0.2, step = 0.1),
-      sliderInput(NS(id, "confidence"), "Confidence", min = 0, max = 1, value = 0.95, step = 0.1),
-      actionButton(NS(id, "calculate"), "Calculate sample size")),
-    mod_preview_ui(NS(id, "preview"))
+      sidebarPanel(
+        # Panel title
+        h3("Determine your sample size", class = "subtitle"),
+        # Method description
+        p("valami"),
+        # Calculation settings
+        ## Closeness input
+        sliderInput(
+          NS(id, "closeness"),
+          "Closeness",
+          min = 0,
+          max = 1,
+          value = 0.2,
+          step = 0.1),
+        ## Confidence input
+        sliderInput(
+          NS(id, "confidence"),
+          "Confidence",
+          min = 0, 
+          max = 1,
+          value = 0.95, 
+          step = 0.1),
+        # Run calculation
+        actionButton(NS(id, "calculate"), "Calculate sample size", class = "calculate-btn"),
+        # Show the results of the calculation
+        wellPanel(
+          class = "panel-calculate",
+          htmlOutput(NS(id, "calculate_output")))),
+      # Output
+      mainPanel(
+        wellPanel(
+          tabsetPanel(
+            type = "tabs",
+            tabPanel(
+              "Justification",
+              # Panel title
+              h3("Justify your sample size"),
+              # Create justification text
+              actionButton(NS(id, "justification"), "Create justification report", class = "calculate-btn"),
+              # Show justification text
+              mod_preview_ui(NS(id, "preview"))),
+            tabPanel(
+              "Code",
+              # Panel title
+              h3("Function call to use in R"),
+              mod_code_ui(NS(id, "code")))
+          )
+        )
+      )
     )
   )
 }
@@ -32,9 +76,7 @@ mod_ssp_app_ui <- function(id){
 #' @rdname mod_ssp_app
 #' @export
 #' @keywords internal
-    
-mod_ssp_app_server <- function(id){
-  
+mod_ssp_app_server <- function(id) {
[…]

### 2021-03-19 R/mod_ssp_bf_thresh.R

@@ -1,57 +1,0 @@
-# Module UI
-  
-#' @title   mod_ssp_bf_thresh_ui and mod_ssp_bf_thresh_server
-#' @description  A shiny Module.
-#'
-#' @param id shiny id
-#' @param input internal
-#' @param output internal
-#' @param session internal
-#'
-#' @rdname mod_ssp_bf_thresh
-#'
-#' @keywords internal
-#' @export 
-#' @importFrom shiny NS tagList 
-mod_ssp_bf_thresh_ui <- function(id){
-  tagList(
-    sidebarLayout(
-    sidebarPanel(
-      h1("Decide BF threshold"),
-      sliderInput(NS(id, "delta"), "Delta", min = 0, max = 2, value = 0.5, step = 0.1),
-      selectInput(NS(id, "thresh"), "Threshold", choices = c(10, 6, 3), selected = 10),
-      actionButton(NS(id, "calculate"), "Calculate sample size")),
-    mod_preview_ui(NS(id, "preview"))
-    )
-  )
-}
-    
-# Module Server
-    
-#' @rdname mod_ssp_bf_thresh
-#' @export
-#' @keywords internal
-    
-mod_ssp_bf_thresh_server <- function(id){
-  
-  moduleServer(id, function(input, output, session) {
-    # Setup loadingbar
-    # waitress <- waiter::Waitress$new("#bf_thresh-preview-show_preview", theme = "overlay", infinite = TRUE, hide_on_render = TRUE)
-    
-    # Calculate results
-    bf_thresh_result <- eventReactive(input$calculate, {
-      # Waitress$start()
-      ssp_bf_thresh(opt = 0.8, band = 0.2, delta = input$delta, thresh = as.integer(input$thresh), report_text = TRUE)
-      })
-    
-    # Render preview
-    mod_preview_server("preview", activate = reactive(input$calculate), output_text = bf_thresh_result, method = "bf_thresh")
-  })
-}
-    
-## To be copied in the UI
-# mod_ssp_bf_thresh_ui("bf_thresh")
-    
-## To be copied in the server
-# mod_ssp_bf_thresh_server("bf_thresh")
- 

### 2021-03-19 R/mod_ssp_bfda.R

@@ -13,19 +13,85 @@
 #' @keywords internal
 #' @export 
 #' @importFrom shiny NS tagList 
-mod_ssp_bfda_ui <- function(id){
-
+mod_ssp_bfda_ui <- function(id) {
   tagList(
+    # Method
+    h1("BFDA (Bayes Factor Design Analysis)"),
     sidebarLayout(
-    sidebarPanel(
-      h1("BFDA"),
-      tags$p("To compute this sample size estimation method takes too long, therefore we do not provide an online tool for it within the app. However, you can set your parameters that will update the code accordingly and copy paste the code into an R session with the \"clip\" button. After this, you can run the computation on your own local computer."),
-      sliderInput(NS(id, "delta"), "Delta", min = 0, max = 2, value = 0, step = 0.1),
-      selectInput(NS(id, "tresh"), "Threshold", choices = c(3, 6, 10), selected = 10),
-      actionButton(NS(id, "clip"), "Clip")),
-    mainPanel(
-      verbatimTextOutput(NS(id, "bfda_output"))
-    )
+      sidebarPanel(
+        # Panel title
+        h3("Determine your sample size", class = "subtitle"),
+        # Method description
+        p("valami"),
+        # Calculation settings
+        ## Delta input
+        sliderInput(
+          NS(id, "delta"),
+          "Delta",
+          min = 0, 
+          max = 2, 
+          value = 0,
+          step = 0.1),
+        sliderInput(
+          NS(id, "tpr"),
+          "TPR",
+          min = 0,
+          max = 1, 
+          value = 0.8, 
+          step = 0.1),
+        selectInput(
+          NS(id, "thresh"), 
+          "Threshold", 
+          choices = 3,
+          selected = 3),
+        # Run calculation
+        actionButton(NS(id, "calculate"), "Calculate sample size", class = "calculate-btn"),
+        # Show the results of the calculation
+        wellPanel(
+          class = "panel-calculate",
+          htmlOutput(NS(id, "calculate_output")))),
+      # Output
+      mainPanel(
+        wellPanel(
+          tabsetPanel(
+            type = "tabs",
+            tabPanel(
+              "Justification",
+              # Panel title
+              h3("Justify your sample size"),
+              # Justification for TPR
+              selectizeInput(NS(id, "tpr_justification"),
+                             label = "TPR",
+                             choices = c(
+                               "it is the common standard in the field",
+                               "it is the journal publishing requirement",
[…]

### 2021-03-19 R/mod_ssp_power_curve.R

@@ -13,19 +13,80 @@
 #' @keywords internal
 #' @export 
 #' @importFrom shiny NS tagList 
-mod_ssp_power_curve_ui <- function(id){
+mod_ssp_power_curve_ui <- function(id) {
   tagList(
+    # Method
+    h1("Power curve"),
     sidebarLayout(
-    sidebarPanel(
-      h1("Power curve"),
-      sliderInput(NS(id, "opt"), "Power", min = 0, max = 1, value = 0.8, step = 0.1),
-      sliderInput(NS(id, "delta"), "Delta", min = 0, max = 2, value = c(0.1, 0.9), step = 0.1),
-      actionButton(NS(id, "calculate"), "Calculate sample size")),
-    mainPanel(
-      plotly:: plotlyOutput(NS(id, "preview")),
-      div(class = "download-btn",
-          downloadButton(NS(id, "report")))
-    )
+      sidebarPanel(
+        # Panel title
+        h3("Determine your sample size", class = "subtitle"),
+        # Method description
+        p("valami"),
+        # Calculation settings
+        ## TPR input
+        sliderInput(
+          NS(id, "tpr"),
+          "TPR",
+          min = 0,
+          max = 1,
+          value = 0.8,
+          step = 0.1),
+        sliderInput(
+          NS(id, "delta"),
+          "Delta",
+          min = 0,
+          max = 2,
+          value = c(0.1, 0.9),
+          step = 0.1),
+        numericInput(
+          NS(id, "max_n"),
+          "Maximum N",
+          min = 10,
+          max = 20000,
+          value = 5000,
+          step = 1),
+        # Run calculation
+        actionButton(NS(id, "calculate"), "Calculate sample size", class = "calculate-btn"),
+        # Show the results of the calculation
+        wellPanel(
+          class = "panel-calculate",
+          htmlOutput(NS(id, "calculate_output")))),
+      # Output
+      mainPanel(
+        wellPanel(
+          tabsetPanel(
+            type = "tabs",
+            tabPanel(
+              "Justification",
+              # Panel title
+              h3("Justify your sample size"),
+              # Justification for TPR
+              selectizeInput(
+                NS(id, "tpr_justification"),
+                label = "Justify TPR",
+                choices = c(
+                  "it is the common standard in the field",
+                  "it is the journal publishing requirement",
+                  "other..."),
+                multiple = FALSE,
+                options = list(create = TRUE)),
+              # Create justification text
[…]

### 2021-03-19 R/mod_ssp_tost.R

@@ -13,19 +13,105 @@
 #' @keywords internal
 #' @export 
 #' @importFrom shiny NS tagList 
-mod_ssp_tost_ui <- function(id){
+mod_ssp_tost_ui <- function(id) {
   tagList(
+    # Method
+    h1("Two One‐Sided Tests (TOST)", class = "method-title"),
     sidebarLayout(
-    sidebarPanel(
-      h1("TOST"),
-      sliderInput(NS(id, "opt"), "Power", min = 0, max = 1, value = 0.8, step = 0.1),
-      sliderInput(NS(id, "band"), "Band", min = 0, max = 1, value = 0.2, step = 0.1),
-      sliderInput(NS(id, "delta"), "Delta", min = 0, max = 2, value = 0, step = 0.1),
-      actionButton(NS(id, "calculate"), "Calculate sample size")
-      ),
-    mod_preview_ui(NS(id, "preview"))
+      sidebarPanel(
+        # Panel title
+        h3("Determine your sample size", class = "subtitle"),
+        # Method description
+        p("TOST is a frequentist statistical testing approach aimed at establishing equivalence between two groups."),
+        # Calculation settings
+        ## TPR input
+        sliderInput(
+          NS(id, "tpr"),
+          list(
+            "TPR",
+            HTML('<i class="fas fa-info"; title="The desired long run probability of obtaining a significant result with TOST, given Delta."></i>')),
+          min = 0,
+          max = 1,
+          value = 0.8,
+          step = 0.1),
+        ## EqBand input
+        sliderInput(
+          NS(id, "eq_band"),
+          list(
+            "EqBand",
+            HTML('<i class="fas fa-info"; title="The chosen width of the region for practical equivalence, i.e. the SESOI."></i>')),
+          min = 0,
+          max = 1,
+          value = 0.2,
+          step = 0.1),
+        ## Delta input
+        sliderInput(
+          NS(id, "delta"),
+          list(
+            "Delta",
+            HTML('<i class="fas fa-info"; title="The expected population effect size. In most cases, this value will be zero."></i>')),
+          min = 0,
+          max = 2, 
+          value = 0, 
+          step = 0.1),
+        # Run calculation
+        actionButton(NS(id, "calculate"), "Calculate sample size", class = "calculate-btn"),
+        # Show the results of the calculation
+        wellPanel(
+          class = "panel-calculate",
+          htmlOutput(NS(id, "calculate_output")))),
+      # Output
+      mainPanel(
[…]

### 2021-03-19 R/power_optim.R

@@ -1,24 +1,19 @@
-power_optim <- function(fun, range, delta, opt, ...) {
+power_optim <- function(fun, range, delta, tpr, ...) {
   Ns = range
   Res = c(fun(Ns[1], delta, ...), 
           fun(Ns[2], delta, ...))
-  if (!(opt > min(Res) & opt < max(Res))) {
-    return(
-      list(
-        n1 = NA_real_,
-        npower = NA_real_
-        )
-      )
+  if (!(tpr > min(Res) & tpr < max(Res))) {
+    stop("Your chosen power level cannot be achieved for n < 15000!")
     }
   
-  while ((min(Ns[Res>opt])-max(Ns[Res<opt]))>1) {
-    ResL = Res[which(Ns==max(Ns[Res<opt]))]
-    ResH = Res[which(Ns==min(Ns[Res>opt]))]
-    NL = max(Ns[Res<opt])
-    NH = min(Ns[Res>opt])
-    NewN = round((NH-NL)*((opt-ResL)/(ResH-ResL))+NL)
+  while ((min(Ns[Res>tpr])-max(Ns[Res < tpr]))>1) {
+    ResL = Res[which(Ns == max(Ns[Res < tpr]))]
+    ResH = Res[which(Ns == min(Ns[Res > tpr]))]
+    NL = max(Ns[Res < tpr])
+    NH = min(Ns[Res > tpr])
+    NewN = round((NH - NL) * ((tpr - ResL) / (ResH - ResL)) + NL)
     while (NewN%in%Ns) {
-      NewN = ifelse(Res[length(Res)]>opt, NewN - 1, NewN + 1)
+      NewN = ifelse(Res[length(Res)] > tpr, NewN - 1, NewN + 1)
       }
     Ns = c(Ns, NewN)
     Res = c(Res, fun(Ns[length(Ns)], delta, ...))
@@ -26,8 +21,8 @@ power_optim <- function(fun, range, delta, opt, ...) {
   
   return(
     list(
-      n1 = min(Ns[Res>opt]),
-      npower = Res[which(Ns==min(Ns[Res>opt]))]
+      n1 = min(Ns[Res > tpr]),
+      npower = Res[which(Ns == min(Ns[Res > tpr]))]
       )
     )
 }

### 2021-03-19 R/show_code.R

@@ -1,0 +1,78 @@
+show_code <- function(method, code_parameters) {
+  # Setup code
+  code_text <- 
+    glue::glue(
+      "# Download package
+      install.packages(\"SampleSizePlanner\")
+      # Load package
+      library(SampleSizePlanner)
+      # Run calculation")
+  
+  # Interactive function call by method
+  if (method == "aipe") {
+    function_call <-
+      glue::glue(
+        "ssp_aipe(delta = {delta}, confidence_level = {confidence_level}, width = {width}, which_width = \"Full\", certainty = NULL)",
+        width = code_parameters$width,
+        delta = code_parameters$delta,
+        confidence_level = code_parameters$confidence_level)
+  } else if (method == "app") {
+    function_call <-
+      glue::glue(
+        "ssp_app(confidence = {confidence}, closeness = {closeness})",
+        confidence = code_parameters$confidence,
+        closeness = code_parameters$closeness)
+  } else if (method == "tost") {
+    function_call <-
+      glue::glue(
+        "ssp_tost(trp = {tpr}, eq_band = {eq_band}, delta = {delta}, sigma = 1, nr = 1, alpha = .05)",
+        eq_band = code_parameters$eq_band,
+        delta = code_parameters$delta,
+        tpr = code_parameters$tpr)
+  } else if (method == "eq_bf") {
+    function_call <-
+      glue::glue("ssp_eq_bf(trp = {tpr}, eq_band = {eq_band}, delta = {delta}, thresh = {thresh})",
+                 eq_band = code_parameters$eq_band,
+                 thresh = code_parameters$thresh, 
+                 delta = code_parameters$delta,
+                 tpr = code_parameters$tpr)
+  } else if (method == "traditional") {
+    function_call <-
+      glue::glue(
+        "ssp_power_traditional(max_n = {max_n}, delta = {delta}, tpr = {tpr}, alpha = 0.05)",
+        delta = code_parameters$delta,
+        tpr = code_parameters$tpr,
+        max_n = code_parameters$max_n)
+  } else if (method == "bf_predetermined") {
+    function_call <-
+      glue::glue(
+        "ssp_bf_predetermined(trp = {tpr}, eq_band = {eq_band}, delta = {delta}, thresh = {thresh}, max_n = {max_n})",
+        eq_band = code_parameters$eq_band,
+        thresh = code_parameters$thresh, 
+        delta = code_parameters$delta,
[…]

### 2021-03-19 R/ssp_aipe.R

@@ -1,8 +1,25 @@
-ssp_aipe <- function(n_min, n_max, delta, confidence, alpha = 0.05) {
-  result <- power_optim(fun = traditional, range = c(n_min, n_max), delta = delta, opt = confidence, alpha =alpha)
-  if(!all(is.na(result))) {
-    result
-    } else {
-      stop("Your chosen confidence level cannot be achieved within this range of sample sizes!")
-      }
+#' Determine sample size with Accuracy In Parameter Estimation (AIPE) method
+#' 
+#' A sample size estimation method used for parameter estimation.
+#' The approach aims to find the required sample size, such that
+#' the confidence interval has a certain expected width. The
+#' function is a simple wrapper function around the \code{\link[MBESS]{ss.aipe.smd}}
+#' function.
+#' 
+#' @param delta Numeric. The expected population effect size.
+#' @param confidence_levell Numeric. The desired level of confidence.
+#' @param width Numeric. The desired width of the confidence interval, given Delta.
+#' @param which_width Character. The desired width of interest. Either `Full`, `Lower` or `Upper`.
+#' @param certainity Numeric. The desired certainty of the confidence interval width.
+#' @param ... Other arguments to be passed to the \code{\link[MBESS]{ss.aipe.smd}} function.
+#' 
+#' @return The function returns a list of one named element. The determined
+#' sample size for group one and two called `n1`.
+#' @examples 
+#' \dontrun{
+#' SampleSizePlanner::ssp_aipe(delta = 0.5, width = 0.2, confidence_level = 0.8)
+#' @export
+ssp_aipe <- function(delta, confidence_level, width, which_width = "Full", certainty = NULL, ...) {
+  n1 <- MBESS::ss.aipe.smd(delta = delta, conf.level = confidence_level, width = width, which.width = which_width, certainty = certainty, ...)
+  list(n1 = n1)
 }

### 2021-03-19 R/ssp_bf_predetermined.R

@@ -1,3 +1,19 @@
+ssp_bf_predetermined <- function(tpr = 0.8, delta, thresh, max_n = 5000) {
+  n1 <- NULL
+  npower <- NULL
+  for (i in 1:length(delta))
+  {
+    res = power_optim(fun = super_bf, range = c(5, max_n),  delta = delta[i], tpr = tpr, thresh = thresh)
+    n1[i] = res$n1
+    npower[i] = res$npower
+  }
+  list(
+    delta = delta,
+    n1 = n1,
+    npower = npower
+  )
+}
+
 super_bf <- function(n1, delta, thresh = 10, tol = 1e-4, granularity = 300) {
   n2 = n1
   t = seq(qt(.005, n1+n2-2, delta/sqrt(1/n1+1/n2)), 
@@ -17,10 +33,4 @@ super_bf <- function(n1, delta, thresh = 10, tol = 1e-4, granularity = 300) {
   npower = pt(t[j], n1+n2-2, delta/sqrt(1/n1+1/n2)) - pt(t[i], n1+n2-2, delta/sqrt(1/n1+1/n2))
   
   npower
-}
-
-ssp_bf_thresh <- function(opt = 0.8, band, delta, thresh) {
-  est <- ssp_tost(opt = opt, band = band, delta = delta) %>% purrr::pluck(., "n1")
-  result <- power_optim(fun = super_bf, range = c(50, est), delta = delta, opt = opt, thresh = thresh)
-  result
 }

### 2021-03-19 R/ssp_bfda.R

@@ -1,26 +1,52 @@
-ssp_bfda <- function(delta, thresh = 10, power = 0.8, n_rep = 1000) {
-  ns = NULL
-  bfs = NULL
+#' Determine sample size with Bayes Factor Design Analysis (BFDA)
+#' 
+#' lorem ipsum
+#' 
+#' @param delta Numeric. the expected population effect size
+#' @param thresh Integer. the Bayes factor threshold for inference
+#' @param tpr Numeric. the long-run probability of obtaining a Bayes factor at least
+#'   as high as the critical threshold favoring superiority, given Delta
+#' @param n_rep Integer. the number of simulations
+#' 
+#' @return A list of two numeric values.
+#' @export
+ssp_bfda <- function(delta, thresh = 10, tpr = 0.8, n_rep = 1000) {
+  Ns = NULL
+  BFs = NULL
   for (i in 1:n_rep) {
-    n1 = n2 = 4
-    plac = rnorm(n1, 0, 1)
-    treat = rnorm(n2, delta, 1)
-    bf = as.vector(BayesFactor::ttestBF(x = treat, y = plac, nullInterval = c(0, Inf), rscale = 1/sqrt(2)))[1]
+    n = 10
+    Plac = rnorm(n, 0, 1)
+    Treat = rnorm(n, delta, 1)
+    BF = BayesFactor::ttest.tstat(
+      t = t.test(Treat, Plac)$statistic,
+      n1 = n,
+      n2 = n, 
+      rscale = 1 / sqrt(2),
+      nullInterval = c(0, Inf),
+      simple = T)
     
-    while (bf < thresh & bf>(1/thresh) & n1<1500) {
-      n1 = n2 = n1 + 1
-      plac = c(plac, rnorm(1, 0, 1))
-      treat = c(treat, rnorm(1, delta, 1))
-      bf = as.vector(BayesFactor::ttestBF(x = treat, y = plac, nullInterval = c(0, Inf), rscale = 1/sqrt(2)))[1]
+    while (BF > (1 / thresh) & BF < thresh & n < 1500) {
+      n = n + 1
+      Plac = c(Plac, rnorm(1, 0, 1))
+      Treat = c(Treat, rnorm(1, delta, 1))
+      BF = BayesFactor::ttest.tstat(
+        t = t.test(Treat, Plac)$statistic,
+        n1 = n,
+        n2 = n, 
+        rscale = 1 / sqrt(2),
[…]

### 2021-03-19 R/ssp_eq_bf.R

@@ -1,42 +1,44 @@
-eq_bf <- function(n1, band, delta, thresh = 10, tol = 1e-4, granularity = 300, prior_df = 1, prior_scale = 1/sqrt(2), prior_location = 0) {
+ssp_eq_bf <- function(tpr, eq_band, delta, thresh) {
+  est <- ssp_tost(tpr = tpr, eq_band = eq_band, delta = delta) %>% purrr::pluck("n1")
+  result <- power_optim(fun = eq_bf, range = c(10, est), delta = delta, tpr = tpr, eq_band = eq_band, thresh = thresh)
+  result
+}
+
+eq_bf <- function(n1, eq_band, delta, thresh = 10, tol = 1e-4, granularity = 300) {
   n2 = n1
   t = seq(qt(.005, n1+n2-2, delta/sqrt(1/n1+1/n2)), 
           qt(.995, n1+n2-2, delta/sqrt(1/n1+1/n2)), length = granularity)
   bf = 1
   i = 0
-  while (bf < thresh)
-  {
+  while (bf < thresh & i < granularity) {
     i = i + 1
-    upper = cdf_t(band, t[i], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
+    upper = cdf_t(eq_band, t[i], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
                    prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)
-    lower = cdf_t(-band, t[i], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
+    lower = cdf_t(-eq_band, t[i], n1 = n1, n2 = n2, independentSamples = T, prior.location = 0, 
                    prior.scale = 1/sqrt(2), prior.df = 1, rel.tol = tol)
     post_dens = upper - lower
-    prior_dens = pcauchy(band, scale = 1/sqrt(2)) - 
-      pcauchy(-band, scale = 1/sqrt(2))
+    prior_dens = pcauchy(eq_band, scale = 1/sqrt(2)) - 
+      pcauchy(-eq_band, scale = 1/sqrt(2))
     bf = (post_dens / prior_dens) / ((1 - post_dens) / (1 - prior_dens))
   }
+  if (i==granularity) {
+    npower = 0
+  } else {
   bf = 1
   j = length(t) + 1
   while (bf < thresh)
   {
     j = j-1
[…]

### 2021-03-19 R/ssp_power_curve.R

@@ -1,28 +1,26 @@
-ssp_power_curve <- function(delta1, delta2, opt, animated = FALSE) {
-  noptim <- NULL
-  delta <- seq(delta1, delta2, 0.01)
+#' Calculating a power curve
+#' 
+#' lorem ipsum
+#' 
+#' @param delta Numeric. the range of hypothetical population effect sizes
+#' @param tpr Numeric. the desired long run probability of obtaining a significant result with TOST, given Delta
+#' @param animated Logical. if TRUE the output plot is animated
+#' 
+#' @return The function returns a ggplot2 object.
+#' @export
+ssp_power_curve <- function(delta, tpr, max_n = 5000) {
+  n1 <- NULL
+  npower <- NULL
+  # delta <- seq(delta1, delta2, 0.01)
   for (i in 1:length(delta)) {
-    noptim[i] = power_optim(fun = traditional, range = c(5, 2000), delta[i], opt = opt)$n1
-    }
-  gg <- ggplot2::ggplot() +
-    ggplot2::aes(x = delta,
-                 y = noptim,
-                 text = paste("n:", noptim,
-                              "<br>delta: ", delta)) +
-    ggplot2::geom_point(shape = 21, colour = "black", fill = "white", size = 3, stroke = 0.5) +
-    ggplot2::scale_y_continuous(limits = c(0, 1500),
-                                labels = c("0", "500", "1000", "1500"),
-                                breaks = c(0, 500, 1000, 1500)) +
-    ggplot2::scale_x_continuous(limits = c(0, delta2),
-                                labels = as.character(delta),
-                                breaks = delta) +
-    ggplot2::labs(x = "Delta",
-                  y = "Planned sample size") +
-    ggplot2::theme_classic()
-  if (animated) {
-    gg %>% 
-    plotly::ggplotly(tooltip = "text")
-  } else {
-    gg
+    res = power_optim(fun = traditional, range = c(5, max_n), delta[i], tpr = tpr)
+    n1[i] = res$n1
+    npower[i] = res$npower
   }
+  
+  list(
+    delta = delta,
+    n1 = n1,
+    npower = npower
+  )
 }

### 2021-03-19 app.R

@@ -1,0 +1,7 @@
+# Launch the ShinyApp (Do not remove this comment)
+# To deploy, run: rsconnect::deployApp()
+# Or use the blue button on top of this file
+
+pkgload::load_all(export_all = FALSE, helpers = FALSE, attach_testthat = FALSE)
+options( "golem.app.prod" = TRUE)
+SampleSizePlanner::run_app() # add parameters here (if any)

### 2021-03-19 data-raw/rope_precalculation_results.R

@@ -40,35 +40,65 @@ rope_data %>%
   count(error_message)
 
 # Create the ouput table
-rope_data <-
-  rope_data %>% 
-  select(iterate, power, delta, band, n1, npower) %>% 
-  arrange(iterate)
+# rope_data <-
+#   rope_data %>% 
+#   select(iterate, power, delta, band, n1, npower) %>% 
+#   arrange(iterate)
 
 rope_result <-
   rope_data %>% 
-  filter(!is.na(n1))
+  filter(!is.na(n1)) %>% 
+  select(iterate, power, delta, band, n1, npower, error_message)
 
-rope_new <- 
-  rope_options %>% 
-  slice(62001:77326)
+# rope_new <- 
+#   rope_options %>% 
+#   slice(62001:77326)
 
-rope_rerun <- 
-  rope_data %>% 
-  filter(is.na(n1)) %>% 
-  select(-n1, -npower) %>% 
-  bind_rows(rope_new)
+# rope_rerun <- 
+#   rope_data %>% 
+#   filter(is.na(n1)) %>% 
+#   select(-n1, -npower) %>% 
+#   bind_rows(rope_new)
 
+# rope_precalculation_results <- 
+#   rope_data %>% 
+#   select(iterate, n1, npower, error_message, power, delta, band) %>% 
+#   rename(tpr = power,
+#          eq_band = band)
+
+rope_rerun_data <- 
+  tibble(filename = list.files(path = "./rope-res-rerun/", pattern = ".rds")) %>%
+  mutate(file = here::here("rope-res-rerun", filename),
+         data = map(file, readRDS)) %>% 
+  unnest(data) %>% 
+  mutate(result = map(data, "result"),
+         error = map(data, "error"),
+         error_message = map_chr(error, ~ pluck(.x, "message", .default = NA_character_)),
+         result_not_null = if_else(map_lgl(result, ~ !is.null(.x)),
+                                   1L,
+                                   0L),
+         n1 = map_dbl(result, ~ pluck(.x, "n1", .default = NA_real_)),
+         npower = map_dbl(result, ~ pluck(.x, "npower", .default = NA_real_)),
+         iterate = as.integer(names(data))) %>% 
+  select(-file, -filename) %>% 
+  left_join(., rope_options, by = "iterate")
+
+# Calculate the number of missing results
+rope_rerun_data %>% 
+  count(result_not_null)
+
+# Join datasets
 rope_precalculation_results <- 
-  rope_data %>% 
[…]

### 2021-03-19 dev/02_dev.R

@@ -9,7 +9,6 @@
 
 golem::add_module(name = "question")
 golem::add_module(name = "article")
-golem::add_module(name = "flowchart")
 golem::add_module(name = "about")
 golem::add_module(name = "tost")
 golem::add_module(name = "menu")
@@ -22,7 +21,8 @@ golem::add_module(name = "bf_threshold")
 golem::add_module(name = "AIPE")
 golem::add_module(name = "BFDA")
 golem::add_module(name = "download")
-golem::add_module(name = "preview")
+golem::add_module(name = "justification")
+golem::add_module(name = "code")
 
 ## 2.2 Add dependencies
 
@@ -36,6 +36,8 @@ usethis::use_package("shinyjs")
 usethis::use_package("shinyBS")
 usethis::use_package("BayesFactor")
 usethis::use_package("assertthat")
+usethis::use_package("MBESS")
+usethis::use_package("rclipboard")
 
 ## 2.3 Add tests
 

### 2021-03-19 informedTtest_functions.R

@@ -1,419 +1,0 @@
-
-integrand_t <- function(delta, t, n, nu, mu.delta, gamma, kappa) {
-  
-  suppressWarnings(
-    dt(x = t, df = nu, ncp = sqrt(n) * delta) *
-    1 / gamma * dt( (delta - mu.delta) / gamma, df = kappa)
-  )
-  
-}
-
-posterior_t <- function(delta,
-                        t,
-                        n1,
-                        n2 = NULL,
-                        independentSamples = FALSE,
-                        prior.location,
-                        prior.scale,
-                        prior.df,
-                        rel.tol = .Machine$double.eps^0.25) {
-  
-  neff <- ifelse(independentSamples, n1 * n2 / (n1 + n2), n1)
-  nu <- ifelse(independentSamples, n1 + n2 - 2, n1 - 1)
-  
-  mu.delta <- prior.location
-  gamma <- prior.scale
-  kappa <- prior.df
-  
-  numerator <- suppressWarnings(
-    dt(x = t, df = nu, ncp = sqrt(neff) * delta) *
-    1 / gamma * dt( (delta - mu.delta) / gamma, df = kappa)
-  )
-  
-  denominator <- integrate(integrand_t,
-                           lower = -Inf, upper = Inf,
-                           t = t, n = neff, nu = nu,
-                           mu.delta = mu.delta,
-                           gamma = gamma,
-                           kappa = kappa,
-                           rel.tol = rel.tol)$value
-  
-  out <- numerator / denominator
-  out[is.na(out)] <- 0
-  
-  return(out)
-  
-}
-
-cdf_t <- function(x,
-                  t,
-                  n1,
-                  n2 = NULL,
-                  independentSamples = FALSE,
-                  prior.location,
-                  prior.scale,
-                  prior.df,
-                  rel.tol = .Machine$double.eps^0.25) {
-  
-  out <- integrate(posterior_t,
-                   lower = -Inf, upper = x,
-                   t = t, n1 = n1, n2 = n2,
-                   independentSamples = independentSamples,
-                   prior.location = prior.location,
-                   prior.scale = prior.scale,
-                   prior.df = prior.df,
-                   rel.tol = rel.tol)$value
-  
-  # catch numerical errors
-  if (out > 1 & out < 1.001) {
-    out <- 1
-    warning(
-      "Numerical integration yields a CDF value slightly larger than 1. The CDF value has been replaced by 1.",
-      call. = FALSE
-    )
-  }
-  
[…]

### 2021-03-23 README.Rmd
updated readme

@@
 
-The goal of SampleSizePlanner is to ...
+The goal of SampleSizePlanner is to help researchers determine the sample size for two group designs.
+The present application and R package offers 9 different sample size planning methods.
+In addition, the web application allows users to create a justification report for their sample size plan.
 
-## Installation
+## Available sample size planning methods
+* Accuracy In Parameter Estimation (AIPE)
+* A-priori precision (APP)
+* Predetermined sample size with Bayes factor
+* Power curve
+* Classical power calculation
+* Interval Equivalence Bayes factor
+* Region of Practical Equivalence (ROPE)
+* Two One‐Sided Tests (TOST)
+* Bayes Factor Design Analysis (BFDA)
 
-You can install the released version of SampleSizePlanner from [CRAN](https://CRAN.R-project.org) with:
+## Usage
+
+SampleSizePlanner can be used either via the web app or via R.
+
+### Using the web app
+
+You can use the app at [https://martonbalazskovacs.shinyapps.io/SampleSizePlanner/](https://martonbalazskovacs.shinyapps.io/SampleSizePlanner/).
+
+You can alternatively run the app locally on your own computer by following these instructions:
+
+Install the development version (SampleSizePlanner is not available from CRAN) from [GitHub](https://github.com/) with:
 
 ``` r
-install.packages("SampleSizePlanner")
+# install.packages("devtools")
+devtools::install_github("marton-balazs-kovacs/SampleSizePlanner")
 ```
 
-## Example
+Running the app.
 
-This is a basic example which shows you how to solve a common problem:
-
-```{r example}
-library(SampleSizePlanner)
-## basic example code
+``` r
+SampleSizePlanner::run_app()
 ```
 
-What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:
+### Using the package
 
-```{r cars}
-summary(cars)
-```
+You can use the R package to determine the sample size with any of the 9 methods that present in the web app.
+However, it is not possible currently to generate a justification report from the R package.
+Please see the tutorial paper for example sample size justifications for each method.
 
-You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.
+## Note
 
-You can also embed plots, for example:
+The tutorial paper that presents the web application and R package is not reviewed yet.
 
-```{r pressure, echo = FALSE}
-plot(pressure)
-```
+## Code of Conduct
 
-In that case, don't forget to commit and push the resulting figure files, so they display on GitHub!
+We are open to new ideas and feature requests.
+
+Please note that the SampleSizePlanner project is released with a [Contributor Code of Conduct](https://marton-balazs-kovacs.github.io/SampleSizePlanner/CODE_OF_CONDUCT.html).
+By contributing to this project, you agree to abide by its terms.

### 2021-03-23 inst/app/www/about.Rmd
updated readme

@@
 
-### Available sample size planning methods
-- AIPE [REF](link)
-- APP [REF](link)
-- BF threshold [REF](link)
-- Power curve [REF](link)
-- Traditional power calculation [REF](link)
-- Interval equiv bf [REF](link)
-- Rope [REF](link)
-- TOST [REF](link)
-- BFDA [REF](link)
-
 ### Contributors
 - Don van Ravenzwaaij
+- Marton Kovacs
 - Rink Hoekstra
-- Marton Kovacs
 - Balazs Aczel

### 2021-03-23 R/mod_ssp_bf_predetermined.R
bug fixes in the app

@@ -24,16 +24,6 @@ mod_ssp_bf_predetermined_ui <- function(id) {
         # Method description
         p("The present method calculates the corresponding default Bayes factor for a t-test statistic with Cauchy prior distribution centered on zero with scale parameter 1/sqrt(2) for several sample sizes."),
         # Calculation settings
-        ## Delta input
-        sliderInput(
-          NS(id, "delta"),
-          list(
-            "Delta",
-            HTML('<i class="fas fa-info"; title="The expected population effect size."></i>')),
-          min = 0,
-          max = 2,
-          value = 0.5,
-          step = 0.1),
         ## TPR input
         sliderInput(
           NS(id, "tpr"),
@@ -43,6 +33,16 @@ mod_ssp_bf_predetermined_ui <- function(id) {
           min = 0,
           max = 1,
           value = 0.8,
+          step = 0.1),
+        ## Delta input
+        sliderInput(
+          NS(id, "delta"),
+          list(
+            "Delta",
+            HTML('<i class="fas fa-info"; title="The expected population effect size."></i>')),
+          min = 0,
+          max = 2,
+          value = 0.5,
           step = 0.1),
         ## Maximum N input
         numericInput(

### 2021-03-23 R/mod_ssp_power_traditional.R

@@ -118,7 +118,7 @@ mod_ssp_power_traditional_server <- function(id) {
     # Calculate results
     traditional_result <- eventReactive(input$calculate, {
       # waitress$start()
-      ssp_power_traditional(max_n = input$max_n, delta = input$delta, tpr = input$tpr)
+      ssp_power_traditional(tpr = input$tpr, delta = input$delta, max_n = input$max_n)
     })
     
     # Show calculated results

### 2021-03-23 R/ssp_bfda.R

@@ -21,7 +21,7 @@
 #' \dontrun{
 #' SampleSizePlanner::ssp_bfda(tpr = 0.8, delta = 0.5, thresh = 10, n_rep = 1000)
 #' }
-ssp_bfda <- function(delta, thresh = 10, tpr = 0.8, n_rep = 1000) {
+ssp_bfda <- function(tpr = 0.8, delta, thresh = 10, n_rep = 1000) {
   Ns = NULL
   BFs = NULL
   for (i in 1:n_rep) {

### 2021-03-23 R/ssp_power_traditional.R

@@ -18,8 +18,7 @@
 #' SampleSizePlanner::ssp_power_traditional(tpr = 0.8, delta = 0.5, max_n = 5000, alpha = 0.05)
 #' }
 ssp_power_traditional <- function(tpr, delta, max_n, alpha = 0.05) {
-  n1 <- power_optim(fun = traditional, range = c(4, max_n), delta = delta, tpr = tpr, alpha = alpha)
-  list(n1 = n1)
+  power_optim(fun = traditional, range = c(4, max_n), delta = delta, tpr = tpr, alpha = alpha)
 }
 
 traditional <- function(n1, delta, alpha = .05) {

### 2021-03-24 R/mod_home.R
updated ui cleaned folders. last vignette build did not run because it is not a finished vignette. it does not have to be perfect now just have to run

@@ -1,0 +1,34 @@
+#' home UI Function
+#'
+#' @description A shiny Module.
+#'
+#' @param id,input,output,session Internal parameters for {shiny}.
+#'
+#' @noRd 
+#'
+#' @importFrom shiny NS tagList 
+mod_home_ui <- function(id){
+  tagList(
+    div(
+      id = "home-page",
+      style="text-align: justify;",
+      h2("How to use the app", style = "margin-bottom: 20px; margin-top: 0; padding-top: 0;"),
+      HTML(
+      "<p>This application provides 9 different methods to determine the sample size
+      for a two group design. Moreover, the app provides help to create an accompanying
+      report text for your sample size determination. The figure below shows the 9 sample size
+      estimation methods in the app and their relations. To get started with the sample size estimation
+      choose a method under the <b>Methods tab</b>. Please visit the <b>About tab</b> if you would like to know more
+      about the project or if you have any suggestions or issues with the app.</p>"),
+      img(src = "www/flowchart.jpg"
+          , width = "100%", style = "margin-bottom: 20px; margin-top: 20px;"
+          ),
+      p("*This version of the application is presented in a tutorial paper that is not reviewed.",
+        style = "text-align: right;")
+    )
+  )
+}
+    
+## To be copied in the UI
+# mod_home_ui("home")
+ 

### 2021-03-24 dev/02_dev.R

@@ -7,11 +7,9 @@
 
 ## 2.1 Add modules
 
-golem::add_module(name = "question")
-golem::add_module(name = "article")
 golem::add_module(name = "about")
+golem::add_module(name = "home")
 golem::add_module(name = "tost")
-golem::add_module(name = "menu")
 golem::add_module(name = "traditional_power")
 golem::add_module(name = "power_curve")
 golem::add_module(name = "apriori_precision")
@@ -38,6 +36,8 @@ usethis::use_package("BayesFactor")
 usethis::use_package("assertthat")
 usethis::use_package("MBESS")
 usethis::use_package("rclipboard")
+usethis::use_package("papaja")
+usethis::use_package("formatR")
 
 ## 2.3 Add tests
 

### 2021-03-30 README.Rmd
last update before submission

@@
 ## Available sample size planning methods
+* Two One‐Sided Tests (TOST)
+* Interval Equivalence Bayes factor
+* Classical power analysis
+* Power curve
+* Bayes Factor Design Analysis (BFDA)
+* Predetermined sample size with Bayes factor
 * Accuracy In Parameter Estimation (AIPE)
 * A-priori precision (APP)
-* Predetermined sample size with Bayes factor
-* Power curve
-* Classical power calculation
-* Interval Equivalence Bayes factor
 * Region of Practical Equivalence (ROPE)
-* Two One‐Sided Tests (TOST)
-* Bayes Factor Design Analysis (BFDA)
 
@@
 However, it is not possible currently to generate a justification report from the R package.
-Please see the tutorial paper for example sample size justifications for each method.
+Please see the tutorial paper or the web app for example sample size justifications for each method.
 
@@
 
-The tutorial paper that presents the web application and R package is not reviewed yet.
+The tutorial paper that presents the web application and R package is not reviewed yet. More details are available in Kovacs, van Ravenzwaaij, Hoekstra & Aczel (2021).
 

### 2021-03-30 inst/app/www/about.Rmd
last update before submission

@@
 The __SampleSizePlanner__ shiny application enables researchers to calculate the sample size with 9 different methods in app and provides code to use an additional method from R.
-The aim of the app is to make the various sample size calculating methods available for a wide range of researchers and propagate the transparent reporting of the choose method.
+The aim of the app is to make the various sample size calculating methods available for a wide range of researchers and propagate the transparent reporting of the chosen method.
 Therefore, a formatted template text containing the used parameters and the output of the calculations is available for download.
@@
 ### Contributors
+- Marton Kovacs
 - Don van Ravenzwaaij
-- Marton Kovacs
 - Rink Hoekstra

### 2021-03-30 R/mod_ssp_aipe.R

@@ -28,9 +28,11 @@ mod_ssp_aipe_ui <- function(id) {
         ## Confidence level input
         sliderInput(
           NS(id, "confidence_level"),
-          list(
-            "Confidence Level",
-            HTML('<i class="fas fa-info"; title="The desired level of confidence."></i>')),
+          HTML(
+            '<div title="The desired level of confidence.">',
+            'Confidence Level',
+            '<i class="fas fa-info"></i>',
+            '</div>'),
           min = 0,
           max = 1,
           value = 0.8,
@@ -38,9 +40,11 @@ mod_ssp_aipe_ui <- function(id) {
         ## Delta input
         sliderInput(
           NS(id, "delta"),
-          list(
-            "Delta",
-            HTML('<i class="fas fa-info"; title="The expected population effect size."></i>')),
+          HTML(
+            '<div title="The expected population effect size.">',
+            'Delta',
+            '<i class="fas fa-info"></i>',
+            '</div>'),
           min = 0,
           max = 2,
           value = 0.5,
@@ -48,9 +52,11 @@ mod_ssp_aipe_ui <- function(id) {
         ## Width input
         sliderInput(
           NS(id, "width"),
-          list(
-            "Width",
-            HTML('<i class="fas fa-info"; title="The desired width of the confidence interval, given Delta."></i>')),
+          HTML(
+            '<div title="The desired width of the confidence interval, given Delta.">',
+            'Width',
+            '<i class="fas fa-info"></i>',
+            '</div>'),
           min = 0,
           max = 1,
           value = 0.2,

### 2021-03-30 R/mod_ssp_app.R

@@ -27,9 +27,11 @@ mod_ssp_app_ui <- function(id) {
         ## Closeness input
         sliderInput(
           NS(id, "closeness"),
-          list(
-            "Closeness",
-            HTML('<i class="fas fa-info"; title="The desired closeness of the sample mean to the population mean defined in standard deviation."></i>')),
+          HTML(
+            '<div title="The desired closeness of the sample mean to the population mean defined in standard deviation.">',
+            'Closeness',
+            '<i class="fas fa-info"></i>',
+            '</div>'),
           min = 0,
           max = 1,
           value = 0.2,
@@ -37,9 +39,11 @@ mod_ssp_app_ui <- function(id) {
         ## Confidence input
         sliderInput(
           NS(id, "confidence"),
-          list(
-            "Confidence",
-            HTML('<i class="fas fa-info"; title="The desired probability of obtaining the sample mean with the desired closeness to the population mean."></i>')),
+          HTML(
+            '<div title="The desired probability of obtaining the sample mean with the desired closeness to the population mean.">',
+            'Confidence',
+            '<i class="fas fa-info"></i>',
+            '</div>'),
           min = 0, 
           max = 1,
           value = 0.95, 

### 2021-03-30 R/mod_ssp_rope.R

@@ -27,27 +27,33 @@ mod_ssp_rope_ui <- function(id){
         ## TPR input
         sliderInput(
           NS(id, "tpr"),
-          list(
-            "True Positive Rate (TPR)",
-            HTML('<i class="fas fa-info"; title="The desired long run probability of having the HDI fully contained within the ROPE interval, given Delta."></i>')),
+          HTML(
+            '<div title="The desired long run probability of having the HDI fully contained within the ROPE interval, given Delta.">',
+            'True Positive Rate (TPR)',
+            '<i class="fas fa-info"></i>',
+            '</div>'),
           min = 0.5,
           max = 0.95,
           value = 0.8,
           step = 0.01),
         sliderInput(
           NS(id, "eq_band"),
-          list(
-            "Equivalence Band (EqBand)",
-            HTML('<i class="fas fa-info"; title="The chosen ROPE interval."></i>')),
+          HTML(
+            '<div title="The chosen ROPE interval.">',
+            'Equivalence Band (EqBand)',
+            '<i class="fas fa-info"></i>',
+            '</div>'),
           min = 0.1,
           max = 0.5,
           value = 0.2,
           step = 0.01),
         sliderInput(
           NS(id, "delta"),
-          list(
-            "Delta",
-            HTML('<i class="fas fa-info"; title="The expected population effect size."></i>')),
+          HTML(
+            '<div title="The expected population effect size.">',
+            'Delta',
+            '<i class="fas fa-info"></i>',
+            '</div>'),
           min = 0,
           max = 0.5,
           value = 0,
@@ -138,9 +144,9 @@ mod_ssp_rope_server <- function(id){
     output$calculate_output <- renderUI({
       HTML(
         glue::glue(
-          "<b>n1:</b> {n1}<br/><b>npower:</b> {npower}",
+          "<b>n1:</b> {n1}<br/><b>Resulting TPR:</b> {npower}",
           n1 = rope_result()$n1,
-          npower = rope_result()$npower
+          npower = round(rope_result()$npower, 2)
         )
       )
     })

### 2021-03-30 R/mod_ssp_tost.R

@@ -27,19 +27,23 @@ mod_ssp_tost_ui <- function(id) {
         ## TPR input
         sliderInput(
           NS(id, "tpr"),
-          list(
+          HTML(
+            '<div title="The desired long run probability of obtaining a significant result with TOST, given Delta.">',
             "True Positive Rate (TPR)",
-            HTML('<i class="fas fa-info"; title="The desired long run probability of obtaining a significant result with TOST, given Delta."></i>')),
+            '<i class="fas fa-info"></i>',
+            '</div>'),
           min = 0,
           max = 1,
           value = 0.8,
-          step = 0.1),
+          step = 0.01),
         ## EqBand input
         sliderInput(
           NS(id, "eq_band"),
-          list(
-            "Equivalence Band (EqBand)",
-            HTML('<i class="fas fa-info"; title="The chosen width of the region for practical equivalence, i.e. the SESOI."></i>')),
+          HTML(
+            '<div title="The chosen width of the region for practical equivalence, i.e. the SESOI.">',
+            'Equivalence Band (EqBand)',
+            '<i class="fas fa-info"></i>',
+            '</div>'),
           min = 0,
           max = 1,
           value = 0.2,
@@ -47,9 +51,11 @@ mod_ssp_tost_ui <- function(id) {
         ## Delta input
         sliderInput(
           NS(id, "delta"),
-          list(
+          HTML(
+            '<div title="The expected population effect size. In most cases, this value will be zero.">',
             "Delta",
-            HTML('<i class="fas fa-info"; title="The expected population effect size. In most cases, this value will be zero."></i>')),
+            '<i class="fas fa-info"></i>',
+            '</div>'),
           min = 0,
           max = 2, 
           value = 0, 
@@ -84,7 +90,7 @@ mod_ssp_tost_ui <- function(id) {
                 NS(id, "eq_band_justification"),
                 label = "Equivalence Band (EqBand)",
                 choices = c(
-                  "previous studies reported a similar SESOI ...",
+                  "previous studies reported a similar equivalence band ...",
                   "of the following substantive reasons: ...",
                   "other..."),
                 multiple = FALSE,
@@ -134,9 +140,9 @@ mod_ssp_tost_server <- function(id) {
     output$calculate_output <- renderUI({
       HTML(
         glue::glue(
-          "<b>n1:</b> {n1}<br/><b>npower:</b> {npower}",
[…]

### 2021-03-30 R/ssp_bfda.R

@@ -1,8 +1,9 @@
 #' Determine sample size with Bayes Factor Design Analysis (BFDA)
 #' 
-#' The present method estimates the long-run rates of misleading
-#' evidence that one can expect for a specific research design if
-#' using preset Bayes Factor thresholds and allowing for sequential testing.
+#' The present method provides an expected sample size such that
+#' compelling evidence in the form of a Bayes factor can be collected
+#' for a given effect size with a certain long-run probability when
+#' allowing for sequential testing.
 #' 
 #' @param delta Numeric. The expected population effect size.
 #' @param thresh Integer. The Bayes factor threshold for inference.

### 2021-03-30 R/app_ui.R
added papaja vignette

@@ -12,7 +12,7 @@ app_ui <- function() {
                           mod_ssp_tost_ui("tost")),
                  tabPanel("Interval Equivalence BF",
                           mod_ssp_eq_bf_ui("eq_bf")),
-                 tabPanel("Traditional power",
+                 tabPanel("Classical power",
                           mod_ssp_power_traditional_ui("traditional")),
                  tabPanel("Power curve",
                           mod_ssp_power_curve_ui("curve")),

### 2021-03-30 R/mod_home.R

@@ -20,7 +20,7 @@ mod_home_ui <- function(id){
       estimation methods in the app and their relationships. To get started with the sample size estimation
       choose a method under the <b>Methods tab</b>. Please visit the <b>About tab</b> if you would like to know more
       about the project or if you have any suggestions or issues with the app.</p>"),
-      img(src = "www/flowchart.jpg"
+      img(src = "www/flowchart_web.jpg"
           , width = "100%", style = "margin-bottom: 20px; margin-top: 20px;"
           ),
       p("*This version of the application is presented in a tutorial paper that is not reviewed.</br>

### 2021-07-01 inst/manuscript/manuscript.Rmd
updated logo, documentation, prior arguments added for Bayesian methods,  rearranged informed t test functions in R folder

---
title             : "SampleSizePlanner: A Tool to Estimate and Justify Sample Size for Two-Group Studies"
shorttitle        : "SAMPLESIZEPLANNER: A TOOL TO ESTIMATE AND JUSTIFY SAMPLE SIZE"

author: 
  - name          : "Marton Kovacs"
    affiliation   : "1,2"
    corresponding : yes    
    email         : "marton.balazs.kovacs@gmail.com"
    address       : "Hungary 1075 Budapest Kazinczy street 23-27"
  - name          : "Don van Ravenzwaaij"
    affiliation   : "3"
  - name          : "Rink Hoekstra"
    affiliation   : "3"
  - name          : "Balazs Aczel"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Institute of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary"
  - id            : "2"
    institution   : "Doctoral School of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary"
  - id            : "3"
    institution   : "University of Groningen, Groningen, The Netherlands"

authornote: |

  Marton Kovacs and Don van Ravenzwaaij are shared first authors.

abstract: |
  Planning sample size often requires researchers to identify a statistical technique
  and to make several choices during their calculations. Currently, there is a lack of clear guidelines
  for researchers to find and use the applicable procedure. In the present tutorial,
  we introduce a web app and R package that offer nine different procedures to determine and
  justify the sample size for independent two-group study designs. The application highlights
  the most important decision points for each procedure and suggests example justifications for them.
  The resulting sample size report can serve as a template for preregistrations and manuscripts.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "sample size determination; power analysis; study design"

bibliography      : ["ssp.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
linkcolor         : "blue"
mask              : no
draft             : no
figsintext        : yes
numbersections    : true

params:
  highlight_changes: TRUE

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_docx
---

```{r setup, include = FALSE}
if(!"tinytex" %in% rownames(installed.packages())) install.packages("tinytex")

# tinytex::install_tinytex()

# Install devtools package if necessary
if(!"devtools" %in% rownames(installed.packages())) install.packages("devtools")

# Install the stable development verions from GitHub
# devtools::install_github("crsh/papaja")

library("papaja")
library("formatR")
library("kableExtra")
devtools::install_github("marton-balazs-kovacs/SampleSizePlanner")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```

# Introduction {.unnumbered}

Social and behavioral sciences are known to be plagued by undersampling [@ioannidis_why_2005]. In the traditional statistical framework, even when the effect exists, undersampled studies yield either nonsignificant results or significant results due to overestimating the size of the effect. Because nonsignificant results are less likely to reach publications than significant ones, results of undersampled studies either remain unpublished or impose a substantial bias on our body of published empirical findings. In addition, the low informational value of undersampled studies may not justify the cost or potential risk they induce [@halpern_continuing_2002]. To mitigate these issues, authors are increasingly expected to plan and justify the sample size of their study [@maxwell_persistence_2004]. However, such sample size justifications are only meaningful if they provide sufficient information to the readers to judge the adequacy of the author's decisions.

In the statistical literature, a few methods have been proposed to determine and justify sample size. In practice, however, authors are short of practical guides on how to navigate among the different sample size methods. The aim of our tutorial is to point out for each method the essential decision points that a researcher has to face during this process. We also provide a collection of ready-to-use analysis code and a ShinyApp that helps researchers use and report the main sample size estimation techniques for different scenarios. The tutorial is focused exclusively on the scenario of the comparison of two independent groups (i.e., the independent *t*-test design) with a one-sided test.

# Sample Size Determination and Justification {.unnumbered}

A lot of factors go into the determination of the sample size for an independent two group study design. In this section, we will first provide a birds-eye view of the most important decisions. Next, we will go into more detail on the specific inference tool that results from the combination of the larger choices.

It is crucial to not just state how we determined our planned sample size but to also give the reader insight into the reasons behind our choices. In a recent overview, @lakens_sample_2021 lists six types of general approaches to justify sample size in quantitative empirical studies: (1) Measure entire population; (2) Resource constraints; (3) A-priori power analysis; (4) Accuracy; (5) Heuristics; and (6) No justification. For the first approach, no quantitative justification is necessary; and for the second approach, the researcher has no freedom to increase the sample size. Power analysis, or more generally the estimation of true positive rate, is used when one plans to conduct hypothesis testing; accuracy justifications are used when one plans to conduct parameter estimation. Our tutorial mainly focuses on approaches two, three, and four, and is aimed at providing a hands-on approach for the mechanical part of the sample size determination (i.e., the calculation). For a deeper discussion of justification of these approaches, or for other approaches (i.e., using heuristics or not providing justification), we refer the reader to @lakens_sample_2021.

## Choosing a method in case of sample size justification {.unnumbered}

In an ideal world, the choice for the number of participants would be solely determined by scientific considerations, and depending on the chosen technique the collection of data would continue until either the desired sample size or a desired outcome has been reached. In practice, researchers are limited by time (collecting data is quite demanding), money (participants or people collecting the data may be paid, and the same may hold for renting space or equipment), or availability of participants (the population may be relatively small, and/or the participation rate quite low).

When constrained by limited resources, it is important to be transparent about those limitations. It is also important to be open about scientific considerations. Depending on the nature of the study (perhaps it is an initial exploration?), small sample sizes need not be a dealbreaker. So although more data are always preferred from an informational point of view, by owning the limitations of our study, we improve future readers' understanding of the process leading up to the eventual paper, and we also answer in advance to those who think the chosen sample size was insufficient.

Whether or not authors have limited resources, two important choices need to be made: (1) whether they are interested in *statistical testing* or in *parameter estimation*; and (2) whether they want to conduct their statistical inference within the *frequentist* framework or within the *Bayesian* framework. Starting with the first decision, statistical testing is the primary framework when one is interested in establishing whether an underlying population effect is equal to, different from, larger than, or smaller than a certain value. In essence, statistical testing lends itself to binary decision making. Typically, testing is concerned with a fixed point null hypothesis (e.g., there is no difference between two groups), although using intervals for testing is also possible. Alternatively, one might be interested in parameter estimation that is less interested in establishing the existence of a difference and instead is concerned with establishing the magnitude of the difference.

The second important decision concerns the statistical framework. Choosing to conduct statistical tests within a frequentist framework, one is usually interested in balancing the type I (false positive) and type II (false negative) error rates. Practitioners choosing to conduct statistical tests within a Bayesian framework are typically interested in being able to quantify the relative probability of hypotheses or models being true given the data and in including prior information.

Within the realm of statistical testing, there are some other factors that affect the preferred inference tool: Do you prefer to test for equivalence (no difference in mean) or for superiority (mean of one group larger than mean of other group), are you interested in calculating a required sample size for a specific hypothetical effect size or for a range of possible values, and do you wish to employ sequential testing (applicable to Bayesian testing)? For frequentist estimation, the preferred inference tool might differ depending on whether we evaluate uncertainty for each group separately or jointly. We will describe these specific factors when we go into detail about each of the preferred methods. A flow-chart representing all of these choices is given in Figure \@ref(fig:flowchart).

```{r flowchart, fig.cap = "(ref:flowchart-caption)", fig.align = "center", echo = FALSE, out.width="100%"}
knitr::include_graphics("./flowchart.jpg", dpi = 300)
```

(ref:flowchart-caption) The figure depicts the decisions that one faces when choosing among sample size estimation methods. The nine sample size estimation methods discussed in this paper are listed in the bottom row. Some decisions are determined by the investigated question and the design of the study while others are based on the preferred statistical framework.

## How to use this guide {.unnumbered}

In the next section, we will illustrate the specific inference tools and resulting sample size calculations in more detail using a ShinyApp and an R package we have developed. Throughout this section, we recurrently use two terms that have different meanings for different techniques. These are the *true positive rate* (TPR), and the *equivalence band* (EqBand). The TPR reflects the long-run probability of concluding there is an effect, given that it does exist. For traditional null hypothesis testing, this is typically referred to as power, but related concepts exist for different inference tools. The EqBand refers to an effect size region, typically around zero, that is deemed clinically insignificant or irrelevant. Different names are given to this region depending on the technique that employs them, such as statistical effect size of interest (SESOI) or region of practical equivalence (ROPE). For both TPR and EqBand, we explain the specific meaning in context of the relevant inference tool below.

For each method, only the main parameters can be adjusted with a certain range of values in the ShinyApp by using a slider. These parameters are presented in the text in bold. Other parameters are set to preset values in the application but can be adjusted in the accompanying R package to any sensible value. These parameters are highlighted in italics in the tutorial. Both the app and the package allow the users to save or copy a text template with the results of the sample size determination. We offer a list of possible justifications at the decision points for each method (indicated between square brackets), but users are able to provide their own justification as free-text. The provided justification text could serve as a stub for the description of the chosen sample size in a paper, a preregistration or registered report, or a grant proposal.

Throughout, we will use the example story of Mary the educational psychologist. Mary has come up with a new set of games that challenge spatial insight. She believes that distributed and targeted engagement with these games for a period of six months for children in the age range of 8 to 12 will lead to lasting improvements on their IQ score as measured through Raven's progressive matrices test (population mean 100, population SD 15). Mary collects data for a control sample that gets regular education and for an experimental sample and plans to compare those samples. For illustrative purposes, we will have Mary's hypothetical study goal depend on the earlier presented decisions to illustrate different scenarios for each sample size planning method.

The ShinyApp is available on <https://martonbalazskovacs.shinyapps.io/SampleSizePlanner> and the R package can be installed by running the following command in R `devtools::install_github("marton-balazs-kovacs/SampleSizePlanner")`. There is more information about the R package and the ShinyApp on the projects' Github page <https://github.com/marton-balazs-kovacs/SampleSizePlanner>, or on the website <https://marton-balazs-kovacs.github.io/SampleSizePlanner/>.

## 1. Testing {.unnumbered}

### 1.1. Effect size = 0 {.unnumbered}

#### 1.1.1. Two One‐Sided Tests (TOST) {.unnumbered}

##### Study context {.unnumbered}

Mary would like to know what sample size she needs for a power of .80 to study whether the mean IQ score of the experimental group's population is practically equivalent to the mean IQ score of the control group. She tests this assumption in a frequentist framework, and considers a population effect size between -0.2 and 0.2 to be 'practically equivalent' to no difference. This would correspond to IQ scores between 97 (100+15\*-.2) and 103 (100+15\*.2).

##### Description {.unnumbered}

TOST is a frequentist equivalence testing approach that adopts two one-sided hypotheses to designate an interval hypothesis [@schuirmann_comparison_1987]. The lower and upper boundaries of the interval are determined by the equivalence band (i.e. SESOI) around the expected population effect size (e.g., 0). @lakens_equivalence_2018 lists several methods that can be used to determine the SESOI . In case of TOST, the two null hypotheses state that the effect size is equal to the lower and upper equivalence band values, whereas the alternative hypotheses state that the effect size is significantly smaller than the upper equivalence band value and significantly larger than the lower equivalence band value. In case both one-sided tests reject the null-hypotheses at a given significance level, the group means are considered to be practically equivalent. See @lakens_equivalence_2018, for further reading.

##### Parameters {.unnumbered}

**Delta:** The expected population effect size. In most cases, this value will be zero.\
**TPR:** The desired long run probability of obtaining a significant result with TOST, given Delta.\
**EqBand:** The chosen width of the region for practical equivalence, i.e. the SESOI.\
*Alpha:* The level of significance. The alpha level in the application is preset to 0.05.\

##### How to use the package {.unnumbered}

```{r, eval = FALSE, tidy = TRUE, echo = TRUE}
SampleSizePlanner::ssp_tost(tpr = 0.8, eq_band = 0.2, delta = 0)
```

##### How to report your sample size estimation {.unnumbered}

In order to calculate an appropriate sample size for testing whether the two groups are practically equivalent, we used the Two One-Sided Tests of Equivalence [TOST; @schuirmann_comparison_1987] method. We set the aimed TPR to be 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement]. We consider all effect sizes below 0.2 equivalent to zero, because [1) previous studies reported a similar equivalence band; 2) of the following substantive reasons: ...]. The expected delta was 0 because [1) we expected no difference between the groups]. Based on these parameters, a sample size of 429 per group was estimated in order to reach a TPR of 0.8 with our design.

#### 1.1.2. Equivalence interval Bayes factor {.unnumbered}

##### Study context {.unnumbered}

Mary would like to know what sample size she needs to have a long-run probability of .80 of obtaining a Bayes factor larger than 10. Mary would like to test whether the mean IQ score of the experimental group's population is practically equivalent to the mean IQ score of the control group. Mary hypothesizes that there is no difference (i.e., H~0~ is true). Mary tests this assumption in a Bayesian framework. Mary considers a population effect size between -0.2 and under 0.2 to be 'practically equivalent'. This would correspond to IQ scores between 97 (100+15\*-.2) and 103 (100+15\*.2).

##### Description {.unnumbered}

Equivalence interval Bayes factors contrast an equivalence hypothesis to a non-equivalence hypothesis and quantify the evidence with Bayes factors. Typically, H~0~ constitutes the equivalence interval (comparable to SESOI in the TOST framework), and H~a~ constitutes the complementary non-equivalence regions. Formally, the Bayes factor is calculated by dividing the fraction *posterior area inside the interval/posterior area outside the interval* (i.e., the posterior odds) by the fraction *prior area inside the interval/prior area outside the interval* (i.e., the prior odds). The resulting value quantifies how much more likely it is that the data occurred under a population effect size deemed 'equivalent' relative to the data having occurred under a population effect size deemed non-equivalent. The current implementation uses a default Cauchy prior on effect size with scale parameter 1/$\sqrt{2}$. For further reading, see @morey_bayes_2011 and @van_ravenzwaaij_bayes_2019.

##### Parameters {.unnumbered}

**Delta:** The expected population effect size.\
**TPR:** The desired long-run probability of obtaining a Bayes factor at least as high as the Threshold, given Delta.\
**EqBand:** The chosen width of the equivalence region.\
*Threshold:* Critical threshold for the Bayes factor. The threshold level in the application can be set to 10, 6, or 3.\

##### How to use the package {.unnumbered}

```{r, eval = FALSE, tidy = TRUE, echo = TRUE}
SampleSizePlanner::ssp_eq_bf(
  tpr = 0.8, delta = 0,
  eq_band = 0.2, thresh = 10
  )
```

##### How to report your sample size estimation {.unnumbered}

In order to estimate the sample size, we used the interval equivalent Bayes factor [@morey_bayes_2011; @van_ravenzwaaij_bayes_2019] method. We set the aimed TPR at 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement]. We consider all effect sizes below 0.2 equivalent to zero, because [1) previous studies reported a similar equivalence region; 2) of the following substantive reasons: ...]. The expected delta was 0 because [1) we expected no difference between the groups]. Our Bayes factor threshold for concluding equivalence was 10. Based on these parameters, a minimal sample size of 144 per group was estimated in order to reach 0.8 TPR for our design.

### 1.2. Effect size \>0 {.unnumbered}

#### 1.2.1. Frequentist {.unnumbered}

##### 1.2.1.1. Classical power analysis {.unnumbered}

###### Study context {.unnumbered}

Mary would like to know what sample size she needs for a power of .80 to study whether the mean IQ score of the experimental group's population is significantly higher than the mean IQ score of the control group. She tests this assumption in a frequentist framework for a hypothetical population effect size of 0.5. This corresponds to a mean IQ score of 107.5 in the experimental group (100+15\*.5), assuming a mean IQ score of 100 in the control group.

###### Description {.unnumbered}

The classical power analysis approach allows to calculate the required sample size in order to obtain a significant result for the null hypothesis test a certain proportion of times in the long run given an assumed population effect size.

###### Parameters {.unnumbered}

**Delta:** The expected population effect size.\
**TPR:** The desired long-run probability of obtaining a significant result with a one-sided *t*-test, given Delta.\
**Maximum N:** The maximum number of participants per group (both groups are assumed to have equal sample size).\
*Alpha:* The level of significance. Alpha is preset to 0.05 in the application.\

###### How to use the package {.unnumbered}

```{r, eval = FALSE, tidy = TRUE, echo = TRUE}
SampleSizePlanner::ssp_power_traditional(
  tpr = 0.8, delta = 0.5,
  max_n = 5000, alpha = 0.05
  )
```

###### How to report your sample size estimation {.unnumbered}

We used a power analysis to estimate the sample size. We set the aimed TPR at 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement]. The expected delta was 0.5 because [1) previous results published in ...; 2) of the following substantive reasons: ...]. Based on these parameters, a minimal sample size of 51 per group was estimated in order to reach 0.8 TPR for our design.

##### 1.2.1.2. Power curve {.unnumbered}

###### Study context {.unnumbered}

Mary would like to know what sample size she needs for a power of .80 to study whether the mean IQ score of the experimental group's population is significantly higher than the mean IQ score of the control group. She tests this assumption in a frequentist framework. However, she is reluctant to commit to a single hypothetical population effect size a-priori, preferring to calculate required sample size for a range of hypothetical deltas between 0.1 and 0.9.

###### Description {.unnumbered}

The power curve method is similar to a classical power analysis but instead of calculating the appropriate sample size for one hypothesized population effect size, the method calculates the required sample size for a range of plausible population effect sizes.

###### Parameters {.unnumbered}

**Delta:** A range of hypothetical population effect sizes.\
**TPR:** The desired long-run probabilities of obtaining a significant result with a one-sided *t*-test, given each value of Delta.\
**Maximum N:** The maximum number of participants per group (both groups are assumed to have equal sample size).\
*Alpha:* The level of significance. Alpha is preset to 0.05 in the application.\

###### How to use the package {.unnumbered}

```{r, eval = FALSE, tidy = TRUE, echo = TRUE}
# Determine the sample sizes for each delta
curve_data <- SampleSizePlanner::ssp_power_curve(
  tpr = 0.8, delta = seq(0.1, 0.9, 0.01), max_n = 5000
  )

# Plot the power curve
SampleSizePlanner::plot_power_curve(
  delta = curve_data$delta, n1 = curve_data$n1, animated = FALSE
  )
```

###### How to report your sample size estimation {.unnumbered}

We used a power analysis to estimate the sample size. We set the aimed TPR at 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement]. Because [1) we have no clear expectation of the magnitude of delta 2) we expected the delta to be around...], we include power calculations for delta ranging from 0.1 to 0.9. Based on these parameters, minimal sample sizes per group for different hypothetical effect sizes to reach 0.8 TPR can be found in Figure \@ref(fig:powercurve).

```{r powercurve, echo = FALSE, fig.cap = "(ref:powercurve-caption)", fig.align = "center",}
# Determine the sample sizes for each delta
curve_data <- SampleSizePlanner::ssp_power_curve(
  tpr = 0.8, delta = seq(0.1, 0.9, 0.01), max_n = 5000
  )
# Plot the power curve
SampleSizePlanner::plot_power_curve(
  delta = curve_data$delta, n1 = curve_data$n1, animated = FALSE
  )
```

(ref:powercurve-caption) The figure shows the resulting power curve created by the application. The X-axis shows the range of deltas from the example, while the Y-axis shows the corresponding sample sizes determined by the power curve method.

#### 1.2.2. Bayesian {.unnumbered}

##### 1.2.2.1. Predetermined sample size with Bayes factor {.unnumbered}

###### Study context {.unnumbered}

Mary would like to test whether the mean IQ score of the experimental group's population is higher than the mean IQ score of the control group. She'd like to know what sample size she needs to have for a long-run probability of .80 of obtaining a Bayes factor larger than 10. Mary plans to collect all her data in one batch without testing sequentially. Mary expects the population effect size to be 0.5. This corresponds to a mean IQ score of 107.5 (100+15\*.5) in the experimental group, assuming a mean IQ score of 100 in the control group.

###### Description {.unnumbered}

The present method calculates the corresponding default Bayes factor for a *t*-test statistic with Cauchy prior distribution centered on zero with scale parameter 1/$\sqrt{2}$ for several sample sizes [the so-called Jeffrey-Zellner-Siow Bayes factor, see e.g., @rouder_bayesian_2009]. The function returns the optimal sample size needed to reach the TPR for a given Bayes factor threshold to detect an expected population effect size. If a range of possible population effect sizes are plausible under the given hypothesis, the function can calculate the optimal sample sizes for the given range of effect sizes and present the results in a figure (analogous to the Power Curve method).

###### Parameters {.unnumbered}

**Delta:** The expected population effect size or a range of expected effect sizes.\
**TPR:** The long-run probability of obtaining a Bayes factor at least as high as the critical threshold favoring superiority, given Delta.\
**Maximum N:** The maximum number of participants per group (both groups are assumed to have equal sample size).\
**Threshold:** Critical threshold for the Bayes factor. Three threshold levels are available in the app: 3, 6, and 10.\

###### How to use the package {.unnumbered}

```{r, eval = FALSE, tidy = TRUE, echo = TRUE}
SampleSizePlanner::ssp_bf_predetermined(
  tpr = 0.8, delta = 0.5, thresh = 10, max_n = 5000
  )
```

###### How to report your sample size estimation {.unnumbered}

We used the Jeffrey-Zellner-Siow Bayes factor method to estimate the sample size. We set the aimed TPR at 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement]. The expected delta was 0.5 because [1) previous results published in ...; of the following substantive reasons: ...]. Our evidence threshold was 10. Based on these parameters, a minimal sample size of 105 per group was estimated in order to reach a 0.8 TPR for our design.

##### 1.2.2.2. Bayes Factor Design Analysis (BFDA) {.unnumbered}

###### Study context {.unnumbered}

Mary would like to know what sample size she needs to have a long-run probability of .80 of obtaining a Bayes factor larger than 10. Mary would like to test whether the mean IQ score of the experimental group's population is higher than the mean IQ score of the control group in a Bayesian framework. Mary plans to collect all her data incrementally and as such is interested in using the advantage of not testing more than strictly necessary offered by sequential testing in her Bayesian analysis. Mary expects the population effect size to be 0.5. This corresponds to a mean IQ score of 107.5 in the experimental group (100+15\*.5), assuming a mean IQ score of 100 in the control group.

###### Description {.unnumbered}

The description of the BFDA method is functionally identical to the one provided in section 'Predetermined sample size with Bayes factor', but gains in TPR due to the addition of sequential testing. In the app, H~0~ and H~a~ indicate the proportion of times sequential testing leads to Bayes factors providing evidence with the given threshold for the null hypothesis and for the alternative hypothesis, respectively. For further reading, see @schonbrodt_bayes_2018 and @schonbrodt_sequential_2017.

###### Parameters {.unnumbered}

**Delta:** The expected population effect size.\
**TPR:** The long run probability of obtaining a Bayes factor at least as high as the critical threshold favoring superiority, given Delta.\
**Threshold:** Critical threshold for the Bayes factor. Three threshold levels are available in the app: 3, 6, and 10.\

###### How to use the package {.unnumbered}

```{r, eval = FALSE, tidy = TRUE, echo = TRUE}
SampleSizePlanner::ssp_bfda(
  tpr = 0.8, delta = 0.5, thresh = 10, n_rep = 1000
  )
```

###### How to report your sample size estimation {.unnumbered}

We used the BFDA method to estimate the sample size. We set the aimed TPR at 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement]. The expected delta was 0.5 because [1) previous results published in ...; 2) of the following substantive reasons: ...]. Our evidence threshold was 10. Based on these parameters, a minimal sample size of 81 per group was estimated in order to reach a 0.8 TPR for our design.

## 2. Estimation {.unnumbered}

### 2.1. Frequentist {.unnumbered}

#### 2.1.1. Accuracy In Parameter Estimation (AIPE) {.unnumbered}

##### Study context {.unnumbered}

Mary would like to know what sample size she needs, such that the 95% confidence interval for the population effect size has an expected width of 0.4. She estimates the population effect size to be 0.2.

##### Description {.unnumbered}

Accuracy in parameter estimation aims to determine the sufficient sample size to obtain a confidence interval with a desired width (precision) around the expected effect size [@kelley_sample_2006]. Note that the width of the calculated confidence interval will depend on the sample variance. As a result, it is possible that for a given sample the variance is relatively large, leading to a resulting confidence interval that is larger than the width of the desired interval for a given sample. Thus, the AIPE method aims to establish the expected value of the calculated confidence interval, which can be thought of as the 50% long-run probability of obtaining a confidence interval no wider than the provided width.

#### Parameters {.unnumbered}

**Delta:** The expected population effect size.\
**Width:** The desired width of the confidence interval, given Delta.\
**Confidence level:** The desired level of confidence.\

##### How to use the package {.unnumbered}

```{r, eval = FALSE, tidy = TRUE, echo = TRUE}
SampleSizePlanner::ssp_aipe(
  delta = 0.5, width = 0.2, confidence_level = 0.8
  )
```

##### How to report your sample size estimation {.unnumbered}

In order to estimate the sample size, we used the accuracy in parameter estimation [AIPE; @kelley_sample_2006] method. We aimed for a 95% confidence level, because [1) it is the common standard in the field; 2) it is the journal publishing requirement]. The desired width was 0.4 because [1) previous studies on this topic reported a similar region of practical equivalence; 2) of the following substantive reasons: ...]. We expected an underlying population effect size of 0.3, because [1) previous results published in ...; 2) of the following substantive reasons: ...]. Based on these parameters, a minimal sample size of 195 per group was estimated for our design.

#### 2.1.2. A Priori Precision (APP) {.unnumbered}

###### Study context {.unnumbered}

Mary would like to know the sample size for which she will have a 95% long-run probability that the sample means in both the experimental and the control group lie within 0.2 standard deviations (3 IQ points) of the true population mean.

###### Description {.unnumbered}

APP aims to determine the sample size needed to have a certain long-run probability of both sample means being within a certain range of their respective population means, expressed in terms of standard deviations [@trafimow_performing_2017]. As a result, APP is not reliant on the expected effect size.

##### Parameters {.unnumbered}

**Closeness:** The desired closeness of the sample mean to the population mean defined in standard deviation.\
**Confidence:** The desired probability of obtaining the sample mean with the desired closeness to the population mean.\

###### How to use the package {.unnumbered}

```{r, eval = FALSE, tidy = TRUE, echo = TRUE}
SampleSizePlanner::ssp_app(
  closeness = 0.2, confidence = 0.95
  )
```

###### How to report your sample size estimation {.unnumbered}

In order to estimate the sample size, we used the a-priori precision [APP; @trafimow_performing_2017] method. Before data collection, we wanted to be 95% confident that both sample means lie within 0.2 SD of the true population means. Based on these parameters, the resulting minimum sample size was 126 per group for our design.

### 2.2. Bayesian testimation {.unnumbered}

#### 2.2.1. Region of Practical Equivalence (ROPE) {.unnumbered}

###### Study context {.unnumbered}

Mary would like to conduct parameter estimation to see whether the mean IQ score of her experimental group's population is practically equivalent to 100. She would like to know what sample size she needs to have a long-run probability of .80 of obtaining a 95% highest density interval that is contained within her predefined region of practical equivalence (ROPE). Mary hypothesizes that there is no difference (i.e., H~0~ is true). She considers a population effect size between -0.2 and under 0.2 to be 'practically equivalent'. This would correspond to IQ scores between 97 (100+15\*-.2) and 103 (100+15\*.2).

###### Description {.unnumbered}

The highest density interval region of practical equivalence technique (HDI-ROPE, often just referred to as ROPE) shares some features with the equivalence interval Bayes factor procedure. Both define an equivalence interval, construct a prior for the population effect size, and update to a posterior after the data comes in. The equivalence interval Bayes factor procedure then focuses on the posterior and prior odds under complementary hypotheses. The ROPE procedure, on the other hand, identifies the 95% highest density interval (HDI; other percentages are permissible as well) and determines whether or not the HDI is fully contained within the equivalence interval. For further reading, see @kruschke_rejecting_2018 and @kruschke_bayesian_2011.

###### Parameters {.unnumbered}

**Delta:** The expected population effect size.\
**TPR:** The desired long run probability of having the HDI fully contained within the ROPE interval, given Delta.\
**EqBand:** The chosen ROPE interval.\

###### How to use the package {.unnumbered}

```{r, eval = FALSE, tidy = TRUE, echo = TRUE}
SampleSizePlanner::ssp_bf_predetermined(
  tpr = 0.8, delta = 0.5, thresh = 10, max_n = 5000
  )
```

###### How to report your sample size estimation {.unnumbered}

In order to estimate the sample size, we used the Region of Practical Equivalence [@kruschke_rejecting_2018] method. We set the aimed TPR at 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement]. We consider all effect sizes below 0.2 equivalent to zero, because [1) previous studies reported a similar region of practical equivalence; 2) of the following substantive reasons: ...]. The expected delta was 0 because [1) we expected no difference between the groups]. Based on these parameters, a minimal sample size of 517 per group was estimated in order to reach a 0.8 TPR for our design.

# Summary {.unnumbered}

Justifying the decisions made during the sample size planning process presents valuable information when one evaluates the inferences drawn from a study. The Shiny app and R package presented in this paper aim to help researchers to choose and employ their sample size estimation method. In addition, the tool provides assistance in reporting the process and justification behind sample size choices. We encourage users and experts of the field to provide feedback and recommendations towards further developments.

# Authors contribution {.unnumbered}

**Conceptualization:** Marton Kovacs, Don van Ravenzwaaij, Rink Hoekstra, and Balazs Aczel.\
**Methodology:** Don van Ravenzwaaij.\
**Project Administration:** Marton Kovacs.\
**Software:** Marton Kovacs and Don van Ravenzwaaij.\
**Supervision:** Balazs Aczel.\
**Writing - Original Draft Preparation:** Marton Kovacs, Don van Ravenzwaaij, Rink Hoekstra, and Balazs Aczel.\
**Writing - Review & Editing:** Marton Kovacs, Don van Ravenzwaaij, Rink Hoekstra, and Balazs Aczel.

# Notes {.unnumbered}

## Glossary {.unnumbered}

*Accuracy in Parameter Estimation (AIPE):* A sample size estimation method used for parameter estimation. The approach aims to find the required sample size, such that the confidence interval has a certain expected width.\
*Priori Procedure (APP):* The approach aims to plan a sample size based on how close the researcher wishes both sample means to be to their respective population parameter, and how confident the researcher wants to be in this.\
*Bayesian inference:* A general framework for updating one's prior beliefs in light of new data.\
*Bayes Factor Design Analysis (BFDA):* This technique provides an expected sample size such that compelling evidence in the form of a Bayes factor can be collected for a given effect size with a certain long-run probability when allowing for sequential testing.\
*Testing vs. Estimation:* Two schools of inference, focusing on establishing whether or not an effect exists versus establishing the magnitude of an effect, respectively.\
*Equivalence band (EqBand):* The region of effect sizes considered practically equivalent to zero. In our paper, SESOI and ROPE are subsumed under EqBand.\
*Frequentist inference:* A general framework in which probabilities are defined as frequencies in hypothetical repeated events. In context of statistical testing, frequentist inference is concerned with long-run error rates of rejecting the null hypothesis for the observed or more extreme parameters in a given design when the model assumptions (e.g., independence of observations) are true.\
*Statistical power:* The long-run probability of finding a significant effect given a certain population effect size.\
*True positive rate (TPR):* The long-run probability of finding evidence for an effect, given that it exists. In our paper, statistical power is subsumed under TPR.\
*Classical power analysis:* This method is used to estimate the minimum sample size that a design needs to reach a certain level of statistical power, given a desired significance level and expected effect size.\
*Power-curve:* This curve shows how changes in effect size modify the statistical power of a test.\
*Region Of Practical Equivalence (ROPE):* The region of effect sizes considered practically equivalent to zero under the HDI-ROPE method.\
*Smallest Effect Size Of Interest (SESOI):* The region of effect sizes considered practically equivalent to zero under the TOST method.\
*Sequential testing:* The practice of incrementally testing as data comes in, typically until some pre-determined level of evidence is obtained.\
*Two One‐Sided Tests (TOST):* A frequentist statistical testing approach aimed at establishing equivalence between two groups.\
*Equivalence interval BF:* A Bayesian statistical testing approach aimed at establishing equivalence between two groups.\

\newpage

# References {.unnumbered}

```{r create_r-references, message = FALSE, warning = FALSE}
r_refs(file = "ssp.bib")
```

```{=tex}
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
```
::: {#refs custom-style="Bibliography"}
:::

\endgroup

### 2021-07-01 R/cdf_normal.R

@@ -1,28 +1,0 @@
-cdf_normal <- function(x,
-                       t,
-                       n1,
-                       n2 = NULL,
-                       independentSamples = FALSE,
-                       prior.mean,
-                       prior.variance,
-                       rel.tol = .Machine$double.eps^0.25) {
-  
-  out <- integrate(posterior_normal, lower = -Inf, upper = x,
-                   t = t, n1 = n1, n2 = n2,
-                   independentSamples = independentSamples,
-                   prior.mean = prior.mean,
-                   prior.variance = prior.variance,
-                   rel.tol = rel.tol)$value
-  
-  # catch numerical errors
-  if (out > 1 & out < 1.001) {
-    out <- 1
-    warning(
-      "Numerical integration yields a CDF value slightly larger than 1. The CDF value has been replaced by 1.",
-      call. = FALSE
-    )
-  }
-  
-  return(out)
-  
-}

### 2021-07-01 R/mod_ssp_tost.R

@@ -16,7 +16,7 @@
 mod_ssp_tost_ui <- function(id) {
   tagList(
     # Method
-    h1("Two One‐Sided Tests (TOST)", class = "method-title"),
+    h1("Two One-Sided Tests (TOST)", class = "method-title"),
     sidebarLayout(
       sidebarPanel(
         # Panel title

### 2021-07-01 R/plot_power_curve.R

@@ -1,9 +1,9 @@
 #' Function to plot the results of the power curve calculation
 #' 
 #' The function plots the power curve based on the results
-#' of the \code{\link{ssp_power_curve}} function.
+#' of the \code{\link{ssp_power_curve}} function or the \code{\link{ssp_bf_predetermined}} function.
 #' 
-#' @param delta Numeric. The range of sample size provided for the \code{\link{ssp_power_curve}} function.
+#' @param delta Numeric. The range of effect sizes.
 #' @param n1 Numeric. The determined sample sizes for each `delta`.
 #' @param animated Logical. if TRUE the output plot is animated.
 #' 
@@ -12,7 +12,10 @@
 #' @examples 
 #' \dontrun{
 #' # Determine the sample sizes for each delta
-#' curve_data <- SampleSizePlanner::ssp_power_curve(tpr = 0.8, delta = seq(0.1, 0.9, 0.01), max_n = 5000)
+#' curve_data <- SampleSizePlanner::ssp_power_curve(
+#' tpr = 0.8,
+#' delta = seq(0.1, 0.9, 0.01),
+#' max_n = 5000)
 #' # Plot the power curve
 #' SampleSizePlanner::plot_power_curve(delta = curve_data$delta, n1 = curve_data$n1, animated = FALSE)
 #' }

### 2021-07-01 R/posterior_t.R

@@ -1,36 +1,0 @@
-posterior_t <- function(delta,
-                        t,
-                        n1,
-                        n2 = NULL,
-                        independentSamples = FALSE,
-                        prior.location,
-                        prior.scale,
-                        prior.df,
-                        rel.tol = .Machine$double.eps^0.25) {
-  
-  neff <- ifelse(independentSamples, n1 * n2 / (n1 + n2), n1)
-  nu <- ifelse(independentSamples, n1 + n2 - 2, n1 - 1)
-  
-  mu.delta <- prior.location
-  gamma <- prior.scale
-  kappa <- prior.df
-  
-  numerator <- suppressWarnings(
-    dt(x = t, df = nu, ncp = sqrt(neff) * delta) *
-      1 / gamma * dt( (delta - mu.delta) / gamma, df = kappa)
-  )
-  
-  denominator <- integrate(integrand_t,
-                           lower = -Inf, upper = Inf,
-                           t = t, n = neff, nu = nu,
-                           mu.delta = mu.delta,
-                           gamma = gamma,
-                           kappa = kappa,
-                           rel.tol = rel.tol)$value
-  
-  out <- numerator / denominator
-  out[is.na(out)] <- 0
-  
-  return(out)
-  
-}

### 2021-07-01 R/ssp_app.R

@@ -11,11 +11,14 @@
 #' @return The function returns a list of one named element. The determined
 #' sample size for group one and two called `n1`.
 #' @export
+#' 
+#' @importFrom stats qnorm
+#' 
 #' @examples
 #' \dontrun{
 #' SampleSizePlanner::ssp_app(closeness = 0.2, confidence = 0.95)
 #' }
 ssp_app <- function(confidence, closeness) {
-  n1 <- ceiling((qnorm((sqrt(confidence) + 1) / 2) / closeness)^2)
+  n1 <- ceiling((stats::qnorm((sqrt(confidence) + 1) / 2) / closeness)^2)
   list(n1 = n1)
 }

### 2021-07-01 R/ssp_bf_predetermined.R

@@ -1,15 +1,20 @@
 #' Determine sample size with Predetermined sample size with Bayes factor method
 #' 
 #' The present method calculates the corresponding default Bayes factor for
-#' a t-test statistic with Cauchy prior distribution centered on zero with
-#' scale parameter 1/sqrt(2) for several sample sizes.
+#' a t-test statistic with Cauchy prior distribution for several sample sizes.
+#' The function returns the optimal sample size needed to reach the TPR for a given
+#' Bayes factor threshold to detect an expected population effect size. If a range of
+#' possible population effect sizes are plausible under the given hypothesis,
+#' the function can calculate the optimal sample sizes for the given range of
+#' effect sizes and present the results in a figure by using the \code{\link{plot_power_curve}} function.
 #' 
 #' @param tpr Numeric. The long-run probability of obtaining a Bayes factor at least as high as the critical threshold favoring superiority, given Delta.
 #' @param delta Numeric. The expected population effect size or a range of expected effect sizes.
 #' @param max_n Integer. The maximum number of participants per group (both groups are assumed to have equal sample size).
 #' @param thresh Integer. Critical threshold for the Bayes factor.
-#' @param tol Numeric. TODO
-#' @param granularity Numeric. TODO
+#' @param tol Numeric. Relative accuracy requested.
+#' @param granularity Numeric. Relative precision of the tpr estimates, higher values mean more precision.
+#' @param prior_scale Numeric. Scale of the Cauchy prior distribution.
 #' 
 #' @return The function returns a list of three named numeric vectors.
 #' The first `delta` is the range of deltas provided for the function.
@@ -21,12 +26,19 @@
 #' \dontrun{
 #' SampleSizePlanner::ssp_bf_predetermined(tpr = 0.8, delta = 0.5, thresh = 10, max_n = 5000)
 #' }
-ssp_bf_predetermined <- function(tpr, delta, thresh = 10, max_n = 5000, tol = 1e-4, granularity = 300) {
[…]

### 2021-07-01 R/ssp_power_curve.R

@@ -5,7 +5,8 @@
 #' classical power analysis but instead of calculating the
 #' appropriate sample size for one hypothesized population effect
 #' size, the method calculates the required sample size for a range
-#' of plausible population effect sizes.
+#' of plausible population effect sizes. To plot the results use the
+#' \code{\link{plot_power_curve}} function.
 #' 
 #' @param delta Numeric. A range of hypothetical population effect sizes.
 #' @param tpr Numeric. The desired long-run probabilities of obtaining a significant result with a one-sided t-test, given each value of Delta.

### 2021-07-01 R/ssp_tost.R

@@ -8,24 +8,27 @@
 #' @param tpr Numeric. The desired long run probability of obtaining a significant result with TOST, given Delta.
 #' @param eq_band Numeric. The chosen width of the region for practical equivalence, i.e. the SESOI.
 #' @param delta Numeric. The expected population effect size. In most cases, this value will be zero.
-#' @param sigma Numeric. TODO
-#' @param nr Numeric. TODO
 #' @param alpha Numeric. The level of significance.
 #' 
 #' @return The function returns a list of three named numeric vectors. The
 #'   sample size for group 1 `n1`, the sample size for group 2 `n2` and
 #'   the associated power `npower`.
 #' @export
+#' 
+#' @importFrom stats dchisq pnorm
+#' 
 #' @examples 
 #' \dontrun{
 #' SampleSizePlanner::ssp_tost(tpr = 0.8, eq_band = 0.2, delta = 0)
 #' }
-ssp_tost <- function(tpr, eq_band, delta, sigma = 1, nr = 1, alpha = .05) {
+ssp_tost <- function(tpr, eq_band, delta, alpha = .05) {
   # Validation of function arguments
   assertthat::assert_that(is_positive_number(tpr))
   assertthat::assert_that(is_positive_number(eq_band))
   assertthat::assert_that(is_not_null(delta))
   
+  nr = 1
+  sigma = 1
   max_n = 10001
   n1 = 4
   sigsq = sigma^2
@@ -47,9 +50,9 @@ ssp_tost <- function(tpr, eq_band, delta, sigma = 1, nr = 1, alpha = .05) {
     int = cu - cl
     intl = int / numint
     cvec = cl + intl * (0:numint)
-    wcpdf = (intl / 3) * coevecc * dchisq(cvec, df)
+    wcpdf = (intl / 3) * coevecc * stats::dchisq(cvec, df)
     st = sqrt(cvec / df) * tcrit
-    npower = sum(wcpdf * (pnorm((eq_band - delta) / std - st) - pnorm((-eq_band - delta) / std + st)))
+    npower = sum(wcpdf * (stats::pnorm((eq_band - delta) / std - st) - stats::pnorm((-eq_band - delta) / std + st)))
   }
   
   if (npower == 0) {

### 2021-07-01 dev/02_dev.R

@@ -35,9 +35,15 @@ usethis::use_package("shinyBS")
 usethis::use_package("BayesFactor")
 usethis::use_package("assertthat")
 usethis::use_package("MBESS")
-usethis::use_package("rclipboard")
+usethis::use_package("clipr")
 usethis::use_package("papaja")
 usethis::use_package("formatR")
+usethis::use_package("config")
+usethis::use_package("dplyr")
+usethis::use_package("rlang")
+usethis::use_package("purrr")
+usethis::use_package("rclipboard")
+usethis::use_package("stats")
 
 ## 2.3 Add tests
 

### 2021-07-16 inst/manuscript/manuscript.Rmd
updated manuscript according to the first round of revision

@@
 title             : "SampleSizePlanner: A Tool to Estimate and Justify Sample Size for Two-Group Studies"
-shorttitle        : "SAMPLESIZEPLANNER: A TOOL TO ESTIMATE AND JUSTIFY SAMPLE SIZE"
+shorttitle        : "SAMPLESIZEPLANNER"
 
@@
 figsintext        : yes
-numbersections    : true
-
-params:
-  highlight_changes: TRUE
+numbersections    : false
 
@@
 classoption       : "man"
-output            : papaja::apa6_docx
+output            : papaja::apa6_pdf
 ---
@@
 
-# Introduction {.unnumbered}
+# Introduction
 
@@
 The aim of our tutorial is to point out for each method the essential decision points that a researcher has to face during this process.
+We provide a short description of each method and the corresponding parameters, but we avoid listing their advantages and disadvantages.
+As there are disagreements between the experts of the field regarding the correct use of some of the methods, we intentionally try to remain impartial and do not favor any of the presented methods.
+Researchers who want to know more about each method can find a number of useful references in the description of the methods.
 We also provide a collection of ready-to-use analysis code and a ShinyApp that helps researchers use and report the main sample size estimation techniques for different scenarios.
@@
 
-# Sample Size Determination and Justification {.unnumbered}
+# Sample Size Determination and Justification
 
@@
 
-## Choosing a method in case of sample size justification {.unnumbered}
+## Choosing a method in case of sample size justification
 
@@
 Within the realm of statistical testing, there are some other factors that affect the preferred inference tool: Do you prefer to test for equivalence (no difference in mean) or for superiority (mean of one group larger than mean of other group), are you interested in calculating a required sample size for a specific hypothetical effect size or for a range of possible values, and do you wish to employ sequential testing (applicable to Bayesian testing)?
+In case of testing, some of the methods are designed to find support for the null hypothesis (e.g., TOST, ROPE), while others are designed to find support for the alternative hypothesis (e.g., traditional null hypothesis testing), and some methods are designed to find support for either (e.g., BFDA).
 For frequentist estimation, the preferred inference tool might differ depending on whether we evaluate uncertainty for each group separately or jointly.
@@
 
-## How to use this guide {.unnumbered}
+## How to use this guide
 
@@
 We offer a list of possible justifications at the decision points for each method (indicated between square brackets), but users are able to provide their own justification as free-text.
+It is important to note that the listed justifications are meant to provide guidance for the user, and they are not sufficient without further details provided by the researcher in the context of the given study.
+For example, previously reported values should always be accompanied by a theoretical justification of why these values make sense.
 The provided justification text could serve as a stub for the description of the chosen sample size in a paper, a preregistration or registered report, or a grant proposal.
@@
 For illustrative purposes, we will have Mary's hypothetical study goal depend on the earlier presented decisions to illustrate different scenarios for each sample size planning method.
+That means that in some of the upcoming examples, Mary expects a null effect and in others she expects a positive effect.
+We will also present a justification text based on Mary’s choices for each method.
 
@@
 
-## 1. Testing {.unnumbered}
+## 1. Testing
 
-### 1.1. Effect size = 0 {.unnumbered}
+### 1.1. Effect size = 0
 
-#### 1.1.1. Two One‐Sided Tests (TOST) {.unnumbered}
+#### 1.1.1. Two One‐Sided Tests (TOST)
 
-##### Study context {.unnumbered}
+##### Study context
 
@@
 
-##### Description {.unnumbered}
+##### Description
 
@@
 In case of TOST, the two null hypotheses state that the effect size is equal to the lower and upper equivalence band values, whereas the alternative hypotheses state that the effect size is significantly smaller than the upper equivalence band value and significantly larger than the lower equivalence band value.
-In case both one-sided tests reject the null-hypotheses at a given significance level, the group means are considered to be practically equivalent.
+In case both one-sided tests reject the null-hypothesis at a given significance level, the group means are considered to be practically equivalent.
 See @lakens_equivalence_2018, for further reading.
 
-##### Parameters {.unnumbered}
+##### Parameters
 
@@
 
-##### How to use the package {.unnumbered}
+##### How to use the package
 
@@
 
-##### How to report your sample size estimation {.unnumbered}
+##### How to report your sample size estimation
 
 In order to calculate an appropriate sample size for testing whether the two groups are practically equivalent, we used the Two One-Sided Tests of Equivalence [TOST; @schuirmann_comparison_1987] method.
+We used an alpha of 0.05.
 We set the aimed TPR to be 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement].
-We consider all effect sizes below 0.2 equivalent to zero, because [1) previous studies reported a similar equivalence band; 2) of the following substantive reasons: ...].
+We consider all effect sizes below 0.2 equivalent to zero, because [1) previous studies reported the choice of a similar equivalence band; 2) of the following substantive reasons: ...].
 The expected delta was 0 because [1) we expected no difference between the groups].
@@
 
-#### 1.1.2. Equivalence interval Bayes factor {.unnumbered}
+#### 1.1.2. Equivalence interval Bayes factor
 
-##### Study context {.unnumbered}
+##### Study context
 
@@
 The resulting value quantifies how much more likely it is that the data occurred under a population effect size deemed 'equivalent' relative to the data having occurred under a population effect size deemed non-equivalent.
-The current implementation uses a default Cauchy prior on effect size with scale parameter 1/$\sqrt{2}$.
-For further reading, see @morey_bayes_2011 and @van_ravenzwaaij_bayes_2019.
+The current implementation uses a default Cauchy prior on effect size with the possible scale parameters of medium (r = 1/$\sqrt{2}$), wide (r = 1), or ultra-wide (r = $\sqrt{2}$).
+For further reading, see @morey_bayes_2011, @van_ravenzwaaij_bayes_2019, and @linde_tendeiro_selker_wagenmakers_van_ravenzwaaij_2020.
 
-##### Parameters {.unnumbered}
+##### Parameters
 
@@
 **EqBand:** The chosen width of the equivalence region.\
+**PriorScale:** The scale of the Cauchy prior distribution. The PriorScale in the application can be set to: 1/$\sqrt{2}$, 1, and $\sqrt{2}$.\
 *Threshold:* Critical threshold for the Bayes factor. The threshold level in the application can be set to 10, 6, or 3.\
 
-##### How to use the package {.unnumbered}
+##### How to use the package
 
@@
   tpr = 0.8, delta = 0,
-  eq_band = 0.2, thresh = 10
+  eq_band = 0.2, thresh = 10, prior_scale = 1/sqrt(2)
   )
@@
 
-##### How to report your sample size estimation {.unnumbered}
+##### How to report your sample size estimation
 
 In order to estimate the sample size, we used the interval equivalent Bayes factor [@morey_bayes_2011; @van_ravenzwaaij_bayes_2019] method.
+We used a Cauchy prior distribution centered on zero with a scale of 1/$\sqrt{2}$.
 We set the aimed TPR at 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement].
-We consider all effect sizes below 0.2 equivalent to zero, because [1) previous studies reported a similar equivalence region; 2) of the following substantive reasons: ...].
+We consider all effect sizes below 0.2 equivalent to zero, because [1) previous studies reported the choice of a similar equivalence region; 2) of the following substantive reasons: ...].
 The expected delta was 0 because [1) we expected no difference between the groups].
@@
 
-### 1.2. Effect size \>0 {.unnumbered}
+### 1.2. Effect size \>0
 
-#### 1.2.1. Frequentist {.unnumbered}
+#### 1.2.1. Frequentist
 
-##### 1.2.1.1. Classical power analysis {.unnumbered}
+##### 1.2.1.1. Classical power analysis
 
-###### Study context {.unnumbered}
+###### Study context
 
@@
 
-###### Description {.unnumbered}
+###### Description
 
-The classical power analysis approach allows to calculate the required sample size in order to obtain a significant result for the null hypothesis test a certain proportion of times in the long run given an assumed population effect size.
+The classical power analysis approach allows one to calculate the required sample size in order to obtain a significant result for the null hypothesis test a certain proportion of times in the long run given an assumed population effect size.
 
-###### Parameters {.unnumbered}
+###### Parameters
 
@@
 
-###### How to use the package {.unnumbered}
+###### How to use the package
 
@@
 
-###### How to report your sample size estimation {.unnumbered}
+###### How to report your sample size estimation
 
 We used a power analysis to estimate the sample size.
+We used an alpha of 0.05.
 We set the aimed TPR at 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement].
@@
 
-##### 1.2.1.2. Power curve {.unnumbered}
+##### 1.2.1.2. Power curve
 
-###### Study context {.unnumbered}
+###### Study context
 
@@
 
-###### Description {.unnumbered}
+###### Description
 
@@
 
-###### Parameters {.unnumbered}
+###### Parameters
 
@@
 
-#### 1.2.2. Bayesian {.unnumbered}
+#### 1.2.2. Bayesian
 
-##### 1.2.2.1. Predetermined sample size with Bayes factor {.unnumbered}
+##### 1.2.2.1. Predetermined sample size with Bayes factor
 
-###### Study context {.unnumbered}
+###### Study context
 
@@
 
-###### Description {.unnumbered}
+###### Description
 
-The present method calculates the corresponding default Bayes factor for a *t*-test statistic with Cauchy prior distribution centered on zero with scale parameter 1/$\sqrt{2}$ for several sample sizes [the so-called Jeffrey-Zellner-Siow Bayes factor, see e.g., @rouder_bayesian_2009].
+The present method calculates the corresponding default Bayes factor for a *t*-test statistic with Cauchy prior distribution centered on zero with scale parameter of either 1/$\sqrt{2}$, 1, or $\sqrt{2}$ for several sample sizes [the so-called Jeffrey-Zellner-Siow Bayes factor, see e.g., @rouder_bayesian_2009].
 The function returns the optimal sample size needed to reach the TPR for a given Bayes factor threshold to detect an expected population effect size.
 If a range of possible population effect sizes are plausible under the given hypothesis, the function can calculate the optimal sample sizes for the given range of effect sizes and present the results in a figure (analogous to the Power Curve method).
+This method is designed to determine the sample sizes for the existence of an effect (i.e., Delta > 0).
 
-###### Parameters {.unnumbered}
+###### Parameters
 
@@
 **Maximum N:** The maximum number of participants per group (both groups are assumed to have equal sample size).\
+**PriorScale:** The scale of the Cauchy prior distribution. The PriorScale in the application can be set to: 1/$\sqrt{2}$, 1, and $\sqrt{2}$.\
 **Threshold:** Critical threshold for the Bayes factor. Three threshold levels are available in the app: 3, 6, and 10.\
 
-###### How to use the package {.unnumbered}
+###### How to use the package
 
@@
 SampleSizePlanner::ssp_bf_predetermined(
-  tpr = 0.8, delta = 0.5, thresh = 10, max_n = 5000
+  tpr = 0.8, delta = 0.5, thresh = 10, max_n = 5000, prior_scale = 1/sqrt(2)
   )
@@
 
-###### How to report your sample size estimation {.unnumbered}
+###### How to report your sample size estimation
 
 We used the Jeffrey-Zellner-Siow Bayes factor method to estimate the sample size.
+We used a Cauchy prior distribution centered on zero with a scale of 1/$\sqrt{2}$.
 We set the aimed TPR at 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement].
@@
 
-##### 1.2.2.2. Bayes Factor Design Analysis (BFDA) {.unnumbered}
+##### 1.2.2.2. Bayes Factor Design Analysis (BFDA)
 
-###### Study context {.unnumbered}
+###### Study context
 
@@
 In the app, H~0~ and H~a~ indicate the proportion of times sequential testing leads to Bayes factors providing evidence with the given threshold for the null hypothesis and for the alternative hypothesis, respectively.
+Users of the Shiny app and R package should set Delta to 0 if they wish to determine the sufficient sample size for rejecting an effect, and use Delta > 0 if they wish to find support for the existence of an effect.
 For further reading, see @schonbrodt_bayes_2018 and @schonbrodt_sequential_2017.
 
-###### Parameters {.unnumbered}
+###### Parameters
 
@@
 **TPR:** The long run probability of obtaining a Bayes factor at least as high as the critical threshold favoring superiority, given Delta.\
+**PriorScale:** The scale of the Cauchy prior distribution. The PriorScale in the application can be set to: 1/$\sqrt{2}$, 1, and $\sqrt{2}$.\
 **Threshold:** Critical threshold for the Bayes factor. Three threshold levels are available in the app: 3, 6, and 10.\
 
-###### How to use the package {.unnumbered}
+###### How to use the package
 
@@
 SampleSizePlanner::ssp_bfda(
-  tpr = 0.8, delta = 0.5, thresh = 10, n_rep = 1000
+  tpr = 0.8, delta = 0.5, thresh = 10, n_rep = 10000, prior_scale = 1/sqrt(2)
   )
@@
 
-###### How to report your sample size estimation {.unnumbered}
+###### How to report your sample size estimation
 
 We used the BFDA method to estimate the sample size.
+We used a Cauchy prior distribution centered on zero with a scale of 1/$\sqrt{2}$.
 We set the aimed TPR at 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement].
@@
 
-## 2. Estimation {.unnumbered}
+## 2. Estimation
 
-### 2.1. Frequentist {.unnumbered}
+### 2.1. Frequentist
 
-#### 2.1.1. Accuracy In Parameter Estimation (AIPE) {.unnumbered}
+#### 2.1.1. Accuracy In Parameter Estimation (AIPE)
 
-##### Study context {.unnumbered}
+##### Study context
 
@@
 
-#### Parameters {.unnumbered}
+#### Parameters
 
@@
 
-##### How to report your sample size estimation {.unnumbered}
+##### How to report your sample size estimation
 
@@
 We aimed for a 95% confidence level, because [1) it is the common standard in the field; 2) it is the journal publishing requirement].
-The desired width was 0.4 because [1) previous studies on this topic reported a similar region of practical equivalence; 2) of the following substantive reasons: ...].
+The desired width was 0.4 because [1) previous studies reported the choice of a similar region of practical equivalence; 2) of the following substantive reasons: ...].
 We expected an underlying population effect size of 0.3, because [1) previous results published in ...; 2) of the following substantive reasons: ...].
@@
 
-#### 2.1.2. A Priori Precision (APP) {.unnumbered}
+#### 2.1.2. A Priori Precision (APP)
 
-###### Study context {.unnumbered}
+###### Study context
 
@@
 
-##### Parameters {.unnumbered}
+##### Parameters
 
@@
 
-###### How to report your sample size estimation {.unnumbered}
+###### How to report your sample size estimation
 
@@
 
-### 2.2. Bayesian testimation {.unnumbered}
+### 2.2. Bayesian testimation
 
-#### 2.2.1. Region of Practical Equivalence (ROPE) {.unnumbered}
+#### 2.2.1. Region of Practical Equivalence (ROPE)
 
-###### Study context {.unnumbered}
+###### Study context
 
@@
 **EqBand:** The chosen ROPE interval.\
+**PriorScale:** The scale of the Cauchy prior distribution. The PriorScale in the application can be set to: 1/$\sqrt{2}$, 1, and $\sqrt{2}$.\
 
-###### How to use the package {.unnumbered}
+###### How to use the package
 
 ```{r, eval = FALSE, tidy = TRUE, echo = TRUE}
-SampleSizePlanner::ssp_bf_predetermined(
-  tpr = 0.8, delta = 0.5, thresh = 10, max_n = 5000
+SampleSizePlanner::ssp_rope(
+  tpr = 0.8, delta = 0.5, eq_band = 0.2, prior_scale = 1/sqrt(2)
   )
@@
 
-###### How to report your sample size estimation {.unnumbered}
+###### How to report your sample size estimation
 
 In order to estimate the sample size, we used the Region of Practical Equivalence [@kruschke_rejecting_2018] method.
+We used a Cauchy prior distribution centered on zero with a scale of 1/$\sqrt{2}$.
 We set the aimed TPR at 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement].
-We consider all effect sizes below 0.2 equivalent to zero, because [1) previous studies reported a similar region of practical equivalence; 2) of the following substantive reasons: ...].
+We consider all effect sizes below 0.2 equivalent to zero, because [1) previous studies reported the choice of a similar region of practical equivalence; 2) of the following substantive reasons: ...].
 The expected delta was 0 because [1) we expected no difference between the groups].
@@
 
-# Summary {.unnumbered}
+# Summary
 
@@
 
-# Authors contribution {.unnumbered}
+# Authors contribution
 
@@
 
-# Notes {.unnumbered}
+# Notes
 
-## Glossary {.unnumbered}
+## Glossary
 
@@
 *Frequentist inference:* A general framework in which probabilities are defined as frequencies in hypothetical repeated events.
-In context of statistical testing, frequentist inference is concerned with long-run error rates of rejecting the null hypothesis for the observed or more extreme parameters in a given design when the model assumptions (e.g., independence of observations) are true.\
+In the context of statistical testing, frequentist inference is concerned with long-run error rates of rejecting the null hypothesis for the observed or more extreme parameters in a given design when the model assumptions (e.g., independence of observations) are true.\
 *Statistical power:* The long-run probability of finding a significant effect given a certain population effect size.\
@@
 
-# References {.unnumbered}
+# References
 

### 2021-07-16 inst/manuscript/manuscript.tex
updated manuscript according to the first round of revision

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={SampleSizePlanner: A Tool to Estimate and Justify Sample Size for Two-Group Studies},
  pdfauthor={Marton Kovacs1,2, Don van Ravenzwaaij3, Rink Hoekstra3, \& Balazs Aczel1},
  pdflang={en-EN},
  pdfkeywords={sample size determination; power analysis; study design},
  colorlinks=true,
  linkcolor=blue,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{SAMPLESIZEPLANNER}
\keywords{sample size determination; power analysis; study design}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{SampleSizePlanner: A Tool to Estimate and Justify Sample Size for Two-Group Studies}
\author{Marton Kovacs\textsuperscript{1,2}, Don van Ravenzwaaij\textsuperscript{3}, Rink Hoekstra\textsuperscript{3}, \& Balazs Aczel\textsuperscript{1}}
\date{}

\authornote{

Marton Kovacs and Don van Ravenzwaaij are shared first authors.

Correspondence concerning this article should be addressed to Marton Kovacs, Hungary 1075 Budapest Kazinczy street 23-27. E-mail: \href{mailto:marton.balazs.kovacs@gmail.com}{\nolinkurl{marton.balazs.kovacs@gmail.com}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Institute of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary\\\textsuperscript{2} Doctoral School of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary\\\textsuperscript{3} University of Groningen, Groningen, The Netherlands}

\abstract{
Planning sample size often requires researchers to identify a statistical technique
and to make several choices during their calculations. Currently, there is a lack of clear guidelines
for researchers to find and use the applicable procedure. In the present tutorial,
we introduce a web app and R package that offer nine different procedures to determine and
justify the sample size for independent two-group study designs. The application highlights
the most important decision points for each procedure and suggests example justifications for them.
The resulting sample size report can serve as a template for preregistrations and manuscripts.
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Social and behavioral sciences are known to be plagued by undersampling (Ioannidis, 2005). In the traditional statistical framework, even when the effect exists, undersampled studies yield either nonsignificant results or significant results due to overestimating the size of the effect. Because nonsignificant results are less likely to reach publications than significant ones, results of undersampled studies either remain unpublished or impose a substantial bias on our body of published empirical findings. In addition, the low informational value of undersampled studies may not justify the cost or potential risk they induce (Halpern, Karlawish, \& Berlin, 2002). To mitigate these issues, authors are increasingly expected to plan and justify the sample size of their study (Maxwell, 2004). However, such sample size justifications are only meaningful if they provide sufficient information to the readers to judge the adequacy of the author's decisions.

In the statistical literature, a few methods have been proposed to determine and justify sample size. In practice, however, authors are short of practical guides on how to navigate among the different sample size methods. The aim of our tutorial is to point out for each method the essential decision points that a researcher has to face during this process. We provide a short description of each method and the corresponding parameters, but we avoid listing their advantages and disadvantages. As there are disagreements between the experts of the field regarding the correct use of some of the methods, we intentionally try to remain impartial and do not favor any of the presented methods. Researchers who want to know more about each method can find a number of useful references in the description of the methods. We also provide a collection of ready-to-use analysis code and a ShinyApp that helps researchers use and report the main sample size estimation techniques for different scenarios. The tutorial is focused exclusively on the scenario of the comparison of two independent groups (i.e., the independent \emph{t}-test design) with a one-sided test.

\hypertarget{sample-size-determination-and-justification}{%
\section{Sample Size Determination and Justification}\label{sample-size-determination-and-justification}}

A lot of factors go into the determination of the sample size for an independent two group study design. In this section, we will first provide a birds-eye view of the most important decisions. Next, we will go into more detail on the specific inference tool that results from the combination of the larger choices.

It is crucial to not just state how we determined our planned sample size but to also give the reader insight into the reasons behind our choices. In a recent overview, Lakens (2021) lists six types of general approaches to justify sample size in quantitative empirical studies: (1) Measure entire population; (2) Resource constraints; (3) A-priori power analysis; (4) Accuracy; (5) Heuristics; and (6) No justification. For the first approach, no quantitative justification is necessary; and for the second approach, the researcher has no freedom to increase the sample size. Power analysis, or more generally the estimation of true positive rate, is used when one plans to conduct hypothesis testing; accuracy justifications are used when one plans to conduct parameter estimation. Our tutorial mainly focuses on approaches two, three, and four, and is aimed at providing a hands-on approach for the mechanical part of the sample size determination (i.e., the calculation). For a deeper discussion of justification of these approaches, or for other approaches (i.e., using heuristics or not providing justification), we refer the reader to Lakens (2021).

\hypertarget{choosing-a-method-in-case-of-sample-size-justification}{%
\subsection{Choosing a method in case of sample size justification}\label{choosing-a-method-in-case-of-sample-size-justification}}

In an ideal world, the choice for the number of participants would be solely determined by scientific considerations, and depending on the chosen technique the collection of data would continue until either the desired sample size or a desired outcome has been reached. In practice, researchers are limited by time (collecting data is quite demanding), money (participants or people collecting the data may be paid, and the same may hold for renting space or equipment), or availability of participants (the population may be relatively small, and/or the participation rate quite low).

When constrained by limited resources, it is important to be transparent about those limitations. It is also important to be open about scientific considerations. Depending on the nature of the study (perhaps it is an initial exploration?), small sample sizes need not be a dealbreaker. So although more data are always preferred from an informational point of view, by owning the limitations of our study, we improve future readers' understanding of the process leading up to the eventual paper, and we also answer in advance to those who think the chosen sample size was insufficient.

Whether or not authors have limited resources, two important choices need to be made: (1) whether they are interested in \emph{statistical testing} or in \emph{parameter estimation}; and (2) whether they want to conduct their statistical inference within the \emph{frequentist} framework or within the \emph{Bayesian} framework. Starting with the first decision, statistical testing is the primary framework when one is interested in establishing whether an underlying population effect is equal to, different from, larger than, or smaller than a certain value. In essence, statistical testing lends itself to binary decision making. Typically, testing is concerned with a fixed point null hypothesis (e.g., there is no difference between two groups), although using intervals for testing is also possible. Alternatively, one might be interested in parameter estimation that is less interested in establishing the existence of a difference and instead is concerned with establishing the magnitude of the difference.

The second important decision concerns the statistical framework. Choosing to conduct statistical tests within a frequentist framework, one is usually interested in balancing the type I (false positive) and type II (false negative) error rates. Practitioners choosing to conduct statistical tests within a Bayesian framework are typically interested in being able to quantify the relative probability of hypotheses or models being true given the data and in including prior information.

Within the realm of statistical testing, there are some other factors that affect the preferred inference tool: Do you prefer to test for equivalence (no difference in mean) or for superiority (mean of one group larger than mean of other group), are you interested in calculating a required sample size for a specific hypothetical effect size or for a range of possible values, and do you wish to employ sequential testing (applicable to Bayesian testing)? In case of testing, some of the methods are designed to find support for the null hypothesis (e.g., TOST, ROPE), while others are designed to find support for the alternative hypothesis (e.g., traditional null hypothesis testing), and some methods are designed to find support for either (e.g., BFDA). For frequentist estimation, the preferred inference tool might differ depending on whether we evaluate uncertainty for each group separately or jointly. We will describe these specific factors when we go into detail about each of the preferred methods. A flow-chart representing all of these choices is given in Figure~\ref{fig:flowchart}.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./flowchart}

}

\caption{The figure depicts the decisions that one faces when choosing among sample size estimation methods. The nine sample size estimation methods discussed in this paper are listed in the bottom row. Some decisions are determined by the investigated question and the design of the study while others are based on the preferred statistical framework.}\label{fig:flowchart}
\end{figure}

\hypertarget{how-to-use-this-guide}{%
\subsection{How to use this guide}\label{how-to-use-this-guide}}

In the next section, we will illustrate the specific inference tools and resulting sample size calculations in more detail using a ShinyApp and an R package we have developed. Throughout this section, we recurrently use two terms that have different meanings for different techniques. These are the \emph{true positive rate} (TPR), and the \emph{equivalence band} (EqBand). The TPR reflects the long-run probability of concluding there is an effect, given that it does exist. For traditional null hypothesis testing, this is typically referred to as power, but related concepts exist for different inference tools. The EqBand refers to an effect size region, typically around zero, that is deemed clinically insignificant or irrelevant. Different names are given to this region depending on the technique that employs them, such as statistical effect size of interest (SESOI) or region of practical equivalence (ROPE). For both TPR and EqBand, we explain the specific meaning in context of the relevant inference tool below.

For each method, only the main parameters can be adjusted with a certain range of values in the ShinyApp by using a slider. These parameters are presented in the text in bold. Other parameters are set to preset values in the application but can be adjusted in the accompanying R package to any sensible value. These parameters are highlighted in italics in the tutorial. Both the app and the package allow the users to save or copy a text template with the results of the sample size determination. We offer a list of possible justifications at the decision points for each method (indicated between square brackets), but users are able to provide their own justification as free-text. It is important to note that the listed justifications are meant to provide guidance for the user, and they are not sufficient without further details provided by the researcher in the context of the given study. For example, previously reported values should always be accompanied by a theoretical justification of why these values make sense. The provided justification text could serve as a stub for the description of the chosen sample size in a paper, a preregistration or registered report, or a grant proposal.

Throughout, we will use the example story of Mary the educational psychologist. Mary has come up with a new set of games that challenge spatial insight. She believes that distributed and targeted engagement with these games for a period of six months for children in the age range of 8 to 12 will lead to lasting improvements on their IQ score as measured through Raven's progressive matrices test (population mean 100, population SD 15). Mary collects data for a control sample that gets regular education and for an experimental sample and plans to compare those samples. For illustrative purposes, we will have Mary's hypothetical study goal depend on the earlier presented decisions to illustrate different scenarios for each sample size planning method. That means that in some of the upcoming examples, Mary expects a null effect and in others she expects a positive effect. We will also present a justification text based on Mary's choices for each method.

The ShinyApp is available on \url{https://martonbalazskovacs.shinyapps.io/SampleSizePlanner} and the R package can be installed by running the following command in R \texttt{devtools::install\_github("marton-balazs-kovacs/SampleSizePlanner")}. There is more information about the R package and the ShinyApp on the projects' Github page \url{https://github.com/marton-balazs-kovacs/SampleSizePlanner}, or on the website \url{https://marton-balazs-kovacs.github.io/SampleSizePlanner/}.

\hypertarget{testing}{%
\subsection{1. Testing}\label{testing}}

\hypertarget{effect-size-0}{%
\subsubsection{1.1. Effect size = 0}\label{effect-size-0}}

\hypertarget{two-onesided-tests-tost}{%
\paragraph{1.1.1. Two One‐Sided Tests (TOST)}\label{two-onesided-tests-tost}}

\hypertarget{study-context}{%
\subparagraph{Study context}\label{study-context}}

Mary would like to know what sample size she needs for a power of .80 to study whether the mean IQ score of the experimental group's population is practically equivalent to the mean IQ score of the control group. She tests this assumption in a frequentist framework, and considers a population effect size between -0.2 and 0.2 to be `practically equivalent' to no difference. This would correspond to IQ scores between 97 (100+15*-.2) and 103 (100+15*.2).

\hypertarget{description}{%
\subparagraph{Description}\label{description}}

TOST is a frequentist equivalence testing approach that adopts two one-sided hypotheses to designate an interval hypothesis (Schuirmann, 1987). The lower and upper boundaries of the interval are determined by the equivalence band (i.e.~SESOI) around the expected population effect size (e.g., 0). Lakens, Scheel, and Isager (2018) lists several methods that can be used to determine the SESOI . In case of TOST, the two null hypotheses state that the effect size is equal to the lower and upper equivalence band values, whereas the alternative hypotheses state that the effect size is significantly smaller than the upper equivalence band value and significantly larger than the lower equivalence band value. In case both one-sided tests reject the null-hypothesis at a given significance level, the group means are considered to be practically equivalent. See Lakens, Scheel, and Isager (2018), for further reading.

\hypertarget{parameters}{%
\subparagraph{Parameters}\label{parameters}}

\textbf{Delta:} The expected population effect size. In most cases, this value will be zero.\\
\textbf{TPR:} The desired long run probability of obtaining a significant result with TOST, given Delta.\\
\textbf{EqBand:} The chosen width of the region for practical equivalence, i.e.~the SESOI.\\
\emph{Alpha:} The level of significance. The alpha level in the application is preset to 0.05.\\

\hypertarget{how-to-use-the-package}{%
\subparagraph{How to use the package}\label{how-to-use-the-package}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{SampleSizePlanner}\SpecialCharTok{::}\FunctionTok{ssp\_tost}\NormalTok{(}\AttributeTok{tpr =} \FloatTok{0.8}\NormalTok{, }\AttributeTok{eq\_band =} \FloatTok{0.2}\NormalTok{, }\AttributeTok{delta =} \DecValTok{0}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{how-to-report-your-sample-size-estimation}{%
\subparagraph{How to report your sample size estimation}\label{how-to-report-your-sample-size-estimation}}

In order to calculate an appropriate sample size for testing whether the two groups are practically equivalent, we used the Two One-Sided Tests of Equivalence {[}TOST; Schuirmann (1987){]} method. We used an alpha of 0.05. We set the aimed TPR to be 0.8, because {[}1) it is the common standard in the field; 2) it is the journal publishing requirement{]}. We consider all effect sizes below 0.2 equivalent to zero, because {[}1) previous studies reported the choice of a similar equivalence band; 2) of the following substantive reasons: \ldots{]}. The expected delta was 0 because {[}1) we expected no difference between the groups{]}. Based on these parameters, a sample size of 429 per group was estimated in order to reach a TPR of 0.8 with our design.

\hypertarget{equivalence-interval-bayes-factor}{%
\paragraph{1.1.2. Equivalence interval Bayes factor}\label{equivalence-interval-bayes-factor}}

\hypertarget{study-context-1}{%
\subparagraph{Study context}\label{study-context-1}}

Mary would like to know what sample size she needs to have a long-run probability of .80 of obtaining a Bayes factor larger than 10. Mary would like to test whether the mean IQ score of the experimental group's population is practically equivalent to the mean IQ score of the control group. Mary hypothesizes that there is no difference (i.e., H\textsubscript{0} is true). Mary tests this assumption in a Bayesian framework. Mary considers a population effect size between -0.2 and under 0.2 to be `practically equivalent.' This would correspond to IQ scores between 97 (100+15*-.2) and 103 (100+15*.2).

\hypertarget{description-1}{%
\subparagraph{Description}\label{description-1}}

Equivalence interval Bayes factors contrast an equivalence hypothesis to a non-equivalence hypothesis and quantify the evidence with Bayes factors. Typically, H\textsubscript{0} constitutes the equivalence interval (comparable to SESOI in the TOST framework), and H\textsubscript{a} constitutes the complementary non-equivalence regions. Formally, the Bayes factor is calculated by dividing the fraction \emph{posterior area inside the interval/posterior area outside the interval} (i.e., the posterior odds) by the fraction \emph{prior area inside the interval/prior area outside the interval} (i.e., the prior odds). The resulting value quantifies how much more likely it is that the data occurred under a population effect size deemed `equivalent' relative to the data having occurred under a population effect size deemed non-equivalent. The current implementation uses a default Cauchy prior on effect size with the possible scale parameters of medium (r = 1/\(\sqrt{2}\)), wide (r = 1), or ultra-wide (r = \(\sqrt{2}\)). For further reading, see Morey and Rouder (2011), Ravenzwaaij, Monden, Tendeiro, and Ioannidis (2019), and Linde, Tendeiro, Selker, Wagenmakers, and Ravenzwaaij (2020).

\hypertarget{parameters-1}{%
\subparagraph{Parameters}\label{parameters-1}}

\textbf{Delta:} The expected population effect size.\\
\textbf{TPR:} The desired long-run probability of obtaining a Bayes factor at least as high as the Threshold, given Delta.\\
\textbf{EqBand:} The chosen width of the equivalence region.\\
\textbf{PriorScale:} The scale of the Cauchy prior distribution. The PriorScale in the application can be set to: 1/\(\sqrt{2}\), 1, and \(\sqrt{2}\).\\
\emph{Threshold:} Critical threshold for the Bayes factor. The threshold level in the application can be set to 10, 6, or 3.\\

\hypertarget{how-to-use-the-package-1}{%
\subparagraph{How to use the package}\label{how-to-use-the-package-1}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{SampleSizePlanner}\SpecialCharTok{::}\FunctionTok{ssp\_eq\_bf}\NormalTok{(}\AttributeTok{tpr =} \FloatTok{0.8}\NormalTok{, }\AttributeTok{delta =} \DecValTok{0}\NormalTok{, }\AttributeTok{eq\_band =} \FloatTok{0.2}\NormalTok{,}
    \AttributeTok{thresh =} \DecValTok{10}\NormalTok{, }\AttributeTok{prior\_scale =} \DecValTok{1}\SpecialCharTok{/}\FunctionTok{sqrt}\NormalTok{(}\DecValTok{2}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{how-to-report-your-sample-size-estimation-1}{%
\subparagraph{How to report your sample size estimation}\label{how-to-report-your-sample-size-estimation-1}}

In order to estimate the sample size, we used the interval equivalent Bayes factor (Morey \& Rouder, 2011; Ravenzwaaij, Monden, Tendeiro, \& Ioannidis, 2019) method. We used a Cauchy prior distribution centered on zero with a scale of 1/\(\sqrt{2}\). We set the aimed TPR at 0.8, because {[}1) it is the common standard in the field; 2) it is the journal publishing requirement{]}. We consider all effect sizes below 0.2 equivalent to zero, because {[}1) previous studies reported the choice of a similar equivalence region; 2) of the following substantive reasons: \ldots{]}. The expected delta was 0 because {[}1) we expected no difference between the groups{]}. Our Bayes factor threshold for concluding equivalence was 10. Based on these parameters, a minimal sample size of 144 per group was estimated in order to reach 0.8 TPR for our design.

\hypertarget{effect-size-0-1}{%
\subsubsection{1.2. Effect size \textgreater0}\label{effect-size-0-1}}

\hypertarget{frequentist}{%
\paragraph{1.2.1. Frequentist}\label{frequentist}}

\hypertarget{classical-power-analysis}{%
\subparagraph{1.2.1.1. Classical power analysis}\label{classical-power-analysis}}

Study context

Mary would like to know what sample size she needs for a power of .80 to study whether the mean IQ score of the experimental group's population is significantly higher than the mean IQ score of the control group. She tests this assumption in a frequentist framework for a hypothetical population effect size of 0.5. This corresponds to a mean IQ score of 107.5 in the experimental group (100+15*.5), assuming a mean IQ score of 100 in the control group.

Description

The classical power analysis approach allows one to calculate the required sample size in order to obtain a significant result for the null hypothesis test a certain proportion of times in the long run given an assumed population effect size.

Parameters

\textbf{Delta:} The expected population effect size.\\
\textbf{TPR:} The desired long-run probability of obtaining a significant result with a one-sided \emph{t}-test, given Delta.\\
\textbf{Maximum N:} The maximum number of participants per group (both groups are assumed to have equal sample size).\\
\emph{Alpha:} The level of significance. Alpha is preset to 0.05 in the application.\\

How to use the package

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{SampleSizePlanner}\SpecialCharTok{::}\FunctionTok{ssp\_power\_traditional}\NormalTok{(}\AttributeTok{tpr =} \FloatTok{0.8}\NormalTok{, }\AttributeTok{delta =} \FloatTok{0.5}\NormalTok{,}
    \AttributeTok{max\_n =} \DecValTok{5000}\NormalTok{, }\AttributeTok{alpha =} \FloatTok{0.05}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

How to report your sample size estimation

We used a power analysis to estimate the sample size. We used an alpha of 0.05. We set the aimed TPR at 0.8, because {[}1) it is the common standard in the field; 2) it is the journal publishing requirement{]}. The expected delta was 0.5 because {[}1) previous results published in \ldots; 2) of the following substantive reasons: \ldots{]}. Based on these parameters, a minimal sample size of 51 per group was estimated in order to reach 0.8 TPR for our design.

\hypertarget{power-curve}{%
\subparagraph{1.2.1.2. Power curve}\label{power-curve}}

Study context

Mary would like to know what sample size she needs for a power of .80 to study whether the mean IQ score of the experimental group's population is significantly higher than the mean IQ score of the control group. She tests this assumption in a frequentist framework. However, she is reluctant to commit to a single hypothetical population effect size a-priori, preferring to calculate required sample size for a range of hypothetical deltas between 0.1 and 0.9.

Description

The power curve method is similar to a classical power analysis but instead of calculating the appropriate sample size for one hypothesized population effect size, the method calculates the required sample size for a range of plausible population effect sizes.

Parameters

\textbf{Delta:} A range of hypothetical population effect sizes.\\
\textbf{TPR:} The desired long-run probabilities of obtaining a significant result with a one-sided \emph{t}-test, given each value of Delta.\\
\textbf{Maximum N:} The maximum number of participants per group (both groups are assumed to have equal sample size).\\
\emph{Alpha:} The level of significance. Alpha is preset to 0.05 in the application.\\

How to use the package

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Determine the sample sizes for each delta}
\NormalTok{curve\_data }\OtherTok{\textless{}{-}}\NormalTok{ SampleSizePlanner}\SpecialCharTok{::}\FunctionTok{ssp\_power\_curve}\NormalTok{(}\AttributeTok{tpr =} \FloatTok{0.8}\NormalTok{, }\AttributeTok{delta =} \FunctionTok{seq}\NormalTok{(}\FloatTok{0.1}\NormalTok{,}
    \FloatTok{0.9}\NormalTok{, }\FloatTok{0.01}\NormalTok{), }\AttributeTok{max\_n =} \DecValTok{5000}\NormalTok{)}

\CommentTok{\# Plot the power curve}
\NormalTok{SampleSizePlanner}\SpecialCharTok{::}\FunctionTok{plot\_power\_curve}\NormalTok{(}\AttributeTok{delta =}\NormalTok{ curve\_data}\SpecialCharTok{$}\NormalTok{delta,}
    \AttributeTok{n1 =}\NormalTok{ curve\_data}\SpecialCharTok{$}\NormalTok{n1, }\AttributeTok{animated =} \ConstantTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

How to report your sample size estimation

We used a power analysis to estimate the sample size. We used an alpha of 0.05. We set the aimed TPR at 0.8, because {[}1) it is the common standard in the field; 2) it is the journal publishing requirement{]}. Because {[}1) we have no clear expectation of the magnitude of delta 2) we expected the delta to be around\ldots{]}, we include power calculations for delta ranging from 0.1 to 0.9. Based on these parameters, minimal sample sizes per group for different hypothetical effect sizes to reach 0.8 TPR can be found in Figure~\ref{fig:powercurve}.

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/powercurve-1}

}

\caption{The figure shows the resulting power curve created by the application. The X-axis shows the range of deltas from the example, while the Y-axis shows the corresponding sample sizes determined by the power curve method.}\label{fig:powercurve}
\end{figure}

\hypertarget{bayesian}{%
\paragraph{1.2.2. Bayesian}\label{bayesian}}

\hypertarget{predetermined-sample-size-with-bayes-factor}{%
\subparagraph{1.2.2.1. Predetermined sample size with Bayes factor}\label{predetermined-sample-size-with-bayes-factor}}

Study context

Mary would like to test whether the mean IQ score of the experimental group's population is higher than the mean IQ score of the control group. She'd like to know what sample size she needs to have for a long-run probability of .80 of obtaining a Bayes factor larger than 10. Mary plans to collect all her data in one batch without testing sequentially. Mary expects the population effect size to be 0.5. This corresponds to a mean IQ score of 107.5 (100+15*.5) in the experimental group, assuming a mean IQ score of 100 in the control group.

Description

The present method calculates the corresponding default Bayes factor for a \emph{t}-test statistic with Cauchy prior distribution centered on zero with scale parameter of either 1/\(\sqrt{2}\), 1, or \(\sqrt{2}\) for several sample sizes (the so-called Jeffrey-Zellner-Siow Bayes factor, see e.g., Rouder, Speckman, Sun, Morey, \& Iverson, 2009). The function returns the optimal sample size needed to reach the TPR for a given Bayes factor threshold to detect an expected population effect size. If a range of possible population effect sizes are plausible under the given hypothesis, the function can calculate the optimal sample sizes for the given range of effect sizes and present the results in a figure (analogous to the Power Curve method). This method is designed to determine the sample sizes for the existence of an effect (i.e., Delta \textgreater{} 0).

Parameters

\textbf{Delta:} The expected population effect size or a range of expected effect sizes.\\
\textbf{TPR:} The long-run probability of obtaining a Bayes factor at least as high as the critical threshold favoring superiority, given Delta.\\
\textbf{Maximum N:} The maximum number of participants per group (both groups are assumed to have equal sample size).\\
\textbf{PriorScale:} The scale of the Cauchy prior distribution. The PriorScale in the application can be set to: 1/\(\sqrt{2}\), 1, and \(\sqrt{2}\).\\
\textbf{Threshold:} Critical threshold for the Bayes factor. Three threshold levels are available in the app: 3, 6, and 10.\\

How to use the package

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{SampleSizePlanner}\SpecialCharTok{::}\FunctionTok{ssp\_bf\_predetermined}\NormalTok{(}\AttributeTok{tpr =} \FloatTok{0.8}\NormalTok{, }\AttributeTok{delta =} \FloatTok{0.5}\NormalTok{,}
    \AttributeTok{thresh =} \DecValTok{10}\NormalTok{, }\AttributeTok{max\_n =} \DecValTok{5000}\NormalTok{, }\AttributeTok{prior\_scale =} \DecValTok{1}\SpecialCharTok{/}\FunctionTok{sqrt}\NormalTok{(}\DecValTok{2}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

How to report your sample size estimation

We used the Jeffrey-Zellner-Siow Bayes factor method to estimate the sample size. We used a Cauchy prior distribution centered on zero with a scale of 1/\(\sqrt{2}\). We set the aimed TPR at 0.8, because {[}1) it is the common standard in the field; 2) it is the journal publishing requirement{]}. The expected delta was 0.5 because {[}1) previous results published in \ldots; of the following substantive reasons: \ldots{]}. Our evidence threshold was 10. Based on these parameters, a minimal sample size of 105 per group was estimated in order to reach a 0.8 TPR for our design.

\hypertarget{bayes-factor-design-analysis-bfda}{%
\subparagraph{1.2.2.2. Bayes Factor Design Analysis (BFDA)}\label{bayes-factor-design-analysis-bfda}}

Study context

Mary would like to know what sample size she needs to have a long-run probability of .80 of obtaining a Bayes factor larger than 10. Mary would like to test whether the mean IQ score of the experimental group's population is higher than the mean IQ score of the control group in a Bayesian framework. Mary plans to collect all her data incrementally and as such is interested in using the advantage of not testing more than strictly necessary offered by sequential testing in her Bayesian analysis. Mary expects the population effect size to be 0.5. This corresponds to a mean IQ score of 107.5 in the experimental group (100+15*.5), assuming a mean IQ score of 100 in the control group.

Description

The description of the BFDA method is functionally identical to the one provided in section `Predetermined sample size with Bayes factor,' but gains in TPR due to the addition of sequential testing. In the app, H\textsubscript{0} and H\textsubscript{a} indicate the proportion of times sequential testing leads to Bayes factors providing evidence with the given threshold for the null hypothesis and for the alternative hypothesis, respectively. Users of the Shiny app and R package should set Delta to 0 if they wish to determine the sufficient sample size for rejecting an effect, and use Delta \textgreater{} 0 if they wish to find support for the existence of an effect. For further reading, see Schönbrodt and Wagenmakers (2018) and Schönbrodt, Wagenmakers, Zehetleitner, and Perugini (2017).

Parameters

\textbf{Delta:} The expected population effect size.\\
\textbf{TPR:} The long run probability of obtaining a Bayes factor at least as high as the critical threshold favoring superiority, given Delta.\\
\textbf{PriorScale:} The scale of the Cauchy prior distribution. The PriorScale in the application can be set to: 1/\(\sqrt{2}\), 1, and \(\sqrt{2}\).\\
\textbf{Threshold:} Critical threshold for the Bayes factor. Three threshold levels are available in the app: 3, 6, and 10.\\

How to use the package

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{SampleSizePlanner}\SpecialCharTok{::}\FunctionTok{ssp\_bfda}\NormalTok{(}\AttributeTok{tpr =} \FloatTok{0.8}\NormalTok{, }\AttributeTok{delta =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{thresh =} \DecValTok{10}\NormalTok{,}
    \AttributeTok{n\_rep =} \DecValTok{10000}\NormalTok{, }\AttributeTok{prior\_scale =} \DecValTok{1}\SpecialCharTok{/}\FunctionTok{sqrt}\NormalTok{(}\DecValTok{2}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

How to report your sample size estimation

We used the BFDA method to estimate the sample size. We used a Cauchy prior distribution centered on zero with a scale of 1/\(\sqrt{2}\). We set the aimed TPR at 0.8, because {[}1) it is the common standard in the field; 2) it is the journal publishing requirement{]}. The expected delta was 0.5 because {[}1) previous results published in \ldots; 2) of the following substantive reasons: \ldots{]}. Our evidence threshold was 10. Based on these parameters, a minimal sample size of 81 per group was estimated in order to reach a 0.8 TPR for our design.

\hypertarget{estimation}{%
\subsection{2. Estimation}\label{estimation}}

\hypertarget{frequentist-1}{%
\subsubsection{2.1. Frequentist}\label{frequentist-1}}

\hypertarget{accuracy-in-parameter-estimation-aipe}{%
\paragraph{2.1.1. Accuracy In Parameter Estimation (AIPE)}\label{accuracy-in-parameter-estimation-aipe}}

\hypertarget{study-context-6}{%
\subparagraph{Study context}\label{study-context-6}}

Mary would like to know what sample size she needs, such that the 95\% confidence interval for the population effect size has an expected width of 0.4. She estimates the population effect size to be 0.2.

\hypertarget{description-6}{%
\subparagraph{Description}\label{description-6}}

Accuracy in parameter estimation aims to determine the sufficient sample size to obtain a confidence interval with a desired width (precision) around the expected effect size (Kelley \& Rausch, 2006). Note that the width of the calculated confidence interval will depend on the sample variance. As a result, it is possible that for a given sample the variance is relatively large, leading to a resulting confidence interval that is larger than the width of the desired interval for a given sample. Thus, the AIPE method aims to establish the expected value of the calculated confidence interval, which can be thought of as the 50\% long-run probability of obtaining a confidence interval no wider than the provided width.

\hypertarget{parameters-6}{%
\paragraph{Parameters}\label{parameters-6}}

\textbf{Delta:} The expected population effect size.\\
\textbf{Width:} The desired width of the confidence interval, given Delta.\\
\textbf{Confidence level:} The desired level of confidence.\\

\hypertarget{how-to-use-the-package-6}{%
\subparagraph{How to use the package}\label{how-to-use-the-package-6}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{SampleSizePlanner}\SpecialCharTok{::}\FunctionTok{ssp\_aipe}\NormalTok{(}\AttributeTok{delta =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{width =} \FloatTok{0.2}\NormalTok{, }\AttributeTok{confidence\_level =} \FloatTok{0.8}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{how-to-report-your-sample-size-estimation-6}{%
\subparagraph{How to report your sample size estimation}\label{how-to-report-your-sample-size-estimation-6}}

In order to estimate the sample size, we used the accuracy in parameter estimation {[}AIPE; Kelley and Rausch (2006){]} method. We aimed for a 95\% confidence level, because {[}1) it is the common standard in the field; 2) it is the journal publishing requirement{]}. The desired width was 0.4 because {[}1) previous studies reported the choice of a similar region of practical equivalence; 2) of the following substantive reasons: \ldots{]}. We expected an underlying population effect size of 0.3, because {[}1) previous results published in \ldots; 2) of the following substantive reasons: \ldots{]}. Based on these parameters, a minimal sample size of 195 per group was estimated for our design.

\hypertarget{a-priori-precision-app}{%
\paragraph{2.1.2. A Priori Precision (APP)}\label{a-priori-precision-app}}

Study context

Mary would like to know the sample size for which she will have a 95\% long-run probability that the sample means in both the experimental and the control group lie within 0.2 standard deviations (3 IQ points) of the true population mean.

Description

APP aims to determine the sample size needed to have a certain long-run probability of both sample means being within a certain range of their respective population means, expressed in terms of standard deviations (Trafimow \& MacDonald, 2017). As a result, APP is not reliant on the expected effect size.

\hypertarget{parameters-7}{%
\subparagraph{Parameters}\label{parameters-7}}

\textbf{Closeness:} The desired closeness of the sample mean to the population mean defined in standard deviation.\\
\textbf{Confidence:} The desired probability of obtaining the sample mean with the desired closeness to the population mean.\\

How to use the package

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{SampleSizePlanner}\SpecialCharTok{::}\FunctionTok{ssp\_app}\NormalTok{(}\AttributeTok{closeness =} \FloatTok{0.2}\NormalTok{, }\AttributeTok{confidence =} \FloatTok{0.95}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

How to report your sample size estimation

In order to estimate the sample size, we used the a-priori precision {[}APP; Trafimow and MacDonald (2017){]} method. Before data collection, we wanted to be 95\% confident that both sample means lie within 0.2 SD of the true population means. Based on these parameters, the resulting minimum sample size was 126 per group for our design.

\hypertarget{bayesian-testimation}{%
\subsubsection{2.2. Bayesian testimation}\label{bayesian-testimation}}

\hypertarget{region-of-practical-equivalence-rope}{%
\paragraph{2.2.1. Region of Practical Equivalence (ROPE)}\label{region-of-practical-equivalence-rope}}

Study context

Mary would like to conduct parameter estimation to see whether the mean IQ score of her experimental group's population is practically equivalent to 100. She would like to know what sample size she needs to have a long-run probability of .80 of obtaining a 95\% highest density interval that is contained within her predefined region of practical equivalence (ROPE). Mary hypothesizes that there is no difference (i.e., H\textsubscript{0} is true). She considers a population effect size between -0.2 and under 0.2 to be `practically equivalent.' This would correspond to IQ scores between 97 (100+15*-.2) and 103 (100+15*.2).

Description

The highest density interval region of practical equivalence technique (HDI-ROPE, often just referred to as ROPE) shares some features with the equivalence interval Bayes factor procedure. Both define an equivalence interval, construct a prior for the population effect size, and update to a posterior after the data comes in. The equivalence interval Bayes factor procedure then focuses on the posterior and prior odds under complementary hypotheses. The ROPE procedure, on the other hand, identifies the 95\% highest density interval (HDI; other percentages are permissible as well) and determines whether or not the HDI is fully contained within the equivalence interval. For further reading, see Kruschke (2018) and Kruschke (2011).

Parameters

\textbf{Delta:} The expected population effect size.\\
\textbf{TPR:} The desired long run probability of having the HDI fully contained within the ROPE interval, given Delta.\\
\textbf{EqBand:} The chosen ROPE interval.\\
\textbf{PriorScale:} The scale of the Cauchy prior distribution. The PriorScale in the application can be set to: 1/\(\sqrt{2}\), 1, and \(\sqrt{2}\).\\

How to use the package

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{SampleSizePlanner}\SpecialCharTok{::}\FunctionTok{ssp\_rope}\NormalTok{(}\AttributeTok{tpr =} \FloatTok{0.8}\NormalTok{, }\AttributeTok{delta =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{eq\_band =} \FloatTok{0.2}\NormalTok{,}
    \AttributeTok{prior\_scale =} \DecValTok{1}\SpecialCharTok{/}\FunctionTok{sqrt}\NormalTok{(}\DecValTok{2}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

How to report your sample size estimation

In order to estimate the sample size, we used the Region of Practical Equivalence (Kruschke, 2018) method. We used a Cauchy prior distribution centered on zero with a scale of 1/\(\sqrt{2}\). We set the aimed TPR at 0.8, because {[}1) it is the common standard in the field; 2) it is the journal publishing requirement{]}. We consider all effect sizes below 0.2 equivalent to zero, because {[}1) previous studies reported the choice of a similar region of practical equivalence; 2) of the following substantive reasons: \ldots{]}. The expected delta was 0 because {[}1) we expected no difference between the groups{]}. Based on these parameters, a minimal sample size of 517 per group was estimated in order to reach a 0.8 TPR for our design.

\hypertarget{summary}{%
\section{Summary}\label{summary}}

Justifying the decisions made during the sample size planning process presents valuable information when one evaluates the inferences drawn from a study. The Shiny app and R package presented in this paper aim to help researchers to choose and employ their sample size estimation method. In addition, the tool provides assistance in reporting the process and justification behind sample size choices. We encourage users and experts of the field to provide feedback and recommendations towards further developments.

\hypertarget{authors-contribution}{%
\section{Authors contribution}\label{authors-contribution}}

\textbf{Conceptualization:} Marton Kovacs, Don van Ravenzwaaij, Rink Hoekstra, and Balazs Aczel.\\
\textbf{Methodology:} Don van Ravenzwaaij.\\
\textbf{Project Administration:} Marton Kovacs.\\
\textbf{Software:} Marton Kovacs and Don van Ravenzwaaij.\\
\textbf{Supervision:} Balazs Aczel.\\
\textbf{Writing - Original Draft Preparation:} Marton Kovacs, Don van Ravenzwaaij, Rink Hoekstra, and Balazs Aczel.\\
\textbf{Writing - Review \& Editing:} Marton Kovacs, Don van Ravenzwaaij, Rink Hoekstra, and Balazs Aczel.

\hypertarget{notes}{%
\section{Notes}\label{notes}}

\hypertarget{glossary}{%
\subsection{Glossary}\label{glossary}}

\emph{Accuracy in Parameter Estimation (AIPE):} A sample size estimation method used for parameter estimation. The approach aims to find the required sample size, such that the confidence interval has a certain expected width.\\
\emph{Priori Procedure (APP):} The approach aims to plan a sample size based on how close the researcher wishes both sample means to be to their respective population parameter, and how confident the researcher wants to be in this.\\
\emph{Bayesian inference:} A general framework for updating one's prior beliefs in light of new data.\\
\emph{Bayes Factor Design Analysis (BFDA):} This technique provides an expected sample size such that compelling evidence in the form of a Bayes factor can be collected for a given effect size with a certain long-run probability when allowing for sequential testing.\\
\emph{Testing vs.~Estimation:} Two schools of inference, focusing on establishing whether or not an effect exists versus establishing the magnitude of an effect, respectively.\\
\emph{Equivalence band (EqBand):} The region of effect sizes considered practically equivalent to zero. In our paper, SESOI and ROPE are subsumed under EqBand.\\
\emph{Frequentist inference:} A general framework in which probabilities are defined as frequencies in hypothetical repeated events. In the context of statistical testing, frequentist inference is concerned with long-run error rates of rejecting the null hypothesis for the observed or more extreme parameters in a given design when the model assumptions (e.g., independence of observations) are true.\\
\emph{Statistical power:} The long-run probability of finding a significant effect given a certain population effect size.\\
\emph{True positive rate (TPR):} The long-run probability of finding evidence for an effect, given that it exists. In our paper, statistical power is subsumed under TPR.\\
\emph{Classical power analysis:} This method is used to estimate the minimum sample size that a design needs to reach a certain level of statistical power, given a desired significance level and expected effect size.\\
\emph{Power-curve:} This curve shows how changes in effect size modify the statistical power of a test.\\
\emph{Region Of Practical Equivalence (ROPE):} The region of effect sizes considered practically equivalent to zero under the HDI-ROPE method.\\
\emph{Smallest Effect Size Of Interest (SESOI):} The region of effect sizes considered practically equivalent to zero under the TOST method.\\
\emph{Sequential testing:} The practice of incrementally testing as data comes in, typically until some pre-determined level of evidence is obtained.\\
\emph{Two One‐Sided Tests (TOST):} A frequentist statistical testing approach aimed at establishing equivalence between two groups.\\
\emph{Equivalence interval BF:} A Bayesian statistical testing approach aimed at establishing equivalence between two groups.\\

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-halpern_continuing_2002}{}%
Halpern, S. D., Karlawish, J. H., \& Berlin, J. A. (2002). The continuing unethical conduct of underpowered clinical trials. \emph{Jama}, \emph{288}(3), 358--362.

\leavevmode\hypertarget{ref-ioannidis_why_2005}{}%
Ioannidis, J. P. (2005). Why most published research findings are false. \emph{{PLoS} Medicine}, \emph{2}(8), e124.

\leavevmode\hypertarget{ref-kelley_sample_2006}{}%
Kelley, K., \& Rausch, J. R. (2006). Sample size planning for the standardized mean difference: Accuracy in parameter estimation via narrow confidence intervals. \emph{Psychological Methods}, \emph{11}(4), 363--385.

\leavevmode\hypertarget{ref-kruschke_bayesian_2011}{}%
Kruschke, J. K. (2011). Bayesian assessment of null values via parameter estimation and model comparison. \emph{Perspectives on Psychological Science}, \emph{6}(3), 299--312.

\leavevmode\hypertarget{ref-kruschke_rejecting_2018}{}%
Kruschke, J. K. (2018). Rejecting or accepting parameter values in bayesian estimation. \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(2), 270--280.

\leavevmode\hypertarget{ref-lakens_sample_2021}{}%
Lakens, D. (2021). Sample size justification. Retrieved from \url{https://doi.org/10.31234/osf.io/9d3yf}

\leavevmode\hypertarget{ref-lakens_equivalence_2018}{}%
Lakens, D., Scheel, A. M., \& Isager, P. M. (2018). Equivalence testing for psychological research: A tutorial. \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(2), 259--269.

\leavevmode\hypertarget{ref-linde_tendeiro_selker_wagenmakers_van_ravenzwaaij_2020}{}%
Linde, M., Tendeiro, J., Selker, R., Wagenmakers, E.-J., \& Ravenzwaaij, D. van. (2020). Decisions about equivalence: A comparison of TOST, HDI-ROPE, and the bayes factor. PsyArXiv. \url{https://doi.org/10.31234/osf.io/bh8vu}

\leavevmode\hypertarget{ref-maxwell_persistence_2004}{}%
Maxwell, S. E. (2004). The persistence of underpowered studies in psychological research: Causes, consequences, and remedies. \emph{Psychological Methods}, \emph{9}(2), 147--163.

\leavevmode\hypertarget{ref-morey_bayes_2011}{}%
Morey, R. D., \& Rouder, J. N. (2011). Bayes factor approaches for testing interval null hypotheses. \emph{Psychological Methods}, \emph{16}(4), 406--419.

\leavevmode\hypertarget{ref-van_ravenzwaaij_bayes_2019}{}%
Ravenzwaaij, D. van, Monden, R., Tendeiro, J. N., \& Ioannidis, J. P. (2019). Bayes factors for superiority, non-inferiority, and equivalence designs. \emph{{BMC} Medical Research Methodology}, \emph{19}(1), 1--12.

\leavevmode\hypertarget{ref-rouder_bayesian_2009}{}%
Rouder, J. N., Speckman, P. L., Sun, D., Morey, R. D., \& Iverson, G. (2009). Bayesian t tests for accepting and rejecting the null hypothesis. \emph{Psychonomic Bulletin \& Review}, \emph{16}(2), 225--237.

\leavevmode\hypertarget{ref-schonbrodt_bayes_2018}{}%
Schönbrodt, F. D., \& Wagenmakers, E.-J. (2018). Bayes factor design analysis: Planning for compelling evidence. \emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 128--142.

\leavevmode\hypertarget{ref-schonbrodt_sequential_2017}{}%
Schönbrodt, F. D., Wagenmakers, E.-J., Zehetleitner, M., \& Perugini, M. (2017). Sequential hypothesis testing with bayes factors: Efficiently testing mean differences. \emph{Psychological Methods}, \emph{22}(2), 322--339.

\leavevmode\hypertarget{ref-schuirmann_comparison_1987}{}%
Schuirmann, D. J. (1987). A comparison of the two one-sided tests procedure and the power approach for assessing the equivalence of average bioavailability. \emph{Journal of Pharmacokinetics and Biopharmaceutics}, \emph{15}(6), 657--680.

\leavevmode\hypertarget{ref-trafimow_performing_2017}{}%
Trafimow, D., \& MacDonald, J. A. (2017). Performing inferential statistics prior to data collection. \emph{Educational and Psychological Measurement}, \emph{77}(2), 204--219.

\end{CSLReferences}

\endgroup

\end{document}

### 2021-09-02 inst/app/www/about.Rmd
renaming, tidying, recalculating ssp results, debugging

@@
 
-The __SampleSizePlanner__ shiny application enables researchers to calculate the sample size with 9 different methods in app and provides code to use an additional method from R.
+The **SampleSizePlanner** shiny application enables researchers to calculate the sample size with 9 different methods in app and provides code to use an additional method from R.
 The aim of the app is to make the various sample size calculating methods available for a wide range of researchers and propagate the transparent reporting of the chosen method.
@@
 ### Contributors
+
 -   Marton Kovacs
@@
 
+The design of the app was done by Zsuzsa Szekely [(@szekely_Zsuzsa)](<https://twitter.com/szekely_zsuzsa>).
+
 ### Questions and comments
 
-If you have any questions or comments please write to Marton Kovacs to the following email address: marton.balazs.kovacs@gmail.com
+If you have any questions or comments please write to Marton Kovacs to the following email address: [marton.balazs.kovacs@gmail.com](mailto:marton.balazs.kovacs@gmail.com)
 
-Or submit an issue on the apps Github page: [https://github.com/marton-balazs-kovacs/SampleSizePlanner](https://github.com/marton-balazs-kovacs/SampleSizePlanner)
+Or submit an issue on the apps Github page: <https://github.com/marton-balazs-kovacs/SampleSizePlanner>

### 2021-09-02 inst/manuscript/manuscript.Rmd
renaming, tidying, recalculating ssp results, debugging

@@
 library("papaja")
+library("ggplot2")
 library("formatR")
@@
 Mary has come up with a new set of games that challenge spatial insight.
-She believes that distributed and targeted engagement with these games for a period of six months for children in the age range of 8 to 12 will lead to lasting improvements on their IQ score as measured through Raven's progressive matrices test (population mean 100, population SD 15).
+She would like to test whether distributed and targeted engagement with these games for a period of six months for children in the age range of 8 to 12 will lead to lasting improvements on their IQ score as measured through Raven's progressive matrices test (population mean 100, population SD 15).
 Mary collects data for a control sample that gets regular education and for an experimental sample and plans to compare those samples.
-For illustrative purposes, we will have Mary's hypothetical study goal depend on the earlier presented decisions to illustrate different scenarios for each sample size planning method.
-That means that in some of the upcoming examples, Mary expects a null effect and in others she expects a positive effect.
-We will also present a justification text based on Mary’s choices for each method.
+Mary has good reason to be skeptical about the effectiveness of training on increasing performance as there are several studies questioning the existence of such effects [@owen_putting_2010; @simons_brain-training_2016].
+For illustrative purposes, in some of the upcoming examples Mary expects a null effect and in others Mary expects a positive effect in order to highlight the different research scenarios for each sample size planning method.
+We will also present a justification text for each sample size planning method based on Mary’s choices described in the example research scenario for the given method.
 
@@
   )
+
+# Save the plot in pdf
+ggplot2::ggsave("fig2.pdf", plot = last_plot(), dpi = 300)
 ```

### 2021-09-02 inst/manuscript/manuscript.tex
renaming, tidying, recalculating ssp results, debugging

@@
 It is crucial to not just state how we determined our planned sample size but to also give the reader insight into the reasons behind our choices.
-In a recent overview, Lakens (2021) lists six types of general approaches to justify sample size in quantitative empirical studies: (1) Measure entire population; (2) Resource constraints; (3) A-priori power analysis; (4) Accuracy; (5) Heuristics; and (6) No justification.
+In a recent overview, Daniel Lakens (2021) lists six types of general approaches to justify sample size in quantitative empirical studies: (1) Measure entire population; (2) Resource constraints; (3) A-priori power analysis; (4) Accuracy; (5) Heuristics; and (6) No justification.
 For the first approach, no quantitative justification is necessary; and for the second approach, the researcher has no freedom to increase the sample size.
@@
 Our tutorial mainly focuses on approaches two, three, and four, and is aimed at providing a hands-on approach for the mechanical part of the sample size determination (i.e., the calculation).
-For a deeper discussion of justification of these approaches, or for other approaches (i.e., using heuristics or not providing justification), we refer the reader to Lakens (2021).
+For a deeper discussion of justification of these approaches, or for other approaches (i.e., using heuristics or not providing justification), we refer the reader to Daniel Lakens (2021).
 
@@
 Mary has come up with a new set of games that challenge spatial insight.
-She believes that distributed and targeted engagement with these games for a period of six months for children in the age range of 8 to 12 will lead to lasting improvements on their IQ score as measured through Raven's progressive matrices test (population mean 100, population SD 15).
+She would like to test whether distributed and targeted engagement with these games for a period of six months for children in the age range of 8 to 12 will lead to lasting improvements on their IQ score as measured through Raven's progressive matrices test (population mean 100, population SD 15).
 Mary collects data for a control sample that gets regular education and for an experimental sample and plans to compare those samples.
-For illustrative purposes, we will have Mary's hypothetical study goal depend on the earlier presented decisions to illustrate different scenarios for each sample size planning method.
-That means that in some of the upcoming examples, Mary expects a null effect and in others she expects a positive effect.
-We will also present a justification text based on Mary's choices for each method.
+Mary has good reason to be skeptical about the effectiveness of training on increasing performance as there are several studies questioning the existence of such effects (Owen et al., 2010; Simons et al., 2016).
+For illustrative purposes, in some of the upcoming examples Mary expects a null effect and in others Mary expects a positive effect in order to highlight the different research scenarios for each sample size planning method.
+We will also present a justification text for each sample size planning method based on Mary's choices described in the example research scenario for the given method.
 
@@
 The lower and upper boundaries of the interval are determined by the equivalence band (i.e.~SESOI) around the expected population effect size (e.g., 0).
-Lakens, Scheel, and Isager (2018) lists several methods that can be used to determine the SESOI .
+Daniël Lakens, Scheel, and Isager (2018) lists several methods that can be used to determine the SESOI .
 In case of TOST, the two null hypotheses state that the effect size is equal to the lower and upper equivalence band values, whereas the alternative hypotheses state that the effect size is significantly smaller than the upper equivalence band value and significantly larger than the lower equivalence band value.
 In case both one-sided tests reject the null-hypothesis at a given significance level, the group means are considered to be practically equivalent.
-See Lakens, Scheel, and Isager (2018), for further reading.
+See Daniël Lakens, Scheel, and Isager (2018), for further reading.
 
@@
 \leavevmode\hypertarget{ref-ioannidis_why_2005}{}%
-Ioannidis, J. P. (2005). Why most published research findings are false. \emph{{PLoS} Medicine}, \emph{2}(8), e124.
+Ioannidis, J. P. (2005). Why most published research findings are false. \emph{PLoS Medicine}, \emph{2}(8), e124.
 
@@
 (2006).
-Sample size planning for the standardized mean difference: Accuracy in parameter estimation via narrow confidence intervals.
-\emph{Psychological Methods}, \emph{11}(4), 363--385.
+Sample size planning for the standardized mean difference: {Accuracy} in parameter estimation via narrow confidence intervals.
+\emph{Psychological Methods}, \emph{11}(4), 363.
 
@@
 \leavevmode\hypertarget{ref-kruschke_rejecting_2018}{}%
-Kruschke, J. K. (2018). Rejecting or accepting parameter values in bayesian estimation. \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(2), 270--280.
+Kruschke, J. K. (2018). Rejecting or accepting parameter values in {Bayesian} estimation. \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(2), 270--280.
 
 \leavevmode\hypertarget{ref-lakens_sample_2021}{}%
-Lakens, D. (2021). Sample size justification. Retrieved from \url{https://doi.org/10.31234/osf.io/9d3yf}
+Lakens, Daniel. (2021). Sample {Size} {Justification}.
 
 \leavevmode\hypertarget{ref-lakens_equivalence_2018}{}%
-Lakens, D., Scheel, A.
+Lakens, Daniël, Scheel, A.
 M., \& Isager, P.
@@
 (2018).
-Equivalence testing for psychological research: A tutorial.
+Equivalence testing for psychological research: {A} tutorial.
 \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(2), 259--269.
@@
 \leavevmode\hypertarget{ref-maxwell_persistence_2004}{}%
-Maxwell, S. E. (2004). The persistence of underpowered studies in psychological research: Causes, consequences, and remedies. \emph{Psychological Methods}, \emph{9}(2), 147--163.
+Maxwell, S. E. (2004). The persistence of underpowered studies in psychological research: Causes, consequences, and remedies. \emph{Psychological Methods}, \emph{9}(2), 147.
 
 \leavevmode\hypertarget{ref-morey_bayes_2011}{}%
-Morey, R. D., \& Rouder, J. N. (2011). Bayes factor approaches for testing interval null hypotheses. \emph{Psychological Methods}, \emph{16}(4), 406--419.
+Morey, R. D., \& Rouder, J. N. (2011). Bayes factor approaches for testing interval null hypotheses. \emph{Psychological Methods}, \emph{16}(4), 406.
+
+\leavevmode\hypertarget{ref-owen_putting_2010}{}%
+Owen, A. M., Hampshire, A., Grahn, J. A., Stenton, R., Dajani, S., Burns, A. S., \ldots{} Ballard, C. G. (2010). Putting brain training to the test. \emph{Nature}, \emph{465}(7299), 775--778.
 
@@
 Bayes factors for superiority, non-inferiority, and equivalence designs.
-\emph{{BMC} Medical Research Methodology}, \emph{19}(1), 1--12.
+\emph{BMC Medical Research Methodology}, \emph{19}(1), 1--12.
 
@@
 \leavevmode\hypertarget{ref-schonbrodt_bayes_2018}{}%
-Schönbrodt, F. D., \& Wagenmakers, E.-J. (2018). Bayes factor design analysis: Planning for compelling evidence. \emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 128--142.
+Schönbrodt, F. D., \& Wagenmakers, E.-J. (2018). Bayes factor design analysis: {Planning} for compelling evidence. \emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 128--142.
 
@@
 (2017).
-Sequential hypothesis testing with bayes factors: Efficiently testing mean differences.
-\emph{Psychological Methods}, \emph{22}(2), 322--339.
+Sequential hypothesis testing with {Bayes} factors: {Efficiently} testing mean differences.
+\emph{Psychological Methods}, \emph{22}(2), 322.
 
@@
 \emph{Journal of Pharmacokinetics and Biopharmaceutics}, \emph{15}(6), 657--680.
+
+\leavevmode\hypertarget{ref-simons_brain-training_2016}{}%
+Simons, D.
+J., Boot, W.
+R., Charness, N., Gathercole, S.
+E., Chabris, C.
+F., Hambrick, D.
+Z., \& Stine-Morrow, E.
+A.
+(2016).
+Do {``brain-training''} programs work?
+\emph{Psychological Science in the Public Interest}, \emph{17}(3), 103--186.
 

### 2021-09-02 R/jusfitication.R

@@ -33,16 +33,15 @@ justification <- function(method, output_parameters) {
       glue::glue(
         "In order to calculate an appropriate sample size for testing whether the \\
         two groups are practically equivalent, we used the Two One-Sided Tests of \\
-        Equivalence (TOST; Schuirmann, 1981) method. We used an alpha of {alpha}. We set the aimed TPR to be {tpr}, \\
+        Equivalence (TOST; Schuirmann, 1981) method. We used an alpha of 0.05. We set the aimed TPR to be {tpr}, \\
         because {tpr_justification}. We consider all effect sizes below {eq_band} \\
         equivalent to zero, because {eq_band_justification}. The expected delta was {delta} \\
         because {delta_justification}. Based on these parameters, a sample size of {n1} \\
-        per group was estimated in order to reach a TPR of {round(npower, 1)} with our design.",
+        per group was estimated in order to reach a TPR of {round(tpr_out, 1)} with our design.",
         eq_band = output_parameters$eq_band,
         delta = output_parameters$delta,
-        alpha = output_parameters$alpha,
         n1 = output_parameters$n1,
-        npower = output_parameters$npower,
+        tpr_out = output_parameters$tpr_out,
         tpr = output_parameters$tpr,
         tpr_justification = output_parameters$tpr_justification,
         eq_band_justification = output_parameters$eq_band_justification,
@@ -57,12 +56,12 @@ justification <- function(method, output_parameters) {
         {tpr}, because {tpr_justification}. We consider all effect sizes below {eq_band} equivalent to zero, \\
         because {eq_band_justification}. The expected delta was {delta} because {delta_justification}. \\
         Our Bayes factor threshold for concluding equivalence was {thresh}. Based on these parameters, a minimal \\
-        sample size of {n1} was estimated in order to reach {round(npower, 1)} TPR for our design.",
+        sample size of {n1} was estimated in order to reach {round(tpr_out, 1)} TPR for our design.",
         eq_band = output_parameters$eq_band,
         thresh = output_parameters$thresh, 
         delta = output_parameters$delta,
         n1 = output_parameters$n1,
-        npower = output_parameters$npower,
+        tpr_out = output_parameters$tpr_out,
         tpr = output_parameters$tpr,
         prior_scale = output_parameters$prior_scale,
         tpr_justification = output_parameters$tpr_justification,
@@ -72,13 +71,12 @@ justification <- function(method, output_parameters) {
   } else if (method == "traditional") {
     text <- 
       glue::glue(
[…]

### 2021-09-02 R/mod_ssp_bf_predetermined.R

@@ -104,7 +104,7 @@ mod_ssp_bf_predetermined_ui <- function(id) {
                 multiple = FALSE,
                 options = list(create = TRUE)),
               # Create justification text
-              actionButton(NS(id, "justification"), "Create justification report", class = "calculate-btn"),
+              actionButton(NS(id, "justification"), "Create justification report", class = "calculate-btn justification-btn"),
               # Show justification text
               mod_preview_ui(NS(id, "preview"))),
             tabPanel(
@@ -151,19 +151,21 @@ mod_ssp_bf_predetermined_server <- function(id) {
     output$calculate_output <- renderUI({
       HTML(
         glue::glue(
-          "<b>n1:</b> {n1}<br/><b>Resulting TPR:</b> {npower}",
+          "<b>n1:</b> {n1}<br/><b>Resulting TPR:</b> {tpr_out}",
           n1 = bf_predetermined_result()$n1,
-          npower = round(bf_predetermined_result()$npower, 2)
+          tpr_out = round(bf_predetermined_result()$tpr_out, 2)
         )
       )
     })
     
     # Add justification enable logic
     observe({
-      if (input$calculate) {
+      if (input$calculate && !is.na(bf_predetermined_result()$n1)) {
         shinyjs::enable("justification")
+        shinyjs::runjs("$('.justification-btn').removeAttr('title');")
       } else{
         shinyjs::disable("justification")
+        shinyjs::runjs("$('.justification-btn').attr('title', 'Please run the calculation first');")
       }
     })
 
@@ -175,7 +177,7 @@ mod_ssp_bf_predetermined_server <- function(id) {
         delta = input$delta,
         delta_justification = input$delta_justification,
         n1 = bf_predetermined_result()$n1,
-        npower = bf_predetermined_result()$npower,
+        tpr_out = bf_predetermined_result()$tpr_out,
         thresh = as.integer(input$thresh),
         prior_scale = input$prior_scale
       )
@@ -185,6 +187,7 @@ mod_ssp_bf_predetermined_server <- function(id) {
     mod_preview_server(
       "preview",
       activate = reactive(input$justification),
+      deactivate = reactive(input$calculate),
       output_parameters = output_parameters,
       method = "bf_predetermined")
     

### 2021-09-02 R/mod_ssp_power_curve.R

@@ -88,14 +88,15 @@ mod_ssp_power_curve_ui <- function(id) {
                 multiple = FALSE,
                 options = list(create = TRUE)),
               # Create justification text
-              actionButton(NS(id, "justification"), "Create justification report", class = "calculate-btn"),
+              actionButton(NS(id, "justification"), "Create justification report", class = "calculate-btn justification-btn"),
               # Show justification text
               mod_preview_ui(NS(id, "preview")),
               wellPanel(
                 style = "margin-top: 15px;",
-                plotly:: plotlyOutput(NS(id, "figure_preview")),
-                div(class = "download-btn",
-                    downloadButton(NS(id, "figure_download"), label = "Download figure")))),
+                plotly:: plotlyOutput(NS(id, "figure_preview"))),
+              div(class = "download-btn",
+                  downloadButton(NS(id, "figure_download"), label = "Download figure"))
+              ),
             tabPanel(
               "Code",
               # Panel title
@@ -119,15 +120,6 @@ mod_ssp_power_curve_server <- function(id) {
       # Setup loadingbar
       # waitress <- waiter::Waitress$new("#curve-preview-show_preview", theme = "overlay", infinite = TRUE, hide_on_render = TRUE)
       
-      # Add downloadbutton enable logic
-      observe({
-        if (input$calculate) {
-          shinyjs::enable("justification")
-        } else{
-          shinyjs::disable("justification")
-        }
-      })
-      
       # Create delta vector
       delta <- reactive({
         seq(input$delta[1], input$delta[2], 0.01)
@@ -141,6 +133,7 @@ mod_ssp_power_curve_server <- function(id) {
       
       # Show calculated results
       output$calculate_output <- renderUI({
+        if ("n1" %in% names(curve_result())) {
         HTML(
           glue::glue(
             "<b>n1:</b> {n1} <b>delta:</b> {delta}<br/>",
@@ -148,6 +141,25 @@ mod_ssp_power_curve_server <- function(id) {
             delta = curve_result()$delta
           )
         )
+        } else {
+          HTML(
+            glue::glue(
+              "<b>{error_message}</b>",
+              error_message = curve_result()$message
+            )
+          )
+        }
+      })
+      
+      # Add downloadbutton enable logic
+      observe({
+        if (input$calculate  && "n1" %in% names(curve_result())) {
+          shinyjs::enable("justification")
+          shinyjs::runjs("$('.justification-btn').removeAttr('title');")
+        } else{
[…]

### 2021-09-02 R/mod_ssp_power_traditional.R

@@ -90,7 +90,7 @@ mod_ssp_power_traditional_ui <- function(id) {
                 multiple = FALSE,
                 options = list(create = TRUE)),
               # Create justification text
-              actionButton(NS(id, "justification"), "Create justification report", class = "calculate-btn"),
+              actionButton(NS(id, "justification"), "Create justification report", class = "calculate-btn justification-btn"),
               # Show justification text
               mod_preview_ui(NS(id, "preview"))),
             tabPanel(
@@ -119,26 +119,43 @@ mod_ssp_power_traditional_server <- function(id) {
     # Calculate results
     traditional_result <- eventReactive(input$calculate, {
       # waitress$start()
-      ssp_power_traditional(tpr = input$tpr, delta = input$delta, max_n = input$max_n)
+      safe_ssp_power_traditional <- purrr::safely(ssp_power_traditional)
+      res <- safe_ssp_power_traditional(tpr = input$tpr, delta = input$delta, max_n = input$max_n)
+      if (is.null(res$error)) {
+        res$result
+      } else {
+        res$error
+      }
     })
     
     # Show calculated results
     output$calculate_output <- renderUI({
-      HTML(
-        glue::glue(
-          "<b>n1:</b> {n1}<br/><b>Resulting TPR:</b> {npower}",
-          n1 = traditional_result()$n1,
-          npower = round(traditional_result()$npower, 2)
+      if ("n1" %in% names(traditional_result())) {
+        HTML(
+          glue::glue(
+            "<b>n1:</b> {n1}<br/><b>Resulting TPR:</b> {tpr_out}",
+            n1 = traditional_result()$n1,
+            tpr_out = round(traditional_result()$tpr_out, 2)
+          )
         )
-      )
+      } else {
+        HTML(
+          glue::glue(
+            "<b>{error_message}</b>",
+            error_message = traditional_result()$message
+          )
+        )
+      }
     })
     
     # Add justification enable logic
     observe({
-      if (input$calculate) {
+      if (input$calculate && "n1" %in% names(traditional_result())) {
         shinyjs::enable("justification")
+        shinyjs::runjs("$('.justification-btn').removeAttr('title');")
       } else{
         shinyjs::disable("justification")
+        shinyjs::runjs("$('.justification-btn').attr('title', 'Please run the calculation first');")
       }
     })
 
@@ -150,7 +167,7 @@ mod_ssp_power_traditional_server <- function(id) {
         delta_justification = input$delta_justification,
[…]

### 2021-09-02 R/mod_ssp_rope.R

@@ -42,7 +42,7 @@ mod_ssp_rope_ui <- function(id){
           min = 0.1,
           max = 0.5,
           value = 0.2,
-          step = 0.01),
+          step = 0.05),
         sliderInput(
           NS(id, "delta"),
           name_with_info(
@@ -103,7 +103,7 @@ mod_ssp_rope_ui <- function(id){
                 multiple = FALSE,
                 options = list(create = TRUE)),
             # Create justification text
-            actionButton(NS(id, "justification"), "Create justification report", class = "calculate-btn"),
+            actionButton(NS(id, "justification"), "Create justification report", class = "calculate-btn justification-btn"),
             # Show justification text
             mod_preview_ui(NS(id, "preview"))
           ),
@@ -129,44 +129,59 @@ mod_ssp_rope_server <- function(id){
   moduleServer(id, function(input, output, session) {
     # Setup loadingbar
     # waitress <- waiter::Waitress$new("#rope-preview-show_preview", theme = "overlay", infinite = TRUE, hide_on_render = TRUE)
-    # Calculate results
-    rope_result <- eventReactive(input$calculate, {
-      # waitress$start()
-      prior_scale <- switch(
+    
+    input_prior_scale <- reactive({
+      switch(
         input$prior_scale,
         "1/sqrt(2)" = 1/sqrt(2),
         "1" = 1,
         "sqrt(2)" = sqrt(2)
       )
+    })
+    
+    # Calculate results
+    rope_result <- eventReactive(input$calculate, {
+      # waitress$start()
       
       rope_precalculation_results %>% 
         dplyr::filter(
           dplyr::near(tpr, input$tpr),
           dplyr::near(delta, input$delta),
           dplyr::near(eq_band, input$eq_band),
-          prior_scale == prior_scale
+          dplyr::near(prior_scale, input_prior_scale())
         ) %>% 
-        dplyr::select(n1, npower, error_message) %>% 
+        dplyr::select(n1, tpr_out, error_message) %>% 
         as.list()
       })
     
     # Show calculated results
     output$calculate_output <- renderUI({
+      if (!is.na(rope_result()$n1)) {
       HTML(
         glue::glue(
-          "<b>n1:</b> {n1}<br/><b>Resulting TPR:</b> {npower}",
+          "<b>n1:</b> {n1}<br/><b>Resulting TPR:</b> {tpr_out}",
           n1 = rope_result()$n1,
[…]

### 2021-09-02 R/mod_ssp_tost.R

@@ -100,7 +100,7 @@ mod_ssp_tost_ui <- function(id) {
                 multiple = FALSE,
                 options = list(create = TRUE)),
               # Create justification text
-              actionButton(NS(id, "justification"), "Create justification report", class = "calculate-btn"),
+              actionButton(NS(id, "justification"), "Create justification report", class = "calculate-btn justification-btn"),
               # Show justification text
               mod_preview_ui(NS(id, "preview"))),
             tabPanel(
@@ -128,26 +128,43 @@ mod_ssp_tost_server <- function(id) {
     # Calculate results
     tost_result <- eventReactive(input$calculate, {
       # waitress$start()
-      ssp_tost(tpr = input$tpr, eq_band = input$eq_band, delta = input$delta)
+      safe_ssp_tost <- purrr::safely(ssp_tost)
+      res <- safe_ssp_tost(tpr = input$tpr, eq_band = input$eq_band, delta = input$delta)
+      if (is.null(res$error)) {
+        res$result
+        } else {
+          res$error
+          }
       })
     
     # Show calculated results
     output$calculate_output <- renderUI({
+      if ("n1" %in% names(tost_result())) {
       HTML(
         glue::glue(
-          "<b>n1:</b> {n1}<br/><b>Resulting TPR:</b> {npower}",
+          "<b>n1:</b> {n1}<br/><b>Resulting TPR:</b> {tpr_out}",
           n1 = tost_result()$n1,
-          npower = round(tost_result()$npower, 2)
+          tpr_out = round(tost_result()$tpr_out, 2)
           )
         )
+      } else {
+        HTML(
+          glue::glue(
+            "<b>{error_message}</b>",
+            error_message = tost_result()$message
+          )
+        )
+      }
     })
     
     # Add justification enable logic
     observe({
-      if (input$calculate) {
+      if (input$calculate && "n1" %in% names(tost_result())) {
         shinyjs::enable("justification")
+        shinyjs::runjs("$('.justification-btn').removeAttr('title');")
       } else{
         shinyjs::disable("justification")
+        shinyjs::runjs("$('.justification-btn').attr('title', 'Please run the calculation first');")
       }
     })
 
@@ -161,7 +178,7 @@ mod_ssp_tost_server <- function(id) {
         eq_band_justification = input$eq_band_justification,
         tpr_justification = input$tpr_justification,
         n1 = tost_result()$n1,
-        npower = tost_result()$npower
[…]

### 2021-09-02 data-raw/eq_bf_precalculation_results.R

@@ -1,18 +1,6 @@
 # Load necessary packages
 library(tidyverse)
 
-# Create datatable storing possible options
-eq_bf_options <- 
-  expand.grid(
-    tpr = seq(0.5, 0.95, by = 0.05),
-    eq_band = seq(0.1, 0.5, by = 0.05),
-    delta = seq(0, 2, by = 0.05),
-    thresh = c(3, 6, 10),
-    prior_scale = c(1 / sqrt(2), 1, sqrt(2))
-  )
-
-eq_bf_options$iterate <- 1:nrow(eq_bf_options)
-
 # Read the outputs of all iterations
 eq_bf_data <- 
   tibble(filename = list.files(path = "./eq-bf-res/", pattern = ".rds")) %>%
@@ -21,16 +9,21 @@ eq_bf_data <-
   unnest(data) %>% 
   mutate(result = map(data, "result"),
          error = map(data, "error"),
+         params = map(data, "params"),
          error_message = map_chr(error, ~ pluck(.x, "message", .default = NA_character_)),
          result_not_null = if_else(map_lgl(result, ~ !is.null(.x)),
                                    1L,
                                    0L),
          n1 = map_dbl(result, ~ pluck(.x, "n1", .default = NA_real_)),
-         npower = map_dbl(result, ~ pluck(.x, "npower", .default = NA_real_)),
-         iterate = as.integer(names(data))) %>% 
-  rename(tpr_out = npower) %>% 
-  select(-file, -filename) %>% 
-  left_join(., eq_bf_options, by = "iterate")
+         tpr_out = map_dbl(result, ~ pluck(.x, "tpr_out", .default = NA_real_)),
+         tpr = map_dbl(params, ~ pluck(.x, "tpr", .default = NA_real_)),
+         delta = map_dbl(params, ~ pluck(.x, "delta", .default = NA_real_)),
+         prior_scale = map_dbl(params, ~ pluck(.x, "prior_scale", .default = NA_real_)),
+         eq_band = map_dbl(params, ~ pluck(.x, "eq_band", .default = NA_real_)),
+         thresh = map_dbl(params, ~ pluck(.x, "thresh", .default = NA_real_)),
+         iterate = map_dbl(params, ~ pluck(.x, "iterate", .default = NA_real_))
+         ) %>% 
+  select(-file, -filename)
 
 # Check the number of cases where the function calculated the n1
[…]

### 2021-09-02 data-raw/rope_precalculation_results.R

@@ -1,19 +1,6 @@
 # Load necessary packages
 library(tidyverse)
 
-# Create datatable storing possible ROPE options
-rope_options <- 
-  expand.grid(
-    tpr = seq(0.5, 0.95, by = 0.05),
-    delta = seq(0, 2, by = 0.05),
-    eq_band = seq(0.1, 0.5, by = 0.01),
-    prior_scale = c(1 / sqrt(2), 1, sqrt(2))
-  )
-
-# rope_options <- rope_options[order(rope_options[, 1], rope_options[, 2]), ]
-
-rope_options$iterate <- 1:nrow(rope_options)
-
 # Read the outputs of all iterations
 rope_data <- 
   tibble(filename = list.files(path = "./rope-res/", pattern = ".rds")) %>%
@@ -22,15 +9,20 @@ rope_data <-
   unnest(data) %>% 
   mutate(result = map(data, "result"),
          error = map(data, "error"),
+         params = map(data, "params"),
          error_message = map_chr(error, ~ pluck(.x, "message", .default = NA_character_)),
          result_not_null = if_else(map_lgl(result, ~ !is.null(.x)),
                                    1L,
                                    0L),
          n1 = map_dbl(result, ~ pluck(.x, "n1", .default = NA_real_)),
-         npower = map_dbl(result, ~ pluck(.x, "npower", .default = NA_real_)),
-         iterate = as.integer(names(data))) %>% 
+         tpr_out = map_dbl(result, ~ pluck(.x, "tpr_out", .default = NA_real_)),
+         tpr = map_dbl(params, ~ pluck(.x, "tpr", .default = NA_real_)),
+         eq_band = map_dbl(params, ~ pluck(.x, "eq_band", .default = NA_real_)),
+         delta = map_dbl(params, ~ pluck(.x, "delta", .default = NA_real_)),
+         prior_scale = map_dbl(params, ~ pluck(.x, "prior_scale", .default = NA_real_))
+         ) %>% 
   select(-file, -filename) %>% 
-  left_join(., rope_options, by = "iterate")
+  rowid_to_column(var = "iterate")
 
 # Calculate the number of missing results
 rope_data %>% 
@@ -40,66 +32,11 @@ rope_data %>%
 rope_data %>% 
   count(error_message)
 
-# Create the ouput table
-# rope_data <-
-#   rope_data %>% 
-#   select(iterate, power, delta, band, n1, npower) %>% 
-#   arrange(iterate)
-
-rope_result <-
[…]

### 2021-09-17 inst/manuscript/manuscript.Rmd
small updates in the manuscript

@@
   - name          : "Balazs Aczel"
-    affiliation   : "1"
+    affiliation   : "2"
 
@@
   - id            : "1"
+    institution   : "Doctoral School of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary"
+  - id            : "2"
     institution   : "Institute of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary"
-  - id            : "2"
-    institution   : "Doctoral School of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary"
   - id            : "3"
@@
 In order to estimate the sample size, we used the interval equivalent Bayes factor [@morey_bayes_2011; @van_ravenzwaaij_bayes_2019] method.
-We used a Cauchy prior distribution centered on zero with a scale of 1/$\sqrt{2}$.
+We used a Cauchy prior distribution centered on 0 with a scale parameter of 1/$\sqrt{2}$.
 We set the aimed TPR at 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement].
@@
 
-The present method calculates the corresponding default Bayes factor for a *t*-test statistic with Cauchy prior distribution centered on zero with scale parameter of either 1/$\sqrt{2}$, 1, or $\sqrt{2}$ for several sample sizes [the so-called Jeffrey-Zellner-Siow Bayes factor, see e.g., @rouder_bayesian_2009].
+The present method calculates the corresponding default Bayes factor for a *t*-test statistic with Cauchy prior distribution centered on 0 with scale parameter of either 1/$\sqrt{2}$, 1, or $\sqrt{2}$ for several sample sizes [the so-called Jeffrey-Zellner-Siow Bayes factor, see e.g., @rouder_bayesian_2009].
 The function returns the optimal sample size needed to reach the TPR for a given Bayes factor threshold to detect an expected population effect size.
@@
 We used the Jeffrey-Zellner-Siow Bayes factor method to estimate the sample size.
-We used a Cauchy prior distribution centered on zero with a scale of 1/$\sqrt{2}$.
+We used a Cauchy prior distribution centered on 0 with a scale parameter of 1/$\sqrt{2}$.
 We set the aimed TPR at 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement].
@@
 We used the BFDA method to estimate the sample size.
-We used a Cauchy prior distribution centered on zero with a scale of 1/$\sqrt{2}$.
+We used a Cauchy prior distribution centered on 0 with a scale parameter of 1/$\sqrt{2}$.
 We set the aimed TPR at 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement].
@@
 In order to estimate the sample size, we used the Region of Practical Equivalence [@kruschke_rejecting_2018] method.
-We used a Cauchy prior distribution centered on zero with a scale of 1/$\sqrt{2}$.
+We used a Cauchy prior distribution centered on 0 with a scale parameter of 1/$\sqrt{2}$.
 We set the aimed TPR at 0.8, because [1) it is the common standard in the field; 2) it is the journal publishing requirement].

### 2021-09-17 inst/manuscript/manuscript.tex
small updates in the manuscript

@@
   pdftitle={SampleSizePlanner: A Tool to Estimate and Justify Sample Size for Two-Group Studies},
-  pdfauthor={Marton Kovacs1,2, Don van Ravenzwaaij3, Rink Hoekstra3, \& Balazs Aczel1},
+  pdfauthor={Marton Kovacs1,2, Don van Ravenzwaaij3, Rink Hoekstra3, \& Balazs Aczel2},
   pdflang={en-EN},
@@
 \title{SampleSizePlanner: A Tool to Estimate and Justify Sample Size for Two-Group Studies}
-\author{Marton Kovacs\textsuperscript{1,2}, Don van Ravenzwaaij\textsuperscript{3}, Rink Hoekstra\textsuperscript{3}, \& Balazs Aczel\textsuperscript{1}}
+\author{Marton Kovacs\textsuperscript{1,2}, Don van Ravenzwaaij\textsuperscript{3}, Rink Hoekstra\textsuperscript{3}, \& Balazs Aczel\textsuperscript{2}}
 \date{}
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Institute of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary\\\textsuperscript{2} Doctoral School of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary\\\textsuperscript{3} University of Groningen, Groningen, The Netherlands}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Doctoral School of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary\\\textsuperscript{2} Institute of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary\\\textsuperscript{3} University of Groningen, Groningen, The Netherlands}
 
@@
 In order to estimate the sample size, we used the interval equivalent Bayes factor (Morey \& Rouder, 2011; Ravenzwaaij, Monden, Tendeiro, \& Ioannidis, 2019) method.
-We used a Cauchy prior distribution centered on zero with a scale of 1/\(\sqrt{2}\).
+We used a Cauchy prior distribution centered on 0 with a scale parameter of 1/\(\sqrt{2}\).
 We set the aimed TPR at 0.8, because {[}1) it is the common standard in the field; 2) it is the journal publishing requirement{]}.
@@
 
-The present method calculates the corresponding default Bayes factor for a \emph{t}-test statistic with Cauchy prior distribution centered on zero with scale parameter of either 1/\(\sqrt{2}\), 1, or \(\sqrt{2}\) for several sample sizes (the so-called Jeffrey-Zellner-Siow Bayes factor, see e.g., Rouder, Speckman, Sun, Morey, \& Iverson, 2009).
+The present method calculates the corresponding default Bayes factor for a \emph{t}-test statistic with Cauchy prior distribution centered on 0 with scale parameter of either 1/\(\sqrt{2}\), 1, or \(\sqrt{2}\) for several sample sizes (the so-called Jeffrey-Zellner-Siow Bayes factor, see e.g., Rouder, Speckman, Sun, Morey, \& Iverson, 2009).
 The function returns the optimal sample size needed to reach the TPR for a given Bayes factor threshold to detect an expected population effect size.
@@
 We used the Jeffrey-Zellner-Siow Bayes factor method to estimate the sample size.
-We used a Cauchy prior distribution centered on zero with a scale of 1/\(\sqrt{2}\).
+We used a Cauchy prior distribution centered on 0 with a scale parameter of 1/\(\sqrt{2}\).
 We set the aimed TPR at 0.8, because {[}1) it is the common standard in the field; 2) it is the journal publishing requirement{]}.
@@
 We used the BFDA method to estimate the sample size.
-We used a Cauchy prior distribution centered on zero with a scale of 1/\(\sqrt{2}\).
+We used a Cauchy prior distribution centered on 0 with a scale parameter of 1/\(\sqrt{2}\).
 We set the aimed TPR at 0.8, because {[}1) it is the common standard in the field; 2) it is the journal publishing requirement{]}.
@@
 In order to estimate the sample size, we used the Region of Practical Equivalence (Kruschke, 2018) method.
-We used a Cauchy prior distribution centered on zero with a scale of 1/\(\sqrt{2}\).
+We used a Cauchy prior distribution centered on 0 with a scale parameter of 1/\(\sqrt{2}\).
 We set the aimed TPR at 0.8, because {[}1) it is the common standard in the field; 2) it is the journal publishing requirement{]}.

### 2023-06-19 README.Rmd
updated logo, documentation, prior arguments added for Bayesian methods,  rearranged informed t test functions in R folder
updated logo path
added citation cff and updated readme with citations

@@
 
-# SampleSizePlanner
+# SampleSizePlanner <a href='https://marton-balazs-kovacs.github.io/SampleSizePlanner/'><img src='man/figures/ssp_logo.png' align="right" height="200" /></a>
 
@@
 * Region of Practical Equivalence (ROPE)
+* Non-inferiority Bayes factor
 
@@
 
-## Note
+The tutorial paper for two-group study designs can be found [HERE](https://journals.sagepub.com/doi/pdf/10.1177/25152459211054059).
 
-The tutorial paper that presents the web application and R package is not reviewed yet. More details are available in Kovacs, van Ravenzwaaij, Hoekstra & Aczel (2021).
+## Citation
+
+Please cite `SampleSizePlanner` if you use it for determining your sample size. To cite SampleSizePlanner the software, use:
+
+```{r echo = FALSE, results = "asis"}
+print(citation("SampleSizePlanner"), style = "text")
+```
+
+Or copy the reference information to your BibTeX file:
+
+~~~bibtex
+
+```{r echo = FALSE, results = "asis"}
+print(citation("SampleSizePlanner"), style = "Bibtex")
+```
+
+~~~
+
+To cite the tutorial paper on two-group designs use:
+
+Kovacs, M., van Ravenzwaaij, D., Hoekstra, R., & Aczel, B.
+(2022).
+SampleSizePlanner: A tool to estimate and justify sample size for two-group studies.
+_Advances in Methods and Practices in Psychological Science, 5_(1), 25152459211054059.
 

### 2024-11-27 inst/app/www/about.Rmd
Save local changes

@@
 
-The **SampleSizePlanner** shiny application enables researchers to calculate the sample size with 9 different methods in app and provides code to use an additional method from R.
+The **SampleSizePlanner** shiny application enables researchers to calculate the sample size with 13 different methods in app and provides code to use an additional method from R.
 The aim of the app is to make the various sample size calculating methods available for a wide range of researchers and propagate the transparent reporting of the chosen method.
@@
 -   Balazs Aczel
+-   Andrew Setiono
+-   Sebastian Lortz
 

### 2024-11-27 R/extract_rope_anova.R

@@ -32,13 +32,19 @@ extract_rope_anova <- function(pre_data, mu_ui, sigma_ui, eq_band_ui, tpr_ui, ci
     sorted_mu <- sort_mu(mu_scaled)[[1]]
     effect_swap <- sort_mu(mu_scaled)[[2]]
     
+    # convert effect A to 1 and B to 2
+    if (effect_ui == "Main Effect A") {effect_ui = "Main Effect 1"}
+    if (effect_ui == "Main Effect B") {effect_ui = "Main Effect 2"}
     
     # account for effect swap in sorting
-    if (effect_ui == "Main Effect 1" & effect_swap) {
-      effect_extract <- "Main Effect 2"} else if 
-    (effect_ui == "Main Effect 2" & effect_swap) {
-      effect_extract <- "Main Effect 1"} else {
-        effect_extract <- effect_ui}
+    if (effect_ui == "Main Effect 1" && effect_swap) {
+      effect_extract <- "Main Effect 2"
+    } else if (effect_ui == "Main Effect 2" && effect_swap) {
+      effect_extract <- "Main Effect 1"
+    } else {
+      effect_extract <- effect_ui
+    }
+    
     
     # Filter the results
     result_filter <- rope_anova_data %>%

### 2024-11-27 R/jusfitication.R

@@ -107,15 +107,16 @@ justification <- function(method, output_parameters) {
       glue::glue(
         "We used a power analysis to estimate the sample size. We used an alpha of 0.05. We set the aimed power at {tpr}, because {tpr_justification}. \\
         We set the expected group means to {mu}, which correspond to the subgroups 1|1, 1|2, 2|1, and 2|2 of factors A|B respectively. \\
-        We set the standard deviation of each group to {sigma}.Based on these parameters, a minimal sample size of {n1} was estimated in order \\
+        We set the standard deviation of each group to {sigma}.Based on these parameters, a minimal sample size of {n1} per group is required in order \\
         to reach {round(tpr_out, 2)} power for {effect} in our design."
         ,
         n1 = output_parameters$n1,
         tpr_out = output_parameters$tpr_out,
         tpr = output_parameters$tpr,
         tpr_justification = output_parameters$tpr_justification,
-        # delta = output_parameters$delta,
-        # delta_justification = output_parameters$delta_justification
+        mu = paste(output_parameters$mu, collapse = ", "),
+        sigma = output_parameters$sigma,
+        effect = output_parameters$effect
       )
     
   } else if (method == "bayesian-twoway-anova") {
@@ -125,55 +126,62 @@ justification <- function(method, output_parameters) {
         for evidence of {thresh}. We used a Cauchy prior distribution centered on 0 with a scale parameter of {prior_scale}. We set the \\
         aimed true positive rate at {tpr}, because {tpr_justification}. We set the expected group means to {mu}, which correspond to the \\
         subgroups 1|1, 1|2, 2|1, and 2|2 of factors A|B respectively. We set the standard deviation of each group to {sigma}. Based on these \\
-        parameters, a minimal sample size of {n1} was estimated to reach a {round(tpr_out, 2)} true positive rate for {effect} in our design.",
+        parameters, a minimal sample size of {n1} per group is required to reach a {round(tpr_out,2)} true positive rate for {effect} in our design.",
         thresh = output_parameters$thresh,
         n1 = output_parameters$n1,
         tpr_out = output_parameters$tpr_out,
         tpr = output_parameters$tpr,
         tpr_justification = output_parameters$tpr_justification,
-        # delta = output_parameters$delta,
-        # delta_justification = output_parameters$delta_justification
+        prior_scale = output_parameters$prior_scale,
[…]

### 2024-11-27 R/mod_ssp_bayesian_anova.R

@@ -40,7 +40,7 @@ mod_ssp_bayesian_anova_ui <- function(id) {
           name_with_info(
             "Which effect's power you want to detect?",
             "Determine which effect of the ANOVA analysis, in which you want to check for true positive rate"),
-          c("Main Effect 1", "Main Effect 2", "Interaction Effect")),
+          c("Main Effect A", "Main Effect B", "Interaction Effect")),
         ## Input Mean for Each Group
         shinyMatrix::matrixInput(
           NS(id, "muMatrix"),
@@ -208,7 +208,11 @@ mod_ssp_bayesian_anova_server <- function(id) {
         # delta_justification = input$delta_justification,
         n1 = pre_result()$pre_n1[[1]],
         tpr_out = pre_result()$pre_tpr_out[[1]],
-        thresh = input$thresh
+        thresh = input$thresh,
+        prior_scale = input$prior_scale,
+        mu = pre_result()$pre_mu,
+        sigma = input$sigma,
+        effect = input$effect
       )
     })
     

### 2024-11-27 R/mod_ssp_eq_anova.R

@@ -16,7 +16,7 @@
 mod_ssp_eq_anova_ui <- function(id) {
   tagList(
     # Method
-    h1("Interval equivalence Bayes factor - Two-way ANOVA"),
+    h1("Interval Equivalence Bayes Factor - Two-way ANOVA"),
     sidebarLayout(
       sidebarPanel(
         # Panel title
@@ -223,12 +223,17 @@ mod_ssp_eq_anova_server <- function(id) {
       list(
         tpr = input$tpr,
         tpr_justification = input$tpr_justification,
+        eq_band_justification = input$eq_band_justification,
         # delta = input$delta,
         # delta_justification = input$delta_justification,
         n1 = pre_result()$pre_n1[[1]],
         tpr_out = pre_result()$pre_tpr_out[[1]],
         thresh = input$thresh,
         eq_band = input$eq_band,
+        prior_scale = input$prior_scale,
+        mu = pre_result()$pre_mu,
+        sigma = input$sigma,
+        effect = input$effect
       )
     })
     

### 2025-02-17 inst/app/www/about.Rmd

@@
 
+### Publications
+
+The sample size calculations for the t-tests are presented in a tutorial paper: [Kovacs, van Ravenzwaaij, Hoekstra & Aczel (2022)](https://doi.org/10.1177/25152459211054059)
+
 ### Contributors
@@
 
-The design of the app was done by Zsuzsa Szekely [(@szekely_Zsuzsa)](<https://twitter.com/szekely_zsuzsa>).
+The design of the app was done by Zsuzsa Szekely [(\@szekely_Zsuzsa)](https://twitter.com/szekely_zsuzsa).
 
@@
 
-If you have any questions or comments please write to Marton Kovacs to the following email address: [marton.balazs.kovacs@gmail.com](mailto:marton.balazs.kovacs@gmail.com)
+If you have any questions or comments please write to Marton Kovacs to the following email address: [marton.balazs.kovacs\@gmail.com](mailto:marton.balazs.kovacs@gmail.com)
 

### 2025-02-17 README.Rmd
modified pkgdown config

@@
 
-# SampleSizePlanner <a href='https://marton-balazs-kovacs.github.io/SampleSizePlanner/'><img src='man/figures/ssp_logo.png' align="right" height="200" /></a>
+# SampleSizePlanner <a href='https://marton-balazs-kovacs.github.io/SampleSizePlanner/'><img src='man/figures/ssp_logo.png' align="right" height="200" alt-text="SampleSizePlanner"/></a>
 
